### 2020-01-21 abstract.tex
start on a paper

Many programming languages offer multiple static analysis tools that offer to detect faults in code witthout executing it.  Understanding the strengths and weaknesses of tools, and performing direct comparisons of their effectiveness is diffcult; it usually involves either manual examination of differing warnings on real code, or the bias-prone construction of artificial test cases.  In practice, comparisons tend to be limited to superficial, anecdotal discussions in the informal literature (e.g., blog posts by software developers), or purely research-community-oriented evaluations made by the authors of new tools seeking to publish their results.  This paper proposes a novel automated approach to comparing static analysis tools, based on producing \emph{mutants} of real code, and comparing mutation detection rates for tools to their warning rates on the original code.  In addition to making tool differences quantitatively observable without extensive manual effort, this approach offers a new way to detect and fix omissions in a static analysis tool's set of detectors.  We present an extensive comparison of three well-known Solidity smart contract static analysis tools, and show how using an automatic prioritization of our results allowed us to add three new detectors to the best of the tools.  We also evaluate popular Java and Python static analysis tools and discuss their strengths and weaknesses.

### 2020-01-21 experiments.tex
start on a paper

\section{Comparisons of Static Analysis Tools}

\subsection{Solidity Smart Contract Tools}

\subsection{Java Tools}

\subsection{Python Tools}

\subsection{Threats to Validity}

### 2020-01-21 introduction.tex
start on a paper

\section{Introduction}

Static analysis of code is one of the most effective ways to avoid defects in software.  Static analysis can find problems that are extremely hard to detect by testing, e.g. when the inputs triggering a bug are

Mutation testing \cite{demillo1978hints,budd1980theoretical} (or mutation analysis, the term we will use in this work, for reasons that will become clear) uses small syntactic changes to a program to introduce synthetic ``faults,'' under the assumption that if the original version of a program is (mostly) correct, most small changes will therefore introduce a fault.  For the most part, mutation analysis has been used to evaluate test suites by computing a score (the ratio of mutants the suite detects, or ``kills'').  Most such use has been in research efforts, rather than practical testing efforts, though there has been sporadic use by interested developers.
In an ASE 2015 \cite{groce2015verified} paper and a 2018 journal extension \cite{groce2018verified} of that paper, Groce et al. proposed examining individual mutants that survive a formal verification or automated test generation process to detect and correct weaknesses in a specification or test generator.  The approach was able to expose bugs in a heavily-tested module of the Linux kernel \cite{mutKernel} and improve a heavily used test generator for the {\tt pyfakefs} file system.  Recently, mutation analysis has been adopted in industrial settings, though not for actual examination of all surviving mutants \cite{MutGoogle,ivankovic2018industrial}, a practice that is hard to scale to large code bodies.

\subsection{Differential Mutation Analysis}

We can say that a static analysis tool kills a mutant when the number of (non-informational or optimization related) warnings or errors produced with respect to the code \emph{increases} for the mutated version, compared to the original code.  This difference is most informative and easily interpreted when the original code produces no warnings or errors (it is ``clean''); for non-clean code, a tool conceivably could detect the mutant, but only change a previously generated warning, not add an additional warning, leading to an underestimate of effectiveness.  However, even for non-clean code, most detected mutants will produce a new warning.

The value of the differential comparison lies in a few key points.  First, this is a measure that does not reward a tool that produces too many false positives.  The tool cannot simply flag all code as having a problem or it will perform poorly at the task of \emph{distinguishing} the mutated code from non-mutated (and presumably at least \emph{more} correct) code.  Based on verification and testing uses of mutation, it is safe to say that at least 50, and likely 60-70\% or more, of mutants that are not semantically equivalent to the original code are actually fault-inducing, so the task presented to a static analysis tool is the generalization of the task we ideally expect static analysis to perform:  to identify faulty code, without executing it, and, most critically, \emph{to distinguish faulty from correct code}.  Obviously, many faults cannot be identified statically without a complete specification, or without unreasonable analysis cost and precision, but the measure of performance here is meant to be mostly \emph{relative} to other tools applied to the same code; this is primarily a \emph{differential} approach.

Second, and critically, this is an \emph{automatable} method that can provide an evaluation of static analysis tools over a large number of target source code files, without requiring human effort to classify results as real bugs or false positives.  It is not clear that any other fully automatic method is competitively meaningful; it is possible that methods based on code changes from version control provide some of the same benefits, but these require classification of changes into bug-fixes and non-bug-fixes, and of course require version control history.  Also, history-based methods will be biased towards precisely those faults humans (or tools) were able to detect and fix.

It is the combination of differential comparison and mutation that is key.  Differential comparison of tools, alone, is not meaningful, without additional effort; na\"ive approaches will not work \cite{regehrRandom}.  Consider a comparison of the number of issues detected between two tools over a single program, or over a large set of programs.  If one tool emits more warnings and errors than another, it may mean that the tool is more effective at finding bugs; but it may also mean that it has a higher false positive rate.  Without human examination of the individual issues, it is impossible to be sure, or even (in cases where the tools are reasonably comparable) to make an informed guess.  Using mutants, however, provides a foreground to compare to this background.  In particular, for a large set of programs, the most informative result will be when 1) tool A reports fewer issues on average than tool B over the un-mutated programs but 2) tool A also detects more mutants.  This is strong evidence that A is simply better all-around than B; it likely has a lower false positive rate \emph{and} a lower false negative rate.  While it is not proof of this claim, it is hard to construct another plausible explanation for reporting \emph{fewer} issues on un-mutated contracts while still detecting \emph{more} mutants.  Other than having better precision and recall, how else could a tool effectively distinguish mutated from un-mutated code?

Finally, even when two tools have similar quantitative results, examining individual mutants killed by one tool but not by another allows us to understand strengths and weaknesses of the tools, in a context that makes comprehending the cause of the detection (or lack of it) easy: the difference between the un-mutated code and mutated code will always be small and relatively simple.  This leads to the idea of improving static analysis tools by examining mutants detected by another tool (thus known to be in-principle detectable) but not by the tool to be improved.  Of course, any faults in code, not just mutants, could serve this purpose.  But, again, the automatic nature of mutation generation, and the presumption that the mutation is indeed a fault, is useful.  Moreover, because mutants follow syntactic patterns, searching for similar mutants/faults the tool to be improved does not detect is much easier than with arbitrary faults, and can be partly automated.  Of course, knowing which mutation patterns are of interest requires human effort.  As with efforts to improve test suites, manually searching through all mutants can be an onerous task, especially for large-scale evaluations.  We therefore also introduce the idea of prioritizing mutants, using a Furthest-Point-First \cite{Gonzalez} algorithm and distance metric inspired by previous work on helping developers sort through failing test cases and avoid duplicates \cite{PLDI13}, to help static analysis tool developers find interesting patterns without wading through numerous uninteresting duplicates.

\subsection{A Simple Example}

\begin{figure}
{\scriptsize
\begin{code}
pragma solidity ^0.4.0;

contract SimpleStorage \{
    uint storedData;

    function set(uint x) public \{
        storedData = x;
    \}

    function get() public view returns (uint) \{
        return storedData;
    \}
  \}
\end{code}
}
\caption{A simple example Solidity smart contract from \url{https://solidity.readthedocs.io/en/v0.4.24/introduction-to-smart-contracts.html}.}
\label{fig:sol424intro}
\end{figure}

Consider the code in Figure \ref{fig:sol424intro}, from the Solidity 0.4.24 ``Introduction to Smart Contracts''.  The Universal Mutator tool \cite{universalmutator,regexpMut}, which has been extensively tuned for Solidity mutation (and is the only smart contract mutation tool referenced in the Solidity documentation (\url{https://solidity.readthedocs.io/en/v0.5.12/resources.html})) produces seven valid, non-redundant mutants for this trivial example code.  Both the public version of Slither \cite{slither} and SmartCheck \cite{smartcheck} produce a small number (three and two, respectively) of low-severity, informational, warnings for this code.  Both tools also detect (by increasing the number of warnings produced) four of the seven mutants.  However, only one of the mutants detected is common to both tools: both tools detect changing the {\tt return} statement in the {\tt get} function to a call to {\tt selfdestruct} the smart contract.  Slither, but not SmartCheck, also detects replacing the assignment of {\tt storedData} in {\tt set} with either a {\tt selfdestruct} or {\tt revert}, or simply removing it altogether.  SmartCheck, on the other hand, detects removing the {\tt return} in {\tt get} or replacing it with a {\tt revert}, or removing the {\tt public} visibility modifier for {\tt get} \footnote{Slither's ``missing return'' detector is only available in the private version of slither, or through the {\tt crytic} service provided by Trail of Bits.}.  If we restrict our analyis to issues with a severity greater than informational, SmartCheck detects no mutants of the contract, while Slither still reports that some mutants \emph{allow an arbitrary caller to cause the contract to self destruct}.  This simple example shows how our approach works on a small scale.  Using large numbers of larger, more realistic contracts makes it possible to extract the same kind of information, on a much larger scale.  Prioritization of mutants is not very useful here (ranking three mutant saves little effort), so we will show the utility of that approach in our full Solidity results.

\subsection{Contributions}

This paper offers the following contributions:

\begin{itemize}
\item We propose a differential approach to comparing static analysis tools based on the insight that program mutants provide an automated source of simple, easy-to-understand program changes that are likely faults.
\item We propose a definition of mutant killing that works well in a static analysis context.
\item We introduce a simple scheme for prioritizing mutants that helps users understand the results of analysis and guide efforts to improve, rather than simple compare, tools.
\item We apply our method to an extensive, in-depth comparison of three Solidity smart contract analysis tools, and show how prioritization allowed us to easily identify (and build) three new detectors for the most effective of these tools.
\item We further provide results for comparisons of popular Java and Python static analysis tools, showing the general usefulness of our methods, and giving a new picture of the comparative effectiveness of these tools.
\end{itemize}

### 2020-01-21 relatedwork.tex
start on a paper

\section{Related Work}

The goal of ``analysing the program analyser'' \cite{cadar2016analysing} and applying better automated methods to evaluate and improve analysis tools has become recently more popular and, we suspect, more possible.  The irony of using mostly ad-hoc, manual methods to test and understand static analysis tools is apparent; however, the fundamentally incomplete and heuristic nature of effective analysis tools makes this a challenge similar to testing machine learning algorithms \cite{OnlyOracle}; most tools will not produce ``the right answer'' all the time, by their very nature.  This is a result of both algorithmic constraints and basic engineering trade-offs.

Cuoq et al. \cite{regehrRandom} proposed the generation of programs to test analysis tools aiming for soundness, in limited circumstances, but noted that na\"ive differential testing of analysis tools was not possible.  This paper proposes a non-na\"ive differential comparison (not, exactly, differential testing, however, in that only aggregate results are possible to interpret without human intelligence), based on the observation that the ability to detect program mutants offers an automatable way to tell which of two tools is better (for a given universe of examples, at least) at telling faulty from non-faulty code.  Klinger et al. propose a different approach to differential testing of analysis tools \cite{klinger2019differentially}.  Their approach is in some ways similar to ours, in that it takes as input a set of seed programs, and compares results across new versions generated from that seed.  The primary differences are that their seed programs must be warning-free (which greatly limits the set of input programs available) and their tool must parse and understand the programs, and that the new versions are based on adding new assertions, not ``breaking'' the original code.  We allow arbitarily buggy seed programs (thus many more real programs can be used), and can, due to the any-language nature of the mutation generator we use, operate even in new languages without further development effort.  Further, their approach only identifies problems when tools are outliers compared to numerous other tools in either detecting a bug (precision) or not detecting it (soundness), and so requires comparing multiple tools.  Our approach has some utility for even a single tool (you can just examine prioritized un-detected mutants).  On the other hand, their approach can identify precision issues, while we offer no real help with false positives (in theory, you could apply their majority-vote method to mutants only a few tools flag, but mutants \emph{are} usually faults. in contrast to their introduction of checks that may be guaranteed to pass, so this is probably not very helpful).

### 2020-01-21 relatedwork.tex
csmith

@@
 Cuoq et al.
-\cite{regehrRandom} proposed the generation of programs to test analysis tools aiming for soundness, in limited circumstances, but noted that na\"ive differential testing of analysis tools was not possible.
+\cite{regehrRandom} proposed the generation of random programs (\emph{\'a la} Csmith \cite{csmith}) to test analysis tools aiming for soundness, in limited circumstances, but noted that na\"ive differential testing of analysis tools was not possible.
 This paper proposes a non-na\"ive differential comparison (not, exactly, differential testing, however, in that only aggregate results are possible to interpret without human intelligence), based on the observation that the ability to detect program mutants offers an automatable way to tell which of two tools is better (for a given universe of examples, at least) at telling faulty from non-faulty code.

### 2020-01-21 introduction.tex
much more on diff uses

@@
 
-Static analysis of code is one of the most effective ways to avoid defects in software.
+Static analysis of code is one of the most effective ways to avoid defects in software, and when security is a concern, applying effective static analysis tools is essential.
 Static analysis can find problems that are extremely hard to detect by testing, e.g.
-when the inputs triggering a bug are
+when the inputs triggering a bug are hard to find.
+Static analysis is also often more efficient than testing; a bug that takes a fuzzer like AFL days to find may be immediately identified by a good static analysis tool.
+
+Users of static analysis tools often wonder which of multiple tools available for a programming language are most effective, and, when using more than one tool is an option (e.g.
+with free tools), how much tools overlap in their results.
+Given the human effort required to read static analysis results, the latter can be an important question.
+If two tools find substantially different (non-false-positive) bugs, it is wise to use both if finding all bugs is important.
+On the other hand, if two tools are very similar in what they detect, the effort of wading through duplicate results may not be a good use of time.
+Developers of static analysis tools also want to be able to compare their tools to others targeting the same domain, in order to see what detection methods (or tweaks to precision/soundness tradeoffs) they might want to imitate.
+Unfortunately, comparing tools is hard, and would seem to require vast manual effort to inspect findings and determine ground truth, to obtain any statistical validity.
+
+Differential testing \cite{McKeeman,ICSEDiff} is a popular approach to comparing multiple software systems offering similar functionality, but the wide divergence of possible trade-offs, analysis focuses, and the prevalence of false positives in almost all analysis results makes na\"ive differential testing not applicable to static analysis tools \cite{regehrRandom}.
 
@@
 Recently, mutation analysis has been adopted in industrial settings, though not for actual examination of all surviving mutants \cite{MutGoogle,ivankovic2018industrial}, a practice that is hard to scale to large code bodies.
+
+Combining a differential approach (not differential \emph{testing}, precisely, in that individual differences are not always worth inspecting) and mutation analysis, however, offers a novel way to compare static analysis tools, one useful both to users wishing to select a good tool or set of tools, and to developers hoping to improve their own tools.
 
@@
 It is the combination of differential comparison and mutation that is key.
-Differential comparison of tools, alone, is not meaningful, without additional effort; na\"ive approaches will not work \cite{regehrRandom}.
+Differential comparison of tools, as noted above, is not really meaningful, without additional effort; na\"ive methods simply will not work \cite{regehrRandom}.
 Consider a comparison of the number of issues detected between two tools over a single program, or over a large set of programs.
@@
 Finally, even when two tools have similar quantitative results, examining individual mutants killed by one tool but not by another allows us to understand strengths and weaknesses of the tools, in a context that makes comprehending the cause of the detection (or lack of it) easy: the difference between the un-mutated code and mutated code will always be small and relatively simple.
-This leads to the idea of improving static analysis tools by examining mutants detected by another tool (thus known to be in-principle detectable) but not by the tool to be improved.
+Moreover, simply looking at how much two tools agree on mutants can answer the question of a user of static analysis tools: given that I am using tool A, would adding tool B be likely to add enough new, interesting results to make it worth my time to examine its output?
+When (rarely) one tool subsumes another in terms of mutants, it can be very clear that the tool whose killed mutants are all killed by another tool is likely strictly inferior.
+Interested users, e.g.
+security analystst, can even inspect the differences to get an idea of the particular cases when a tool might be most effective.
+
+Comparing mutant results also leads to the idea of improving static analysis tools by examining mutants detected by another tool (thus known to be in-principle detectable) but not by the tool to be improved.
 Of course, any faults in code, not just mutants, could serve this purpose.

### 2020-01-21 introduction.tex
more intro

@@
 
-Differential testing \cite{McKeeman,ICSEDiff} is a popular approach to comparing multiple software systems offering similar functionality, but the wide divergence of possible trade-offs, analysis focuses, and the prevalence of false positives in almost all analysis results makes na\"ive differential testing not applicable to static analysis tools \cite{regehrRandom}.
+Differential testing \cite{Differential,ICSEDiff} is a popular approach to comparing multiple software systems offering similar functionality, but the wide divergence of possible trade-offs, analysis focuses, and the prevalence of false positives in almost all analysis results makes na\"ive differential testing not applicable to static analysis tools \cite{regehrRandom}.
 

### 2020-01-21 experiments.tex
more

@@
-\section{Comparisons of Static Analysis Tools}
+\section{Comparison and Improvements of Static Analysis Tools}
 
 \subsection{Solidity Smart Contract Tools}
+
+\subsubsection{Smart Contracts and Smart Contract Static Analysis}
+
+Smart contracts are autonomous code instruments, usually operating on a blockchain, that often have critical responsibilities such as facilitating and verifying (large) financial services transactions, tracking high-value physical goods or intellectual property, or even controlling ``decentralized organizations'' with multifarious aspects.
+Security and correctness are thus critical in the smart contract domain, and static analysis is a key way to ensure allocation of high-value resources is not compromised.
+The most popular smart contract platform, by far, is the Ethereum blockchain, and the Solidity smart contract language \cite{buterin2013whitepaper,wood2014yellow}; the Ethereum cryptocurrency has a market capitalization as we write of over \$15 billion dollars, largely fueled by interest in the smart contract functionality.
+Ethereum contracts have been the targets of widely publicized attacks, with large financial consequences  \cite{spank,DAO}.
+A recent paper examining results from 23 professional security audits of Solidity contracts argues that effective static analysis is a major key to avoiding such disasters in the future \cite{FC20}.
+
+We analyzed three well-known tools for static analysis of Solidity smart contracts: Slither \cite{slither}, SmartCheck \cite{smartcheck}, and Securify \cite{Securify}.
+
+\subsubsection{Original Program Selection}
+
+\subsubsection{Analysis Results}
+
+We first compare the two tools that performed best.
+Slither detected 2.37 mean issues over the 100 contracts, and found 39 contracts free of all issues.
+Slither had a mean mutation score of 0.09 over all contracts, and 0.11 over clean (according to Slither) contracts.
+SmartCheck detected 1.89 mean issues over the 100 contracts, but only found 27 contracts free of all issues.
+The mean issue counts for the tools were not statistically significantly different ($p=0.34$).
+SmartCheck had a mean mutation score of 0.05 over all contracts, and 0.03 over clean (according to SmartCheck) contracts.
+The ratio between mean issues (Slither / SmartCheck) is 1.25, and the ratio of numbers of clean contracts is similar (1.44).
+The mutation score ratio, however, is 1.8, suggesting that Slither is not out-scoring SmartCheck \emph{only} because it produces more false positives.
+Slither detects 3,520 mutants not detected by SmartCheck, while SmartCheck only detects 1,629 mutants not detected by Slither (of these, 1,256 were also not detected by Securify) .
+Over the 7 contracts that both tools mark as clean, Slither has a mean mutation score of 0.08, and SmartCheck has a mean mutation score of 0.03; the difference, even with only 7 contracts, is statistically significant by paired Wilcoxon test, at $p=0.03$.
+
+Securify was an outlier, compared to the other tools.
+First, while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick, Securify often required hours to analyze a contract; the full analysis required over a month.
+Second, Securify identified a mean of 24.65 issues on the un-mutated contracts, with a median of 17 issues found; this was a significant difference in issue counts from either other tool, with $p<1.0^{-15}$ in both cases.
+Despite this huge number of non-informational warnings for the original contracts, the mean mutation score was even worse than for SmartCheck, 0.03.
+For contracts where Securify found no issues in the original contract, it \emph{never} detected any mutants.
+Securify, however, did detect 1,449 mutants not detected by Slither, 1,076 of which were not detected by SmartCheck.
+On the other hand, given the ratio of issue counts compared to both tools is over 10.0, it is fairly likely many of these \emph{are} false positives, despite the probability mutants are actual faults.
+
+Our recommendation, based on mutation analysis results, is that smart contract analysis should at least use both Slither and SmartCheck, perhaps applying only Slither until it reports no interesting issues.
+Securify does offer a substantial non-overlapping set of issues reported, however, but produces a very large number of false positives.
+Available human resources and total number of issues produced (as well as high runtime) might justify not using Securify in some cases.
+
+\subsubsection{Improving Slither}
 

### 2020-01-21 analyzejava.py

@@ -23,6 +23,7 @@ with open("Java.csv") as javaResults:
             if row["spotbugs_killed"] == "0":
                 PMD_not_spotbugs += 1
         if row["Infer_killed"] == "1":
+            print(row)
             Infer_killed += 1
         try:
             spotbugs_warnings += int(row["Spotbug"])

### 2020-01-22 experiments.tex
RQs
more on RQ1

@@
-\section{Comparison and Improvements of Static Analysis Tools}
+\section{Experimental Results: Comparison and Improvement of Static Analysis Tools}
+
+For each language, RQs:
+
+RQ1:  actionable, novel differences for tool ranking?
+in particular, does the rate of mutant detection differ from what would be expected by extrapolating from issue counts over original code; that is, is the tendency of mutants to represent faults/bad code useful, and avoids the problem of simply rewarding false positive-prone tools?
+RQ2:  tool overlaps identifiable?
+RQ3:  agrees with other sources of information?
+RQ4:  can identify improvements?
 

### 2020-01-22 analyzejava.py
diagrams

@@ -10,19 +10,26 @@ PMD_warnings = 0
 Infer_warnings = 0
 N = 0.0
 
+spotbugs_file = open("all.spotbugs.killed.txt", 'w')
+PMD_file = open("all.PMD.killed.txt", 'w')
+Infer_file = open("all.Infer.killed.txt", 'w')
+
 with open("Java.csv") as javaResults:
     reader = csv.DictReader(javaResults)
     for row in reader:
         N += 1
         if row["spotbugs_killed"] == "1":
             spotbugs_killed += 1
+            spotbugs_file.write(row["Mutant"] + "\n")
             if row["PMD_killed"] == "0":
                 spotbugs_not_PMD += 1
         if row["PMD_killed"] == "1":
             PMD_killed += 1
+            PMD_file.write(row["Mutant"] + "\n")
             if row["spotbugs_killed"] == "0":
                 PMD_not_spotbugs += 1
         if row["Infer_killed"] == "1":
+            Infer_file.write(row["Mutant"] + "\n")            
             print(row)
             Infer_killed += 1
         try:

### 2020-01-22 experiments.tex
more data, diagrams, RQs

--- 2020-01-21 experiments.tex
@@
-\section{Comparison and Improvements of Static Analysis Tools}
+\section{Experimental Results: Comparison and Improvement of Static Analysis Tools}
+
+Our primary experimental results are a set of comparisons of tools using our method, for three languages: Solidity (the most popular language for smart contracts), Java, and Python.
+We also use these results to answer a set of research questions that consider the utility of our method:
+
+\begin{itemize}
+\item {\bf RQ1:}  Does our approach produce actionable results?
+That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar in terms of mutation detection capability?
+\item {\bf RQ2:}  Does our approach provide additional information beyond simply counting findings for the original, un-mutated analyzed code?
+Do \emph{ratios} differ between programs, or does the number of mutants killed simply reflect the ``verbosity'' of each tool?
+\item {\bf RQ3:}  Do the rankings that raw kills and ratios establish agree with other sources of information about the effectiveness of the evaluated tools?
+Can we confirm results from other evaluation methods, or informal opinion?
+\item {\bf RQ4:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that different tools are good at/bad at?
+Can we extend this information to improve tools by addressing identified weaknesses?
+  \end{itemize}
+
+In particular, we consider {\bf RQ2} to be of critical importance; if the mutant ratios for tools differ, then this is clear evidence that our hypothesis that the tendency of mutants to be faults, and to expect that mutated code will, by a sound and precise tool, be flagged as problematic more often than non-mutated code, holds.
+This expectation that (some subset of the) mutants can serve as proxies for real, detectable faults is the core concept of our approach.
 
@@
 \subsubsection{Analysis Results}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{solidity.png}
+  \caption{Mutants killed by Solidity static analysis tools.}
+  \label{fig:solidityvenn}
+\end{figure}
+
+\begin{table}
+  \begin{tabular}{l|r|r|r|r|r}
+    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
+    Tool & Mean & Median & Mean & Median & Ratio\\
+    \hline
+    \hline
+    Slither & 2.37 & 1.0 & 0.09 & 0.09 & 0.038 \\
+    \hline
+    SmartCheck & 1.89 & 1.0 & 0.05 & 0.05 & 0.026 \\
+    \hline
+    Securify & 24.65 & 17.0 & 0.03 & 0.02 &  0.001 \\
+    \hline
+  \end{tabular}
+  \caption{Solidity tools: number of findings and mutation scores over all contracts.}
+  \label{tab:scoresolidity}
+\end{table}
+
+\begin{table}
+  \begin{tabular}{l|r|r|r|r|r}
+    & & \multicolumn{2}{|c|}{} & \multicolumn{2}{|c}{Clean For All (3)} \\
+    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} &  \multicolumn{2}{|c}{Mutation Score}\\
+    Tool & Contracts & Mean & Median & Mean & Median\\
+    \hline
+    \hline
+    Slither & 39 & 0.11 & 0.11 & 0.09 & 0.09 \\
+    \hline
+    SmartCheck & 27 & 0.03 & 0.01 & 0.03 & 0.00 \\
+    \hline
+    Securify & 5 & 0.00 & 0.00 & 0.00 & 0.00 \\
+    \hline
+  \end{tabular}
+  \caption{Solidity tools: clean contract counts and mutation scores.}
+  \label{tab:cleansolidity}
+\end{table}
+
+Figure \ref{fig:solidityvenn} shows the mutants killed by the Solidity analysis tools.
+Tables \ref{tab:scoresolidity} and \ref{tab:cleansolidity} provide numeric details of the results, including the \emph{ratio} for each tool, adjusting its mutation scores by its' general tendency to produce findings.
+First, a user examining these results would suspect that Slither and SmartCheck are both useful tools, and should likely both be applied in a high-risk security-sensitive context like smart contract development.
+Second, a user might suspect that the large number of findings produced, and smaller number of mutants killed, for Securify, mean that including Securify in the static analysis tool stable is a more difficult decision.
+On the one hand, Securify does detect nearly as many mutants it alone can identify as SmartCheck.
+The large number of findings, and very bad mutant ratio, however, lead us to suspect that many of these ``detected'' mutants are false positives (or, at least, that the problem is not the one Securify identifies).
+Extracting the signal from Securify's noise will be difficult.
+We also note that while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick (it took about 8 days sequential compute time for Slither and 3 days for SmartCheck), Securify often required many hours to analyze a mutant, and frequently required a few days to analyze a mutant; the full analysis required over three months of compute time.
 
@@
 
+\subsubsection{Original Program Selection}
+
+\subsubsection{Analysis Results}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{java.png}
+  \caption{Mutants killed by Java static analysis tools.}
+  \label{fig:javavenn}
+\end{figure}
+
+\begin{table}
+  \begin{tabular}{l|r|r|r|r|r}
+    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & \\
+    Tool & Mean & Median & Mean & Median & Ratio\\
+    \hline
+    \hline
+    SpotBugs & 0.26 & 0.00 & 0.30 & 0.07 & 1.132 \\
+    \hline
+    PMD & 0.48 & 0.00 & 0.09 & 0.00 & 0.183 \\
+    \hline
+    Infer & 0.00 & 0.00 & 0.00 & 0.00 &  0.006 \\
+    \hline
+  \end{tabular}
+  \caption{Java tools: number of findings and mutation scores over all files.}
+  \label{tab:scorejava}
+\end{table}
+
+\begin{table}
+  \begin{tabular}{l|r|r|r|r|r}
+    & & \multicolumn{2}{|c|}{} & \multicolumn{2}{|c}{Clean For All (1,185)} \\
+    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} &  \multicolumn{2}{|c}{Mutation Score}\\
+    Tool & Files & Mean & Median & Mean & Median\\
+    \hline
+    \hline
+    SpotBugs & 1,426 & 0.32 & 0.05 & 0.37 & 0.06 \\
+    \hline
+    PMD & 1,359 & 0.08 & 0.00 & 0.08 & 0.00 \\
+    \hline
+    Infer & 1,598 & 0.00 & 0.00 & 0.00 & 0.00 \\
+    \hline
+  \end{tabular}
+  \caption{Java tools: clean file counts and mutation scores.}
+  \label{tab:cleanjava}
+\end{table}
+
+Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools.
+
 \subsection{Python Tools}

### 2020-01-22 introduction.tex
more data, diagrams, RQs

@@
 Obviously, many faults cannot be identified statically without a complete specification, or without unreasonable analysis cost and precision, but the measure of performance here is meant to be mostly \emph{relative} to other tools applied to the same code; this is primarily a \emph{differential} approach.
+In other words, a key notion is that while most mutants cannot be detected statically, the ones that are tend to be \emph{true positives}: if they were real code changes, they would be faults.
 
@@
 Differential comparison of tools, as noted above, is not really meaningful, without additional effort; na\"ive methods simply will not work \cite{regehrRandom}.
-Consider a comparison of the number of issues detected between two tools over a single program, or over a large set of programs.
+Consider a comparison of the number of findings between two tools over a single program, or over a large set of programs.
 If one tool emits more warnings and errors than another, it may mean that the tool is more effective at finding bugs; but it may also mean that it has a higher false positive rate.
-Without human examination of the individual issues, it is impossible to be sure, or even (in cases where the tools are reasonably comparable) to make an informed guess.
+Without human examination of the individual findings, it is impossible to be sure, or even (in cases where the tools are reasonably comparable) to make an informed guess.
 Using mutants, however, provides a foreground to compare to this background.
-In particular, for a large set of programs, the most informative result will be when 1) tool A reports fewer issues on average than tool B over the un-mutated programs but 2) tool A also detects more mutants.
+In particular, for a large set of programs, the most informative result will be when 1) tool A reports fewer findings on average than tool B over the un-mutated programs but 2) tool A also detects more mutants.
 This is strong evidence that A is simply better all-around than B; it likely has a lower false positive rate \emph{and} a lower false negative rate.
-While it is not proof of this claim, it is hard to construct another plausible explanation for reporting \emph{fewer} issues on un-mutated contracts while still detecting \emph{more} mutants.
+While it is not proof of this claim, it is hard to construct another plausible explanation for reporting \emph{fewer} findings on un-mutated contracts while still detecting \emph{more} mutants.
 Other than having better precision and recall, how else could a tool effectively distinguish mutated from un-mutated code?
 
-Finally, even when two tools have similar quantitative results, examining individual mutants killed by one tool but not by another allows us to understand strengths and weaknesses of the tools, in a context that makes comprehending the cause of the detection (or lack of it) easy: the difference between the un-mutated code and mutated code will always be small and relatively simple.
+We can quantitatively express this relationship between detecting mutants and findings for the original program.
+Since we are comparing over the same program or set of programs, we simply divide the mean \emph{mutation score} ($\frac{|\mathit{killed}||}{|\mathit{mutants}|}$, the ratio of killed mutants to all mutants) by the mean number of findings.
+This \emph {mutant ratio} tells us about the ability of a tool to produce findings \emph{for mutants}, relative to its tendency to produce findings in general (per line of code, per source file, etc.).
+If a tool has a tendency to produce large numbers of findings (compared to other tools), and this is paired with a tendency to detect most mutants, then the tool will not be penalized for producing many findings.
+Assuming that real faults are relatively rare in the original, un-mutated code, the best result (and best ratio score) will be for a tool that produces comparatively few findings for un-mutated code, but detects a larger portion of mutants than other tools; the worst result will be a tool that produces lots of findings, but detects few mutants.
+We will actually see some examples of the worst case in our results for real tools.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{example.png}
+  \caption{Mutants killed by three static analysis tools.}
+  \label{fig:examplevenn}
+\end{figure}
+
+Finally, even when tools have similar quantitative results, including similar ratios, examining individual mutants killed by one tool but not by another allows us to understand strengths and weaknesses of the tools, in a context that makes comprehending the cause of the detection (or lack of it) easy: the difference between the un-mutated code and mutated code will always be small and relatively simple.
 Moreover, simply looking at how much two tools agree on mutants can answer the question of a user of static analysis tools: given that I am using tool A, would adding tool B be likely to add enough new, interesting results to make it worth my time to examine its output?
@@
 Interested users, e.g.
-security analystst, can even inspect the differences to get an idea of the particular cases when a tool might be most effective.
+security analysts, can inspect the differences to get an idea of the particular cases when a tool might be most effective, but a more typical user can simply look at a Venn diagram of kills like that shown in Figure \ref{fig:examplevenn}.
+Assume that tools A, B, and C all produce very similar numbers of findings for un-mutated code, and have similar execution times.
+Tool A is likely the most important tool to make use of; it detected more mutants than any other tool, and more than twice as many mutants were killed by A alone than by B alone.
+However, also running tool B is probably a very good idea, assuming, e.g., it is not a very expensive commercial tool.
+B does not do as well as A, but it is the only tool that detects a large number of mutants, and most mutants it detects are unique to it.
+Finally, Tool C may not be worth running; recall that it produces a similar number of findings to A and B on the un-mutated code, so it is notably bad at detecting faults (at least ones that look like mutants).
+It might be a good idea to just look at the 27 mutants detected by C alone:  if they represent an important class of potential problems (perhaps C specialized in detecting potentially non-terminating loops), then C might be useful, but if the first few mutants inspected are false positives, then C is likely not useful.
 
-Comparing mutant results also leads to the idea of improving static analysis tools by examining mutants detected by another tool (thus known to be in-principle detectable) but not by the tool to be improved.
+Comparing mutant results also leads to the idea of \emph{improving} static analysis tools by examining mutants detected by another tool (thus known to be in-principle detectable) but not by the tool to be improved.
 Of course, any faults in code, not just mutants, could serve this purpose.
@@
 SmartCheck, on the other hand, detects removing the {\tt return} in {\tt get} or replacing it with a {\tt revert}, or removing the {\tt public} visibility modifier for {\tt get} \footnote{Slither's ``missing return'' detector is only available in the private version of slither, or through the {\tt crytic} service provided by Trail of Bits.}.
-If we restrict our analyis to issues with a severity greater than informational, SmartCheck detects no mutants of the contract, while Slither still reports that some mutants \emph{allow an arbitrary caller to cause the contract to self destruct}.
+If we restrict our analyis to findings with a severity greater than informational, SmartCheck detects no mutants of the contract, while Slither still reports that some mutants \emph{allow an arbitrary caller to cause the contract to self destruct}.
 This simple example shows how our approach works on a small scale.

### 2020-01-22 analyzejava.py

@@ -1,4 +1,5 @@
 import csv
+import scipy
 
 spotbugs_killed = 0
 PMD_killed = 0
@@ -14,21 +15,41 @@ spotbugs_file = open("all.spotbugs.killed.txt", 'w')
 PMD_file = open("all.PMD.killed.txt", 'w')
 Infer_file = open("all.Infer.killed.txt", 'w')
 
+files = {}
+
+def readZero(v):
+    try:
+        r = int(v)
+        if r < 0:
+            return 0
+        return r
+    except:
+        return 0
+
 with open("Java.csv") as javaResults:
     reader = csv.DictReader(javaResults)
     for row in reader:
+        fname = row["Project"]+":"+row["File"]
+        if fname not in files:
+            files[fname] = {"mutants":0.0,"spotbugs_kills":0,"spotbugs_issues":readZero(row["Spotbug"]),
+                                "PMD_kills":0,"PMD_issues":readZero(row["PMD"]),
+                                "Infer_kills":0,"Infer_issues":readZero(row["Infer"])}
         N += 1
+        files[fname]["mutants"] += 1
         if row["spotbugs_killed"] == "1":
+            files[fname]["spotbugs_kills"] += 1
             spotbugs_killed += 1
             spotbugs_file.write(row["Mutant"] + "\n")
             if row["PMD_killed"] == "0":
                 spotbugs_not_PMD += 1
         if row["PMD_killed"] == "1":
+            files[fname]["PMD_kills"] += 1
             PMD_killed += 1
             PMD_file.write(row["Mutant"] + "\n")
             if row["spotbugs_killed"] == "0":
                 PMD_not_spotbugs += 1
         if row["Infer_killed"] == "1":
+            files[fname]["Infer_kills"] += 1            
             Infer_file.write(row["Mutant"] + "\n")            
             print(row)
             Infer_killed += 1
@@ -51,3 +72,35 @@ with open("Java.csv") as javaResults:
     print()
     print("spotbugs but not PMD:", spotbugs_not_PMD)
     print("PMD but not spotbugs:", PMD_not_spotbugs)    
+
+    print()
+    print()
+    print(len(files), "FILES")
+    print()
+    for tool in ["spotbugs", "PMD", "Infer"]:
+        print()
+        print("="*48)
+        print("TOOL:", tool)
+        cleanCount = 0
+        cleanScores = []
+        allCleanScores = []
+        allCleanCount = 0
+        issues = []
+        scores = []
+        for f in files:
[…]

### 2020-01-22 introduction.tex
small fixes

@@
 \item We propose a definition of mutant killing that works well in a static analysis context.
-\item We introduce a simple scheme for prioritizing mutants that helps users understand the results of analysis and guide efforts to improve, rather than simple compare, tools.
+\item We introduce a simple scheme for prioritizing mutants that helps users understand the results of analysis and guide efforts to improve, rather than simply compare, tools.
 \item We apply our method to an extensive, in-depth comparison of three Solidity smart contract analysis tools, and show how prioritization allowed us to easily identify (and build) three new detectors for the most effective of these tools.
@@
 \end{itemize}
+
+The Solidity, Java, and Python case studies also serve to answer a set of research questions investigating the utility of our approach, and provide basic evidence that it provides actionable, non-obvious information that is otherwise difficult to produce.
+While there are certainly limitations to using differential mutation analysis to compare (and extend) static analysis tools, the method scales to basing comparisons on large numbers of real software source files, but has some of the advantages of humans establishing ground truth for tool findings.

### 2020-01-22 experiments.tex
mutant
big add code changes

@@
 \subsubsection{Original Program Selection}
+
+We could have used a set of high-transaction contracts, or known-important contracts to validate our approach.
+However, we knew that one of our goals in the Solidity experiments was to actually improve a mutation analysis tool, and the developers of the static analysis tools use exactly such benchmarks to validate their tools.
+Basing our improvements on mutants of the contracts used for evaluation of proposed detectors would introduce a serious bias in our favor: we would be more likely to produce detectors that would have true positives and few false positives on the benchmark contracts.
+We therefore instead selected 100 random contracts for which EtherScan (\url{https://etherscan.io/}) has source code, and used this (quite arbitrary) set of contracts from the actual blockchain to compare tools and identify opportunities for improvement.
 
@@
 
-\subsubsection{Improving Slither}
+\subsubsection{Improving Slither {\bf (RQ 4)}}
+
+Based on the differential mutation analysis, we identified three low-hanging fruit to improve the performance of Slither.
+The process was simple.
+First, we produced a list of all mutants killed by either SmartCheck or Securify, but not killed by Slither.
+We then applied the prioritization method based on the FPF algorithm and the distance metric described in Section \ref{sec:prioritizing}, and examined the mutants in rank order.
+Many of the mutants were difficult to identify as true or false positives, absent context.
+Some opportunities for enhancement were clear, but seemed likely to require considerable effort to implement without producing a large number of false positives.
+For example, Securify often detected when an ERC20 token contract's guard preventing making the special 0x0 address the owner of a contract was removed, and issued the error {\tt Violation for MissingInputValidation}.
+Detecting such missing guards is probably useful, but formulating a way to do it without producing false positives is non-trivial.
+We wanted to show that mutants could identify \emph{useful} but \emph{easy to implement} missing detectors.
+Examining the first few mutants, we identified three such, based on mutants killed by either SmartCheck or Securify, or both:
+
+\begin{enumerate}
+\item {\bf Boolean constant misuse:}  This detector flags code like {\tt if (true}} or {\tt g(b || true)} (where {\tt g} is a function that takes a Boolean input).
+Constant-valued conditionals tend to indicate debugging efforts that have persisted into production code, or other faults; there are almost no circumstances where a conditional should not vary with state or input.
+This detector is actually split into two detectors, one for this serious issue, and an informational/stylistic detector that notes that code such as {\tt if (x == true)}, while semantically harmless, is difficult to read.
+This problem was easily identified from cases such as this mutant, killed by SmartCheck but not Slither:
+
+\begin{code}
+0x598ab825d607ace3b00d8714c0a141c7ae2e6822_Vault.mutant.275.sol: 424contracts//0x598ab825d607ace3b00d8714c0a141c7ae2e6822_Vault.sol:249
+        if (!p.recipient.send(p.amount)) \{  // Make the payment
+ ==>          if (true) \{  // Make the payment
+        if (true) {  // Make the payment
+\end{code}
+
+\item {\bf Type-based tautologies:}  A type-based tautology is again a case where a Boolean expression has a constant value, but this is not due to misuse of a Boolean constant, but is instead due to the \emph{types} in a comparison.
+For example, if {\tt x} is an unsigned integer type, the comparison {\tt x >= 0} is always true and {\tt x < 0} is always false.
+This detector is a generalization of the SmartCheck detector \url{https://github.com/smartdec/smartcheck/blob/master/rule_descriptions/SOLIDITY_UINT_CANT_BE_NEGATIVE/description_en.html}, modified to actually compute the ranges of types and identify more general instances of tautological comparisons, e.g.
+{\tt y < 512} where {\tt y}'s type is {\tt int8}.
+Again, the problem was easily identified by a mutants killed by SmartCheck but not Slither such as:
+
+\begin{code}
+0x968815CD73647C3af02a740a2438D6f8219e7534_TTPresale.mutant.311.sol: 424contracts//0x968815CD73647C3af02a740a2438D6f8219e7534_TTPresale.sol:175
+        require(nextDiscountTTMTokenId6 >= 361 && nextDiscountTTMTokenId6 <= 391);
+ ==>  ...361...==>...0...
+        require(nextDiscountTTMTokenId6 >= 0 && nextDiscountTTMTokenId6 <= 391);
+\end{code}
+
+\item {\bf Loss of precision due to divide before multiply:}  Solidity only supports integer types.
+This means that performing division before multiplication can introduce rounding that is not present when the multiplication is performed first.
+This is a fairly important problem, given the frequency with which Solidity code performs financial calculations where maximum precision is desired.
+SmartCheck provides a detector for such precision losses \url{https://github.com/smartdec/smartcheck/blob/master/rule_descriptions/SOLIDITY_DIV_MUL/description_en.html}, which enabled it to detect mutants such as:
+
+\begin{code}
+0x534ccee849a688581d1b0c65e7ff317ed10c5ed3_NametagToken.mutant.480.sol: 424contracts//0x534ccee849a688581d1b0c65e7ff317ed10c5ed3_NametagToken.sol:650
+        byte char = byte(bytes32(uint(x) * 2 ** (8 * j)));
+ ==>  ...*...==>.../...
+        byte char = byte(bytes32(uint(x) * 2 ** (8 / j)));
+\end{code}
+
+All three of these detectors were submitted as PRs, vetted over an internal benchmark set of contracts used by the Slither developers to evaluate new detectors, and accepted for release in the public version of Slither.
+All three detectors produce some true positives (actual problems, though not always exploitable) in benchmark contracts, have acceptably low false positive rates, and were deemed valuable enough to include as non-informational (medium severity) detectors.
+The first mutants in prioritized rank exhibiting the issues, shown above, were the 2nd, 9th, and 12th non-statement-deletion mutants ranked for SmartCheck, out of over 800 such mutants.
+Using our prioritization, it was possible to identify these issues by examining fewer than 20 unkilled mutants.
+Without prioritization, on average a developer would have to look at more than 200, 80, and 400 mutants, respectively, to find instances of these problems.
+
+There were 92 separately ranked statement deletion mutants also.
+These, however, could all be ignored, as they were almost entirely duplicates related to the missing-return statement detector.
+If this detector were not already present as a private Slither detector, it would also be a good candidate for addition to the tool.
+Our three submitted detectors were not present as private detectors, and only one (the type-based tautology detector) had even been identified, via a GitHub issue, as a potential improvement (and only in the private version of Slither).
+
+Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants.
+The majority of cases involved either the missing return problem or replacing {\tt msg.sender} with {\tt tx.origin}; Slither has a detector for misuses of {\tt tx.origin}, and we believe (but are not sure) that almost all of these are due to intentional behavior: SmartCheck and Securify tend to identify most uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule intended to avoid false positives.
+
+It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ4}.
+
+\end{enumerate}
 
@@
   \begin{tabular}{l|r|r|r|r|r}
-    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & \\
+    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
     Tool & Mean & Median & Mean & Median & Ratio\\

### 2020-01-22 prioritizing.tex
label the section

\section{Prioritizing Mutants}
\label{sec:prioritizing}

### 2020-01-22 experiments.tex
p

@@
 \begin{enumerate}
-\item {\bf Boolean constant misuse:}  This detector flags code like {\tt if (true}} or {\tt g(b || true)} (where {\tt g} is a function that takes a Boolean input).
+\item {\bf Boolean constant misuse:}  This detector flags code like {\tt if (true)} or {\tt g(b || true)} (where {\tt g} is a function that takes a Boolean input).
 Constant-valued conditionals tend to indicate debugging efforts that have persisted into production code, or other faults; there are almost no circumstances where a conditional should not vary with state or input.
@@
 \begin{code}
-0x598ab825d607ace3b00d8714c0a141c7ae2e6822_Vault.mutant.275.sol: 424contracts//0x598ab825d607ace3b00d8714c0a141c7ae2e6822_Vault.sol:249
+0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.mutant.275.sol: 424contracts//0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.sol:249
         if (!p.recipient.send(p.amount)) \{  // Make the payment
  ==>          if (true) \{  // Make the payment
-        if (true) {  // Make the payment
+        if (true) \{  // Make the payment
 \end{code}
@@
 \begin{code}
-0x968815CD73647C3af02a740a2438D6f8219e7534_TTPresale.mutant.311.sol: 424contracts//0x968815CD73647C3af02a740a2438D6f8219e7534_TTPresale.sol:175
-        require(nextDiscountTTMTokenId6 >= 361 && nextDiscountTTMTokenId6 <= 391);
+0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.mutant.311.sol: 424contracts//0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.sol:175
+        require(nextDiscountTTMTokenId6 >= 361 \&\& nextDiscountTTMTokenId6 <= 391);
  ==>  ...361...==>...0...
-        require(nextDiscountTTMTokenId6 >= 0 && nextDiscountTTMTokenId6 <= 391);
+        require(nextDiscountTTMTokenId6 >= 0 \&\& nextDiscountTTMTokenId6 <= 391);
 \end{code}
@@
 \begin{code}
-0x534ccee849a688581d1b0c65e7ff317ed10c5ed3_NametagToken.mutant.480.sol: 424contracts//0x534ccee849a688581d1b0c65e7ff317ed10c5ed3_NametagToken.sol:650
+0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.mutant.480.sol: 424contracts//0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.sol:650
         byte char = byte(bytes32(uint(x) * 2 ** (8 * j)));
@@
 \end{code}
+
+\end{enumerate}
 
@@
 It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ4}.
-
-\end{enumerate}
 

### 2020-01-22 analyzejava.py

@@ -26,10 +26,13 @@ def readZero(v):
     except:
         return 0
 
+projects = set([])
+    
 with open("Java.csv") as javaResults:
     reader = csv.DictReader(javaResults)
     for row in reader:
         fname = row["Project"]+":"+row["File"]
+        projects.add(row["Project"])
         if fname not in files:
             files[fname] = {"mutants":0.0,"spotbugs_kills":0,"spotbugs_issues":readZero(row["Spotbug"]),
                                 "PMD_kills":0,"PMD_issues":readZero(row["PMD"]),
@@ -104,3 +107,5 @@ with open("Java.csv") as javaResults:
         print("CLEAN SCORES:", scipy.mean(cleanScores), scipy.median(cleanScores))
         print("#ALL CLEAN:", allCleanCount),
         print("ALL CLEAN SCORES:", scipy.mean(allCleanScores), scipy.median(allCleanScores))
+
+print(projects, len(projects))

### 2020-01-23 analyzepython.py
py

@@ -1,0 +1,114 @@
+import csv
+import scipy
+
+Pylint_killed = 0
+Pychecker_killed = 0
+Pylint_not_Pychecker = 0
+Pychecker_not_Pylint = 0
+Pyflakes_killed = 0
+Pylint_warnings = 0
+Pychecker_warnings = 0
+Pyflakes_warnings = 0
+N = 0.0
+
+Pylint_file = open("all.Pylint.killed.txt", 'w')
+Pychecker_file = open("all.Pychecker.killed.txt", 'w')
+Pyflakes_file = open("all.Pyflakes.killed.txt", 'w')
+
+files = {}
+
+def readZero(v):
+    try:
+        r = int(v)
+        if r < 0:
+            return 0
+        return r
+    except:
+        return 0
+
+projects = set([])
+
+Mutant = 0
+
+with open("Python.csv") as javaResults:
+    reader = csv.DictReader(javaResults)
+    for row in reader:
+        Mutant += 1
+        fname = row["projectname"]
+        projects.add(row["projectname"])
+        if fname not in files:
+            files[fname] = {"mutants":0.0,"Pylint_kills":0,"Pylint_issues":readZero(row["BeforePylintCount"]),
+                                "Pychecker_kills":0,"Pychecker_issues":readZero(row["BeforePycheckerCount"]),
+                                "Pyflakes_kills":0,"Pyflakes_issues":readZero(row["BeforePyflakesCount"])}
+        N += 1
+        files[fname]["mutants"] += 1
+        if row["PylintKilled"] == "TRUE":
+            files[fname]["Pylint_kills"] += 1
+            Pylint_killed += 1
+            Pylint_file.write(str(Mutant) + "\n")
+            if row["Pycheckerkilled"] == "FALSE":
+                Pylint_not_Pychecker += 1
+        if row["Pycheckerkilled"] == "TRUE":
+            files[fname]["Pychecker_kills"] += 1
+            Pychecker_killed += 1
+            Pychecker_file.write(str(Mutant) + "\n")
+            if row["PylintKilled"] == "FALSE":
+                Pychecker_not_Pylint += 1
+        if row["Pyflakeskilled"] == "TRUE":
+            files[fname]["Pyflakes_kills"] += 1            
+            Pyflakes_file.write(str(Mutant) + "\n")            
+            print(row)
+            Pyflakes_killed += 1
+        try:
+            Pylint_warnings += int(row["BeforePylintCount"])
+        except:
+            pass
+        try:
[…]

### 2020-01-23 experiments.tex
py
python

@@
 
+\subsubsection{Static Analysis Tools Compared}
+
 We analyzed three well-known tools for static analysis of Solidity smart contracts: Slither \cite{slither}, SmartCheck \cite{smartcheck}, and Securify \cite{Securify}.
 
-\subsubsection{Original Program Selection}
+\subsubsection{Smart Contract Selection}
 
@@
 
+\begin{figure}
+  {\scriptsize
+\begin{code}
+  0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.mutant.275.sol:
+  424contracts/0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.sol:249
+        if (!p.recipient.send(p.amount)) \{  // Make the payment
+ ==>          if (true) \{  // Make the payment
+        if (true) \{  // Make the payment
+      \end{code}
+      }
+      \caption{Mutant showing {\bf Boolean constant misuse}.}
+      \label{fig:boolean}
+    \end{figure}
+
+    \begin{figure}
+      {\scriptsize
+\begin{code}
+  0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.mutant.311.sol:
+  424contracts/0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.sol:175
+        require(nextDiscountTTMTokenId6 >= 361 \&\& nextDiscountTTMTokenId6 <= 391);
+ ==>  ...361...==>...0...
+        require(nextDiscountTTMTokenId6 >= 0 \&\& nextDiscountTTMTokenId6 <= 391);
+      \end{code}
+      }
+      \caption{Mutant showing {\bf Type-based tautologies}.}
+      \label{fig:tautology}      
+\end{figure}
+
+\begin{figure}
+  {\scriptsize
+\begin{code}
+  0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.mutant.480.sol:
+  424contracts/0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.sol:650
+        byte char = byte(bytes32(uint(x) * 2 ** (8 * j)));
+ ==>  ...*...==>.../...
+        byte char = byte(bytes32(uint(x) * 2 ** (8 / j)));
+      \end{code}
+      }
+      \caption{Mutant showing {\bf Loss of precision}.}
+      \label{fig:divmul}            
+      \end{figure}
+
 Based on the differential mutation analysis, we identified three low-hanging fruit to improve the performance of Slither.
@@
 This detector is actually split into two detectors, one for this serious issue, and an informational/stylistic detector that notes that code such as {\tt if (x == true)}, while semantically harmless, is difficult to read.
-This problem was easily identified from cases such as this mutant, killed by SmartCheck but not Slither:
-
-\begin{code}
-0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.mutant.275.sol: 424contracts//0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.sol:249
-        if (!p.recipient.send(p.amount)) \{  // Make the payment
- ==>          if (true) \{  // Make the payment
-        if (true) \{  // Make the payment
-\end{code}
+This problem was easily identified from cases such as the mutant in Figure \ref{fig:boolean}, killed by SmartCheck but not Slither:
 
@@
 For example, if {\tt x} is an unsigned integer type, the comparison {\tt x >= 0} is always true and {\tt x < 0} is always false.
-This detector is a generalization of the SmartCheck detector \url{https://github.com/smartdec/smartcheck/blob/master/rule_descriptions/SOLIDITY_UINT_CANT_BE_NEGATIVE/description_en.html}, modified to actually compute the ranges of types and identify more general instances of tautological comparisons, e.g.
+This detector is a generalization of the SmartCheck detector \url{https://github.com/smartdec/smartcheck/blob/master/rule\_descriptions/SOLIDITY\_UINT\_CANT\_BE\_NEGATIVE/description\_en.html}, modified to actually compute the ranges of types and identify more general instances of tautological comparisons, e.g.
 {\tt y < 512} where {\tt y}'s type is {\tt int8}.
-Again, the problem was easily identified by a mutants killed by SmartCheck but not Slither such as:
+Again, the problem was easily identified by mutants killed by SmartCheck but not Slither such as the one in Figure \ref{fig:tautology}.
 
-\begin{code}
-0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.mutant.311.sol: 424contracts//0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.sol:175
-        require(nextDiscountTTMTokenId6 >= 361 \&\& nextDiscountTTMTokenId6 <= 391);
- ==>  ...361...==>...0...
-        require(nextDiscountTTMTokenId6 >= 0 \&\& nextDiscountTTMTokenId6 <= 391);
-\end{code}
-
-\item {\bf Loss of precision due to divide before multiply:}  Solidity only supports integer types.
+\item {\bf Loss of precision:}  Solidity only supports integer types.
 This means that performing division before multiplication can introduce rounding that is not present when the multiplication is performed first.
 This is a fairly important problem, given the frequency with which Solidity code performs financial calculations where maximum precision is desired.
-SmartCheck provides a detector for such precision losses \url{https://github.com/smartdec/smartcheck/blob/master/rule_descriptions/SOLIDITY_DIV_MUL/description_en.html}, which enabled it to detect mutants such as:
-
-\begin{code}
-0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.mutant.480.sol: 424contracts//0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.sol:650
-        byte char = byte(bytes32(uint(x) * 2 ** (8 * j)));
- ==>  ...*...==>.../...
-        byte char = byte(bytes32(uint(x) * 2 ** (8 / j)));
-\end{code}
-
+SmartCheck provides a detector for such precision losses \url{https://github.com/smartdec/smartcheck/blob/master/rule\_descriptions/SOLIDITY\_DIV\_MUL/description\_en.html}, which enabled it to detect mutants such as the one shown in Figure\ref{fig:divmul}.
 \end{enumerate}
@@
 
-\subsubsection{Original Program Selection}
+\subsubsection{Static Analysis Tools Compared}
+
+\subsubsection{Project Selection}
 
@@
 
+\subsubsection{Static Analysis Tools Compared}
+
+\subsubsection{Project Selection}
+
+\subsubsection{Analysis Results}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{python.png}
+  \caption{Mutants killed by Python static analysis tools.}
+  \label{fig:pythonvenn}
+\end{figure}
+
+\begin{table}
+  \begin{tabular}{l|r|r|r|r|r}
+    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
+    Tool & Mean & Median & Mean & Median & Ratio\\
+    \hline
+    \hline
+    Pylint & 2.24 & 1.00 & 0.22 & 0.20 & 0.097 \\
+    \hline
+    pyflakes & 2.4 & 1.00 & 0.03 & 0.00 & 0.012 \\
+    \hline
+    PyChecker & 0.56& 0.00 & 0.10 & 0.00 &  0.183 \\
+    \hline
+  \end{tabular}
+  \caption{Python tools: number of findings and mutation scores over all projects.}
+  \label{tab:scorepython}
+\end{table}
+
+\begin{table}
+  \begin{tabular}{l|r|r|r}
+    & & \multicolumn{2}{|c|}{} \\
+    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} \\
+    Tool & Projects & Mean & Median \\
+    \hline
+    \hline
+    Pylint & 0 & N/A & N/A \\
+    \hline
+    pyflakes & 12 & 0.02 & 0.00 \\
+    \hline
+    PyChecker & 21& 0.01 & 0.00 \\
+    \hline
+  \end{tabular}
+  \caption{Python tools: clean project counts and mutation scores.}
+  \label{tab:cleanpython}
+\end{table}
+
 \subsection{Threats to Validity}

### 2020-01-23 experiments.tex
projects

@@
 
+70,825 mutants
+
 \subsubsection{Static Analysis Tools Compared}
@@
     \hline
-    SpotBugs & 0.26 & 0.00 & 0.30 & 0.07 & 1.132 \\
+    SpotBugs & 0.47 & 0.00 & 0.37 & 0.21 & 0.7847 \\
     \hline
-    PMD & 0.48 & 0.00 & 0.09 & 0.00 & 0.183 \\
+    PMD & 1.27 & 0.00 & 0.07 & 0.07 & 0.0562 \\
     \hline
-    Infer & 0.00 & 0.00 & 0.00 & 0.00 &  0.006 \\
+    Infer & 0.07 & 0.00 & 0.00 & 0.00 &  0.0002 \\
     \hline
   \end{tabular}
-  \caption{Java tools: number of findings and mutation scores over all files.}
+  \caption{Java tools: number of findings and mutation scores over all projects.}
   \label{tab:scorejava}
@@
   \begin{tabular}{l|r|r|r|r|r}
-    & & \multicolumn{2}{|c|}{} & \multicolumn{2}{|c}{Clean For All (1,185)} \\
+    & & \multicolumn{2}{|c|}{} & \multicolumn{2}{|c}{Clean For All (10)} \\
     & \# Clean & \multicolumn{2}{|c|}{Mutation Score} &  \multicolumn{2}{|c}{Mutation Score}\\
-    Tool & Files & Mean & Median & Mean & Median\\
+    Tool & Projects & Mean & Median & Mean & Median\\
     \hline
     \hline
-    SpotBugs & 1,426 & 0.32 & 0.05 & 0.37 & 0.06 \\
+    SpotBugs & 12 & 0.42 & 0.46 & 0.44 & 0.46 \\
     \hline
-    PMD & 1,359 & 0.08 & 0.00 & 0.08 & 0.00 \\
+    PMD & 11 & 0.07 & 0.07 & 0.07 & 0.07 \\
     \hline
-    Infer & 1,598 & 0.00 & 0.00 & 0.00 & 0.00 \\
+    Infer & 14 & 0.00 & 0.00 & 0.00 & 0.00 \\
     \hline
   \end{tabular}
-  \caption{Java tools: clean file counts and mutation scores.}
+  \caption{Java tools: clean project counts and mutation scores.}
   \label{tab:cleanjava}
@@
 \subsection{Python Tools}
+
+167,511 mutants
 

### 2020-01-23 analyzejava.py

@@ -27,11 +27,14 @@ def readZero(v):
         return 0
 
 projects = set([])
+
+Mutant = 0
     
 with open("Java.csv") as javaResults:
     reader = csv.DictReader(javaResults)
     for row in reader:
-        fname = row["Project"]+":"+row["File"]
+        Mutant += 1
+        fname = row["Project"]
         projects.add(row["Project"])
         if fname not in files:
             files[fname] = {"mutants":0.0,"spotbugs_kills":0,"spotbugs_issues":readZero(row["Spotbug"]),
@@ -109,3 +112,4 @@ with open("Java.csv") as javaResults:
         print("ALL CLEAN SCORES:", scipy.mean(allCleanScores), scipy.median(allCleanScores))
 
 print(projects, len(projects))
+print(Mutant-1)

### 2020-01-23 analyzepython.py

@@ -112,3 +112,4 @@ with open("Python.csv") as javaResults:
         print("ALL CLEAN SCORES:", scipy.mean(allCleanScores), scipy.median(allCleanScores))
 
 print(projects, len(projects))
+print(Mutant-1)

### 2020-01-23 conclusions.tex
much closer to submit

\section{Conclusions and Future Work}

In this paper, we showed that program mutants can be used as a proxy
for real faults, to compare (and motivate improvements to) static
analysis tools.  Mutants are attractive in that changing a mostly
correct program usually introduces a bug into it; this is the basis of
mutation testing, after all, and a large body of work supports the
claim that at least 60-70\% of mutants are fault-inducing.   This
means we can assume mutants are faulty, and escape the
ground-truth/false positive problem that makes comparing static
analysis tools so labor-intensive.  It cannot help with precision
problems, directly, but combined with finding counts for un-mutated
code, our approach can identify tools that detect mutants well,
adjusted for their general tendency to flag any code as faulty.  We
evaluated 9 popular static analysis tools, for Solidity smart contracts, Java,
and Python, and offer advice to users of these tools.  We were also
able to use our methods, plus a novel mutant prioritization scheme, to
identify three useful new detectors for the Slither smart contract
analyzer.

As future work, we would like to further validate our approach and
improve our admittedly \emph{ad hoc} mutant distance metric.  Allowing
user feedback \cite{EndUserMistake,OnlyOracle}, or applying metric
learning methods \cite{kulis2012metric} (particularly unsupervised
learning \cite{scholkopf1998nonlinear,tipping1999probabilistic}) are
the most obvious and interesting possibilites for a better metric.
Finally, mutant prioritization should be applicable to improving
software testing, as well \cite{groce2018verified}.

### 2020-01-23 experiments.tex
much closer to submit

@@
 \begin{itemize}
-\item {\bf RQ1:}  Does our approach produce actionable results?
+\item {\bf RQ1:}  Does mutation analysis of static analysis tools produce actionable results?
 That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar in terms of mutation detection capability?
@@
 Can we confirm results from other evaluation methods, or informal opinion?
-\item {\bf RQ4:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that different tools are good at/bad at?
-Can we extend this information to improve tools by addressing identified weaknesses?
+\item {\bf RQ4:}  Do mutation scores for clean programs differ from those for programs for which tools report findings before mutation?
+\item {\bf RQ5:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
+ different tools are good at/bad at?  Can we extend this information to improve tools by addressing identified weaknesses?
+\item {\bf RQ6:}  How easy is it identify low-hanging fruit for improving tools without prioritization to remove conceptually redundant mutants?
   \end{itemize}
@@
 This expectation that (some subset of the) mutants can serve as proxies for real, detectable faults is the core concept of our approach.
+{\bf RQ4} addresses a concern briefly mentioned in the introduction:  it is possible that warnings for the original code interfere with our definition of detection.
+The ideal case for our approach is when a tool report no findings for un-mutated code, and either reports a finding (kills the mutant) or not (does not kill it) when the mutant is introduced.
+Chekam et al.
+showed that the ``clean program assumption'' for testing, the idea that coverage will be similar for faulty and fixed versions of a program, is a threat to the validity of investigations of the relationship between coverage and fault detection \cite{CleanProgram}, and we want to establish that this is unlikely to be the case for our approach.
+{\bf RQ5} and {\bf RQ6} receive only preliminary answers, for the smart contract tools.
 
@@
 We analyzed three well-known tools for static analysis of Solidity smart contracts: Slither \cite{slither}, SmartCheck \cite{smartcheck}, and Securify \cite{Securify}.
+Slither, based on an SSA-based intermediate language (SlithIR \cite{slither}) is a tool produced by Trail of Bits.
+SmartCheck, developed by SmartDec, translates Solidity source directly to an XML-based representation, then uses \emph{XPath} patterns to define problems.
+Securify, from SRI Systems Lab at ETH Zurich, works at the bytecode level, first parsing and decompiling contracts, then translating to \emph{semantic facts} in order to look for predefined problems.
 
@@
 We therefore instead selected 100 random contracts for which EtherScan (\url{https://etherscan.io/}) has source code, and used this (quite arbitrary) set of contracts from the actual blockchain to compare tools and identify opportunities for improvement.
+The collected contracts had a total of 15,980 non-comment source lines, as measured by {\tt cloc}, with a mean size of 159.8 LOC and a median size of 108 LOC.
+The largest single contract had 1,127 lines of code.
+Smart contracts are typically smaller than other programs, for various reasons, which explains our use of 100 contracts, compared to a much smaller set of projects for Java and Python.
+The Universal Mutator generated 46,769 mutants for the 100 contracts.
 
@@
 Extracting the signal from Securify's noise will be difficult.
-We also note that while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick (it took about 8 days sequential compute time for Slither and 3 days for SmartCheck), Securify often required many hours to analyze a mutant, and frequently required a few days to analyze a mutant; the full analysis required over three months of compute time.
+We also note that while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick (it took about 6 days sequential compute time for Slither and 3 days for SmartCheck), Securify often required many hours to analyze a mutant, and frequently required a few days to analyze a mutant; the full analysis required over three months of compute time.
 
-We first compare the two tools that performed best.
-Slither detected 2.37 mean issues over the 100 contracts, and found 39 contracts free of all issues.
-Slither had a mean mutation score of 0.09 over all contracts, and 0.11 over clean (according to Slither) contracts.
-SmartCheck detected 1.89 mean issues over the 100 contracts, but only found 27 contracts free of all issues.
-The mean issue counts for the tools were not statistically significantly different ($p=0.34$).
-SmartCheck had a mean mutation score of 0.05 over all contracts, and 0.03 over clean (according to SmartCheck) contracts.
-The ratio between mean issues (Slither / SmartCheck) is 1.25, and the ratio of numbers of clean contracts is similar (1.44).
-The mutation score ratio, however, is 1.8, suggesting that Slither is not out-scoring SmartCheck \emph{only} because it produces more false positives.
-Slither detects 3,520 mutants not detected by SmartCheck, while SmartCheck only detects 1,629 mutants not detected by Slither (of these, 1,256 were also not detected by Securify) .
-Over the 7 contracts that both tools mark as clean, Slither has a mean mutation score of 0.08, and SmartCheck has a mean mutation score of 0.03; the difference, even with only 7 contracts, is statistically significant by paired Wilcoxon test, at $p=0.03$.
+For our research questions, {\bf RQ1} is clearly answered in the affirmative.
+Figure \ref{fig:solidityvenn} shows that the tools address quite different problems, with all tools reporting far more uniquely detected mutants than mutants in common with other tools.
+There are only 18 mutants detected by all tools, all of them involving replacement of {\tt msg.sender} (the caller of a smart contract, which may be another smart contract) with {\tt tx.origin} (the original initiator of a sequence of blockchain calls, a ``human'' account).
+Use of {\tt tx.origin} is usuallly a bad idea, can lead to incorrect behavior, and may not be relied on to be meaningful at all in future versions of Ethereum (\url{https://ethereum.stackexchange.com/questions/196/how-do-i-make-my-dapp-serenity-proof}).
 
-Securify was an outlier, compared to the other tools.
-First, while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick, Securify often required hours to analyze a contract; the full analysis required over a month.
-Second, Securify identified a mean of 24.65 issues on the un-mutated contracts, with a median of 17 issues found; this was a significant difference in issue counts from either other tool, with $p<1.0^{-15}$ in both cases.
-Despite this huge number of non-informational warnings for the original contracts, the mean mutation score was even worse than for SmartCheck, 0.03.
-For contracts where Securify found no issues in the original contract, it \emph{never} detected any mutants.
-Securify, however, did detect 1,449 mutants not detected by Slither, 1,076 of which were not detected by SmartCheck.
-On the other hand, given the ratio of issue counts compared to both tools is over 10.0, it is fairly likely many of these \emph{are} false positives, despite the probability mutants are actual faults.
+{\bf RQ2} is also answered in the affirmative.
+Counting findings for un-mutated code might suggest that Securify is the best tool, by a wide margin, but in the context of its near-zero mutant ratio, we must suspect that many of the warnings are false positives.
+Slither has the best mutant ratio, but the margin between it and SmartCheck confirms that both tools likely provide value.
 
-Our recommendation, based on mutation analysis results, is that smart contract analysis should at least use both Slither and SmartCheck, perhaps applying only Slither until it reports no interesting issues.
-Securify does offer a substantial non-overlapping set of issues reported, however, but produces a very large number of false positives.
-Available human resources and total number of issues produced (as well as high runtime) might justify not using Securify in some cases.
+For {\bf RQ3}, there are only a few tool comparisons in the literature; this is probably due to the fast-moving nature of the blockchain analysis world; the oldest of these tools' publication dates is 2018.
+The most extensive is that of Durieux et al.
+\cite{durieux2019empirical}.
+Slither detected 17\% of known vulnerabilities in their analyis, vs.
+11\% for SmartCheck and 9\% for Securify.
+Slither and SmartCheck were also among the four (out of 9) tools that detected vulnerabilities in the most categories; Securify was not.
+The overall recommendation of Durieux et al.
+was to use a combination of Slither and Mythril \cite{mythril-code} for contract analysis.
+Parizi et al.
+\cite{Parizi} also offer a ranking of tools, and determined that SmartCheck was the most effective, and far more so than Securify; unfortunately, they did not include Slither in their set of evaluated tools.
 
-\subsubsection{Improving Slither {\bf (RQ 4)}}
+The Slither paper \cite{slither} also provides an evaluation of all three tools.
+Their findings counts differ from ours because of different choices (we threw out merely informational results), but these are unrelated to mutation analysis, in any case.
+The evaluation only considered reentrancy faults \cite{SurveyAttacks,FC20} (which are sometimes, but only rarely, introduced by mutants).
+For reentrancy, Slither performed best on two real-world large contracts, finding subtle bugs in both, SmartCheck detected the problem in one of the two, and Securify detected neither.
+For a dataset of 1,000 contracts, SmartCheck had a high false positive rate (over 70\%) but detected more actual reentrancies (209) than Slither (99) or Securify (6).
+On the other hand, Slither's low false positive rate of 11\%  makes its results possibly more useful in practice.
+It seems safe to say that
+
+For {\bf RQ4}, on the changes seen when restricting analysis to clean contracts, Slither did slightly better at detecting mutants when the original contract was clean for Slither, and the other two tools did somewhat worse on contracts for which they reported no findings.
+For the three contracts clean for all tools, Slither performed almost exactly as it did over contracts in general, and the other tools performed worse, by about the same margin as they did for their own clean contracts.
+For our approach, we only need a week version of the ``clean program assumption'':  the threat is that kills may be under-reported for non-clean programs, due to interference with findings for the original code.
+It is not a problem if mutation scores are \emph{worse} for programs where a tool reports no findings for the un-mutated code.
+We therefore, for smart contracts, find no threat to our approach arising from the presence of findings on un-mutated code.
+We speculate that ``clean'' results for some tools result from contracts where the tool has trouble with the contract code, but does not actually crash; Slither may do better on clean code because it has fewer such failures, and clean contracts are probably generally simpler and easier to analyze.
+
+\subsubsection{Improving Slither {\bf (RQ5 and RQ6)}}
 
@@
 
-Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants.
-The majority of cases involved either the missing return problem or replacing {\tt msg.sender} with {\tt tx.origin}; Slither has a detector for misuses of {\tt tx.origin}, and we believe (but are not sure) that almost all of these are due to intentional behavior: SmartCheck and Securify tend to identify most uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule intended to avoid false positives.
-
-It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ4}.
+Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants, suggesting that for {\bf RQ6} it is indeed hard to use mutation analysis results without prioritization.
+A large majority of the mutants we inspected involved either the missing
+{\tt return} problem noted in the introduction, or replacing {\tt msg.sender} with {\tt tx.origin}; Slither, of course, has a detector for misuses of {\tt tx.origin}, and we believe (but are not sure) that almost all of the differences with respect to such changes are due to intentional behavior.
+SmartCheck and Securify tend to identify most (though not all) uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule, intended to reduce false positives.
+It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ5} and {\bf RQ6}.
 
@@
 \begin{table}
-  \begin{tabular}{l|r|r|r|r|r}
-    & & \multicolumn{2}{|c|}{} & \multicolumn{2}{|c}{Clean For All (10)} \\
-    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} &  \multicolumn{2}{|c}{Mutation Score}\\
-    Tool & Projects & Mean & Median & Mean & Median\\
+  \begin{tabular}{l|r|r|r}
+    & & \multicolumn{2}{|c|}{} \\
+    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} \\
+    Tool & Projects & Mean & Median \\
     \hline
     \hline
-    SpotBugs & 12 & 0.42 & 0.46 & 0.44 & 0.46 \\
+    SpotBugs & 3 & 0.41 & 0.17 \\
     \hline
-    PMD & 11 & 0.07 & 0.07 & 0.07 & 0.07 \\
+    PMD & 0 & N/A & N/A \\
     \hline
-    Infer & 14 & 0.00 & 0.00 & 0.00 & 0.00 \\
+    Infer & 6 & 0.00 & 0.00 \\
     \hline
   \end{tabular}
-  \caption{Java tools: clean project counts and mutation scores.}
+  \caption{Java tools: clean project counts and mutation scores.  No projects were clean for all tools, due to PMD producing findings for all projects.}
   \label{tab:cleanjava}
@@
   \end{tabular}
-  \caption{Python tools: clean project counts and mutation scores.}
+  \caption{Python tools: clean project counts and mutation scores.  No projects were clean for all tools, due to Pylint producing findings for all projects.}
   \label{tab:cleanpython}
@@
 \subsection{Threats to Validity}
+
+The primary threat to validity in terms of generalization is that we only examined nine static analysis tools, and our analyses were restricted to 100 smart contracts, 15 Java projects, and 25 Python projects.
+Because it is hard to identify a ground truth to compare with (the motivation for our approach), we cannot be certain that our rankings of tools are correct even for these tools and this code.
+However, where there are existing discussions of the tools, our results seem to agree with these, but add substantial detail.
+
+A general threat to our approach is that mutants may differ substantially from ``real'' faults, in some way.
+This is certainly true, in a sense \cite{GopinathMutants}, but for static analysis purposes, the risk is that some mutation operators align with patterns a particular tool identifies, biasing the evaluation in favor of that tool.
+Such faults may be dis-proportionately present in mutants vs.
+real code.
+However, we consider this unlikely.
+The vast majority of applied mutation operations for all of our experiments were highly generic, and do not plausibly represent a pattern in which some tool might specialize.
+Statement deletions leave no ``trace'' for a tool to match against, but only an omission.
+Changing arithmetic and comparison operators and numeric constants (incrementing, decrementing, or changing to 0 or 1) account for most of the non-statement-deletion mutants.
+We cannot see how any tool could have a (useful) detection rule that looks for these kinds of changes (a tool obviously would not flag all instances of 0 and 1 constants, or addition).
+Finally, many mutants (though many fewer than in these categories) add {\tt break} or {\tt continue} statements.
+Again, we do not see how a tool could take advantage of this without either genuinely understanding likely-bad control flow patterns (in which case, we think it ``deserves'' to kill more mutants) or flagging so many instances of these it would have a bad mutant ratio.

### 2020-01-23 prioritizing.tex
much closer to submit

@@
 \label{sec:prioritizing}
+
+One goal of our approach is to make it easy for tool developers to
+examine cases where one tool kills a mutant and another fails to, in
+order to identify patterns for new detectors.  Dedicated developers may also
+simple want to scan all mutants their tool does not kill, for the same
+purpose, analogous to what Groce et al. have proposed for automated
+verification and testing \cite{groce2015verified,groce2018verified}.
+Security analysts and other expert users who are not developers may
+also wish to do this, to better understand details of tool strengths
+and weaknesses.
+
+Unfortunately the full list of unkilled mutants, or differentially
+unkilled mutants, is likely to be both large and highly redundant.  In
+our results below, only one of 9 tools we examined killed fewer than
+1,000 mutants it alone detected.  Any cross-tool comparison is thus
+likely to involve hundreds or thousands of mutants.
+
+The problem of identifying unique ``faults'' (tool weaknesses) in this
+situation is very similar to the \emph{fuzzer taming} problem in
+software testing, as defined by Chen et al. \cite{PLDI13}: ``Given a
+potentially large collection of test cases, each of which triggers a
+bug, rank them in such a way that test cases triggering distinct bugs
+are early in the list.'' \cite{PLDI13}.  Their solution was
+to use Gonzalez' Furthest-Point-First \cite{Gonzalez} (FPF) algorithm
+to \emph{rank} test cases so that users can examine very different
+test cases as quickly as possible.  An
+FPF ranking requires a distance metric $d$, and ranks items so that
+dissimilar ones appear earlier.
+The hypothesis of Chen et al. was that dissimilar tests, by a
+well-chosen metric, will also fail due to different faults.  FPF is a
+greedy algorithm that proceeds by repeatedly adding the item with the
+\emph{maximum minimum distance to all previously ranked items}. Given an
+initial seed item $r_0$, a set $S$ of items to rank, and a distance
+metric $d$, FPF computes $r_i$ as
+$s \in S: \forall s' \in S: min_{ j < i}(d(s,r_j)) \geq min_{j <
+  i}(d(s',r_j))$.  The condition on $s$ is obviously true when
+$s = s'$, or when $s' = r_j$ for some $j < i$; the other cases for
+$s'$ force selection of \emph{some}
+max-min-distance $s$.
+
+In order to apply FPF ranking to examining mutants, we implemented a
+simple, somewhat \emph{ad hoc} distance metric and FPF ranker in the
+Universal Mutator \cite{universalmutator} tool.  Our metric $d$ is
+the sum of a set of measurements.  First, it adds a similarity
+ratio based on Levenshtein distance \cite{lev} for (1) the \emph{changes} (Levenshtein edits) from
+the original source code elements to
+the two mutants,  (2) the two original source code elements changed (in
+general, lines), and (3) the actual output mutant code.  These are
+weighted with multipliers of 5.0, 0.1, and 0.1, respectively; the type
+of change (mutation operator, roughly) dominates this part of the
+distance, because it best describes ``what the mutant did''; however,
+because many mutants will have the same change (e.g., changing {\tt +}
+to {\tt -}, the other ratios also often matter.  The Python
+{\tt Levenshtein} library's similarity ratio is used, as it is based
+on true minimal string edits; it reports similarity ratios between 0.0
+and 1.0.
+
+Our metric also incorporates a measure of the distance in the source
+code between the locations of two mutants.  If the mutants are to
+different files, this adds 0.5 to the distance; it also adds 0.25
+times the number of source lines separating the two mutants if they
+are in the same file, divided by 10, but caps the amount added at
+0.25.  The full metric, therefore is:
+
+$$ 5.0 \times r(\mathit{edit}_1, \mathit{edit}_2) +$$
+$$0.1 \times r(\mathit{source}_1, \mathit{source}_2) +$$
+$$0.1 \times r(\mathit{mutant}_1, \mathit{mutant}_2) +$$
+$$0.5 \times \mathit{different\_files} +$$
+$$max(0.25, \frac{\mathit{line\_distance}(\mathit{mutant}_1, \mathit{mutant}_2)}{10})$$
+
+\noindent Where $r$ is a Levenshtein-based string similarity ratio,
+and $\mathit{line\_distance}$ is the distance in a source file between
+two locations, in lines (and zero if the locations are in different
+source files).
+
+We do not claim this is an optimal, or even tuned, metric.  It
+provided useful results for a different problem, ranking mutants in an
+effort similar to that of Groce et
+al. \cite{groce2015verified,groce2018verified} to examine unkilled
+mutants for a smart contract library tested using the Echidna fuzzer,
+and we simply adopted it for our differential mutation analysis.
+Devising a better metric is left as future work, we only wish to show
+that even a hastily-devised and somewhat arbitrary metric provides
+considerable advantage over wading through an un-ordered list of
+mutants, and introduce the idea of using FPF for mutants, not just tests.

### 2020-01-23 relatedwork.tex
more
py
much closer to submit

@@
 This paper proposes a non-na\"ive differential comparison (not, exactly, differential testing, however, in that only aggregate results are possible to interpret without human intelligence), based on the observation that the ability to detect program mutants offers an automatable way to tell which of two tools is better (for a given universe of examples, at least) at telling faulty from non-faulty code.
+
 Klinger et al.
@@
 in contrast to their introduction of checks that may be guaranteed to pass, so this is probably not very helpful).
+Most importantly, however, their approach only applies to tools that check assertions, rather than more general static analysis tools that may only identify bad code patterns, not compute state reachability (e.g, most of the widely used tools).
+
+Finally, the large body of work on using mutants in software testing \cite{demillo1978hints,budd1980theoretical, groce2015verified,groce2018verified,MutGoogle,ivankovic2018industrial,mutKernel} is obviously relevant, in that we basically adopt its approach, but re-define killing a mutation for a static analysis context.
+Methods for improving mutation analysis in testing will often (but not always, e.g.
+those based on coverage or other dynamic information) transfer to our context as well.

### 2020-01-23 analyzejava.py

@@ -28,6 +28,8 @@ def readZero(v):
 
 projects = set([])
 
+projectFiles = {}
+
 Mutant = 0
     
 with open("Java.csv") as javaResults:
@@ -37,11 +39,30 @@ with open("Java.csv") as javaResults:
         fname = row["Project"]
         projects.add(row["Project"])
         if fname not in files:
-            files[fname] = {"mutants":0.0,"spotbugs_kills":0,"spotbugs_issues":readZero(row["Spotbug"]),
-                                "PMD_kills":0,"PMD_issues":readZero(row["PMD"]),
-                                "Infer_kills":0,"Infer_issues":readZero(row["Infer"])}
+            projectFiles[fname] = []
+            files[fname] = {"mutants":0.0,
+                                "spotbugs_kills":0,"spotbugs_issues":0,
+                                "PMD_kills":0,"PMD_issues":0,
+                                "Infer_kills":0,"Infer_issues":0}
+        if row["File"] not in projectFiles[fname]:
+            projectFiles[fname].append(row["File"])
+            files[fname]["spotbugs_issues"] += readZero(row["Spotbug"])
+            files[fname]["PMD_issues"] += readZero(row["PMD"])
+            files[fname]["Infer_issues"] += readZero(row["Infer"])
         N += 1
         files[fname]["mutants"] += 1
+        if row["spotbugs_killed"] not in ["0", "1"]:
+            print("NO RESULT FOR SPOTBUGS")
+        if row["PMD_killed"] not in ["0", "1"]:
+            print("NO RESULT FOR PMD")
+        try:
+            x = int(row["Spotbug"])
+        except:
+            print("NO ISSUES FOR SPOTBUG", row["spotbugs_killed"])
+        try:
+            x = int(row["PMD"])
+        except:
+            print("NO ISSUES FOR PMD")            
         if row["spotbugs_killed"] == "1":
             files[fname]["spotbugs_kills"] += 1
             spotbugs_killed += 1

### 2020-01-23 experiments.tex
Java nearly done, pending UCI

@@
   {\scriptsize
-\begin{code}
+\noindent \begin{code}
  0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.mutant.275.sol:
-  424contracts/0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.sol:249
  if (!p.recipient.send(p.amount)) \{  // Make the payment
@@
  0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.mutant.311.sol:
-  424contracts/0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.sol:175
  require(nextDiscountTTMTokenId6 >= 361 \&\& nextDiscountTTMTokenId6 <= 391);
@@
  0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.mutant.480.sol:
-  424contracts/0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.sol:650
  byte char = byte(bytes32(uint(x) * 2 ** (8 * j)));
@@
 
+Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools, and Tables \ref{tab:scorejava} and \ref{tab:cleanjava} provide numeric results for projects and clean projects, respectively.
+
+In terms of {\bf RQ1}, the raw results are clear:  SpotBugs finds vastly more uniquely killed mutants than the other tools, but PMD contributes a large number of its own unique detections, and clearly offers some independent utility.
+Infer on the other hand, is only able to detect two mutants, but these are unique.
+
+{\bf RQ2} is also answered in the affirmative.
+While the raw kills for SpotBugs are impressive, it also produced fewer findings for clean code, giving it an excellent mutant ratio, making its dominance even more clear.
+PMD's performance was acceptable, and better than any of the Solidity tools (note that comparing ratios across languages is potentially invalid, but not obviously so).
+Infer is confirmed as nearly useless, again.
+
+\begin{figure}
+  \url{https://stackoverflow.com/questions/4297014/what-are-the-differences-between-pmd-and-findbugs}
+  \url{https://www.sw-engineering-candies.com/blog-1/comparison-of-findbugs-pmd-and-checkstyle}
+  \url{https://www.reddit.com/r/java/comments/3i7w6n/checkstyle_vs_pmd_vs_findbugs_for_dummies_why/}
+\caption{Blog posts and forum threads discussing Java static analysis tools.}
+\label{fig:blog}
+\end{figure}
+
+For {\bf RQ3}, to our knowledge there is no academic comparison of these tools.
+However, the user postings listed in Figure \ref{fig:blog}, plus personal communications with security analysts who use these tools (citation removed for now, due to potential to compromise blinding) supported some basic conclusions.
+SpotBugs/FindBugs is indeed the best tool for finding bugs; PMD focuses more on stylistic issues and has a weaker semantic model.
+Mutants are much more likely to be bugs than stylistic issues (they alter semantics, not code style, in almost all cases: deletion or constant/operator changes obviously have no style impact).
+Infer is closer to a model checker focusing on resource leaks than a truly general-purpose tool, arguably.
+
+For {\bf RQ4} there were very few clean projects, but the mutation scores for Infer were unchanged, and for SpotBugs the mean mutation score improved and median mutation score was worse.
+Again, we see no evidence that non-clean code is a source of significant degradation on mutation detection.
+
 \begin{figure}
@@
 
-Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools.
-
 \subsection{Python Tools}
@@
 \subsubsection{Analysis Results}
+
+Figure \ref{fig:pythonvenn} shows the mutants killed by the Java analysis tools, and Tables \ref{tab:scorepython} and \ref{tab:cleanpython} provide numeric results for projects and clean projects, respectively.
+
+For {\bf RQ1}, Pylint and pyflakes both killed more than 15,000 mutants killed by no other tool; Pylint killed more mutants, but the relative difference was much smalller than in our Solidity or Java analysis.
+It is clear both tools are well worth using, assuming these are not false positives.
+Pyflakes' mutation score is lower than unique kills might suggest because it has almost no overlap with the other two tools.
+It is doing something quite different than they are, apparently.
+
+The mutant ratios {\bf RQ2} suggest that PyChecker might be better than pyflakes, in fact, and better than Pylint, because while it killed fewer mutants, it also produced fewer findings by far, overall.
+PyChecker would be a good choice for quick and dirty analysis of code on the fly, perhaps, in that it had excellent mutation scores for some programs (hence a high mean) but is relatively ``quiet'' compared to the other tools.
+However, its very low median mutation score is troubling; it may simply not work for some code.
+A combination of pyflakes and Pylint seems safer, though adding the relatively low-findings PyChecker seems likely to have little cost.
+For Python, we see that all of our measures are required to get a good picture of the tools.
+
+For {\bf RQ4} there were no clean projects for the known-to-be-aggressive Pylint.
+Performance for the other two tools was worse on projects for which they were clean, again failing to support the possibility that our definition of killing a mutant under-reports detection for non-clean code.
 

### 2020-01-23 experiments.tex
more python

@@
 
+\begin{figure}
+  \url{https://stackoverflow.com/questions/1428872/pylint-pychecker-or-pyflakes}
+  \url{https://www.reddit.com/r/Python/comments/ii3gm/experience_with_pylint_pychecker_pyflakes/}
+  \url{https://www.slant.co/versus/12630/12631/~pylint_vs_pyflakes}
+  \url{https://news.ycombinator.com/item?id=12748885}
+  \url{https://doughellmann.com/blog/2008/03/01/static-code-analizers-for-python/}
+\caption{Blog posts and forum threads discussing Python static analysis tools.}
+\label{fig:blogpython}
+\end{figure}
+
+For {\bf RQ3}, there were again no academic comparisons we could find.
+However, opinions on the web were quite common (see Figure \ref{fig:blogpython} lists ones we examined.
+It is hard to summarize the overall opinion here, since it ranges.
+There is probably general agreement that PyChecker is old and maybe less useful, but also terse and sometimes helpful.
+Pylint is the most recommended tool, and the general complaint that it is too picky was mitigated in our results by turning off clearly informative-only findings; users not configuring Pylint will probably see even more killed mutants, but likely a worse mutant ratio because mutants tend to introduce non-stylistic problems.
+Pyflakes is well liked, and supposedly less verbose than Pylint; we did not see this, but that may be because pyflakes aims to focus on ``actual bugs'' so it is equally verbose for mutants, which are likely to be actual bugs.
+
 For {\bf RQ4} there were no clean projects for the known-to-be-aggressive Pylint.

### 2020-01-23 experiments.tex
fixup

@@
 For {\bf RQ3}, there were again no academic comparisons we could find.
-However, opinions on the web were quite common (see Figure \ref{fig:blogpython} lists ones we examined.
-It is hard to summarize the overall opinion here, since it ranges.
+However, opinions on the web were quite common (see Figure \ref{fig:blogpython}, which lists ones we examined).
+It is hard to summarize the overall opinion here, since it ranges considerably.
 There is probably general agreement that PyChecker is old and maybe less useful, but also terse and sometimes helpful.
 Pylint is the most recommended tool, and the general complaint that it is too picky was mitigated in our results by turning off clearly informative-only findings; users not configuring Pylint will probably see even more killed mutants, but likely a worse mutant ratio because mutants tend to introduce non-stylistic problems.
-Pyflakes is well liked, and supposedly less verbose than Pylint; we did not see this, but that may be because pyflakes aims to focus on ``actual bugs'' so it is equally verbose for mutants, which are likely to be actual bugs.
+Pyflakes is also well liked, and supposedly less verbose than Pylint; we did not see this in our results, but that may be because we configured Pylint, and pyflakes aims to focus on ``actual bugs'' so it is equally verbose for mutants, which are likely to be actual bugs.
 

### 2020-01-23 experiments.tex
nearly done

@@
 
-Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools, and Tables \ref{tab:scorejava} and \ref{tab:cleanjava} provide numeric results for projects and clean projects, respectively.
-
-In terms of {\bf RQ1}, the raw results are clear:  SpotBugs finds vastly more uniquely killed mutants than the other tools, but PMD contributes a large number of its own unique detections, and clearly offers some independent utility.
-Infer on the other hand, is only able to detect two mutants, but these are unique.
-
-{\bf RQ2} is also answered in the affirmative.
-While the raw kills for SpotBugs are impressive, it also produced fewer findings for clean code, giving it an excellent mutant ratio, making its dominance even more clear.
-PMD's performance was acceptable, and better than any of the Solidity tools (note that comparing ratios across languages is potentially invalid, but not obviously so).
-Infer is confirmed as nearly useless, again.
-
-\begin{figure}
-  \url{https://stackoverflow.com/questions/4297014/what-are-the-differences-between-pmd-and-findbugs}
-  \url{https://www.sw-engineering-candies.com/blog-1/comparison-of-findbugs-pmd-and-checkstyle}
-  \url{https://www.reddit.com/r/java/comments/3i7w6n/checkstyle_vs_pmd_vs_findbugs_for_dummies_why/}
-\caption{Blog posts and forum threads discussing Java static analysis tools.}
-\label{fig:blog}
-\end{figure}
-
-For {\bf RQ3}, to our knowledge there is no academic comparison of these tools.
-However, the user postings listed in Figure \ref{fig:blog}, plus personal communications with security analysts who use these tools (citation removed for now, due to potential to compromise blinding) supported some basic conclusions.
-SpotBugs/FindBugs is indeed the best tool for finding bugs; PMD focuses more on stylistic issues and has a weaker semantic model.
-Mutants are much more likely to be bugs than stylistic issues (they alter semantics, not code style, in almost all cases: deletion or constant/operator changes obviously have no style impact).
-Infer is closer to a model checker focusing on resource leaks than a truly general-purpose tool, arguably.
-
-For {\bf RQ4} there were very few clean projects, but the mutation scores for Infer were unchanged, and for SpotBugs the mean mutation score improved and median mutation score was worse.
-Again, we see no evidence that non-clean code is a source of significant degradation on mutation detection.
-
 \begin{figure}
@@
 
+Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools, and Tables \ref{tab:scorejava} and \ref{tab:cleanjava} provide numeric results for projects and clean projects, respectively.
+
+In terms of {\bf RQ1}, the raw results are clear:  SpotBugs finds vastly more uniquely killed mutants than the other tools, but PMD contributes a large number of its own unique detections, and clearly offers some independent utility.
+Infer on the other hand, is only able to detect two mutants, but these are unique.
+
+{\bf RQ2} is also answered in the affirmative.
+While the raw kills for SpotBugs are impressive, it also produced fewer findings for clean code, giving it an excellent mutant ratio, making its dominance even more clear.
+PMD's performance was acceptable, and better than any of the Solidity tools (note that comparing ratios across languages is potentially invalid, but not obviously so).
+Infer is confirmed as nearly useless, again.
+
+\begin{figure}
+  {\scriptsize
+  \begin{itemize}
+  \item \url{https://stackoverflow.com/questions/4297014/what-are-the-differences-between-pmd-and-findbugs}
+  \item \url{https://www.sw-engineering-candies.com/blog-1/comparison-of-findbugs-pmd-and-checkstyle}
+  \item \url{https://www.reddit.com/r/java/comments/3i7w6n/checkstyle_vs_pmd_vs_findbugs_for_dummies_why/}
+  \end{itemize}
+  }
+\caption{Blog posts and forum threads discussing Java static analysis tools.}
+\label{fig:blog}
+\end{figure}
+
+For {\bf RQ3}, to our knowledge there is no academic comparison of these tools.
+However, the user postings listed in Figure \ref{fig:blog}, plus personal communications with security analysts who use these tools (citation removed for now, due to potential to compromise blinding) supported some basic conclusions.
+SpotBugs/FindBugs is indeed the best tool for finding bugs; PMD focuses more on stylistic issues and has a weaker semantic model.
+Mutants are much more likely to be bugs than stylistic issues (they alter semantics, not code style, in almost all cases: deletion or constant/operator changes obviously have no style impact).
+Infer is closer to a model checker focusing on resource leaks than a truly general-purpose tool, arguably.
+
+For {\bf RQ4} there were very few clean projects, but the mutation scores for Infer were unchanged, and for SpotBugs the mean mutation score improved and median mutation score was worse.
+Again, we see no evidence that non-clean code is a source of significant degradation on mutation detection.
+
 \subsection{Python Tools}
@@
 \subsubsection{Analysis Results}
-
-Figure \ref{fig:pythonvenn} shows the mutants killed by the Java analysis tools, and Tables \ref{tab:scorepython} and \ref{tab:cleanpython} provide numeric results for projects and clean projects, respectively.
-
-For {\bf RQ1}, Pylint and pyflakes both killed more than 15,000 mutants killed by no other tool; Pylint killed more mutants, but the relative difference was much smalller than in our Solidity or Java analysis.
-It is clear both tools are well worth using, assuming these are not false positives.
-Pyflakes' mutation score is lower than unique kills might suggest because it has almost no overlap with the other two tools.
-It is doing something quite different than they are, apparently.
-
-The mutant ratios {\bf RQ2} suggest that PyChecker might be better than pyflakes, in fact, and better than Pylint, because while it killed fewer mutants, it also produced fewer findings by far, overall.
-PyChecker would be a good choice for quick and dirty analysis of code on the fly, perhaps, in that it had excellent mutation scores for some programs (hence a high mean) but is relatively ``quiet'' compared to the other tools.
-However, its very low median mutation score is troubling; it may simply not work for some code.
-A combination of pyflakes and Pylint seems safer, though adding the relatively low-findings PyChecker seems likely to have little cost.
-For Python, we see that all of our measures are required to get a good picture of the tools.
-
-\begin{figure}
-  \url{https://stackoverflow.com/questions/1428872/pylint-pychecker-or-pyflakes}
-  \url{https://www.reddit.com/r/Python/comments/ii3gm/experience_with_pylint_pychecker_pyflakes/}
-  \url{https://www.slant.co/versus/12630/12631/~pylint_vs_pyflakes}
-  \url{https://news.ycombinator.com/item?id=12748885}
-  \url{https://doughellmann.com/blog/2008/03/01/static-code-analizers-for-python/}
-\caption{Blog posts and forum threads discussing Python static analysis tools.}
-\label{fig:blogpython}
-\end{figure}
-
-For {\bf RQ3}, there were again no academic comparisons we could find.
-However, opinions on the web were quite common (see Figure \ref{fig:blogpython}, which lists ones we examined).
-It is hard to summarize the overall opinion here, since it ranges considerably.
-There is probably general agreement that PyChecker is old and maybe less useful, but also terse and sometimes helpful.
-Pylint is the most recommended tool, and the general complaint that it is too picky was mitigated in our results by turning off clearly informative-only findings; users not configuring Pylint will probably see even more killed mutants, but likely a worse mutant ratio because mutants tend to introduce non-stylistic problems.
-Pyflakes is also well liked, and supposedly less verbose than Pylint; we did not see this in our results, but that may be because we configured Pylint, and pyflakes aims to focus on ``actual bugs'' so it is equally verbose for mutants, which are likely to be actual bugs.
-
-For {\bf RQ4} there were no clean projects for the known-to-be-aggressive Pylint.
-Performance for the other two tools was worse on projects for which they were clean, again failing to support the possibility that our definition of killing a mutant under-reports detection for non-clean code.
 
@@
 
+Figure \ref{fig:pythonvenn} shows the mutants killed by the Java analysis tools, and Tables \ref{tab:scorepython} and \ref{tab:cleanpython} provide numeric results for projects and clean projects, respectively.
+
+For {\bf RQ1}, Pylint and pyflakes both killed more than 15,000 mutants killed by no other tool; Pylint killed more mutants, but the relative difference was much smalller than in our Solidity or Java analysis.
+It is clear both tools are well worth using, assuming these are not false positives.
+Pyflakes' mutation score is lower than unique kills might suggest because it has almost no overlap with the other two tools.
+It is doing something quite different than they are, apparently.
+
+The mutant ratios {\bf RQ2} suggest that PyChecker might be better than pyflakes, in fact, and better than Pylint, because while it killed fewer mutants, it also produced fewer findings by far, overall.
+PyChecker would be a good choice for quick and dirty analysis of code on the fly, perhaps, in that it had excellent mutation scores for some programs (hence a high mean) but is relatively ``quiet'' compared to the other tools.
+However, its very low median mutation score is troubling; it may simply not work for some code.
+A combination of pyflakes and Pylint seems safer, though adding the relatively low-findings PyChecker seems likely to have little cost.
+For Python, we see that all of our measures are required to get a good picture of the tools.
+
+\begin{figure}
+  {\scriptsize
+  \begin{itemize}
+  \item \url{https://stackoverflow.com/questions/1428872/pylint-pychecker-or-pyflakes}
+  \item \url{https://www.reddit.com/r/Python/comments/ii3gm/experience_with_pylint_pychecker_pyflakes/}
+  \item \url{https://www.slant.co/versus/12630/12631/~pylint_vs_pyflakes}
+  \item \url{https://news.ycombinator.com/item?id=12748885}
+  \item \url{https://doughellmann.com/blog/2008/03/01/static-code-analizers-for-python/}
+  \end{itemize}
+  }
+\caption{Blog posts and forum threads discussing Python static analysis tools.}
+\label{fig:blogpython}
+\end{figure}
+
+For {\bf RQ3}, there were again no academic comparisons we could find.
+However, opinions on the web were quite common (see Figure \ref{fig:blogpython}, which lists ones we examined).
+It is hard to summarize the overall opinion here, since it ranges considerably.
+There is probably general agreement that PyChecker is old and maybe less useful, but also terse and sometimes helpful.
+Pylint is the most recommended tool, and the general complaint that it is too picky was mitigated in our results by turning off clearly informative-only findings; users not configuring Pylint will probably see even more killed mutants, but likely a worse mutant ratio because mutants tend to introduce non-stylistic problems.
+Pyflakes is also well liked, and supposedly less verbose than Pylint; we did not see this in our results, but that may be because we configured Pylint, and pyflakes aims to focus on ``actual bugs'' so it is equally verbose for mutants, which are likely to be actual bugs.
+
+For {\bf RQ4} there were no clean projects for the known-to-be-aggressive Pylint.
+Performance for the other two tools was worse on projects for which they were clean, again failing to support the possibility that our definition of killing a mutant under-reports detection for non-clean code.
+
 \subsection{Threats to Validity}

### 2020-01-23 experiments.tex
rep

@@
 {\bf RQ5} and {\bf RQ6} receive only preliminary answers, for the smart contract tools.
+
+We used the Universal Mutator \cite{universalmutator,regexpMut}, which aggressively produces large numbers of mutants, to generate all mutants.
+We will make available a repository containing our full raw results plus details on contracts and projects selected, for replication and further analysis, upon acceptance (to avoid breaking blinding).
 

### 2020-01-24 abstract.tex
initial submission

@@
-Many programming languages offer multiple static analysis tools that offer to detect faults in code witthout executing it.
-Understanding the strengths and weaknesses of tools, and performing direct comparisons of their effectiveness is diffcult; it usually involves either manual examination of differing warnings on real code, or the bias-prone construction of artificial test cases.
+Many programming languages offer multiple static analysis tools that offer to detect faults in code without executing it.
+Understanding the strengths and weaknesses of tools, and performing direct comparisons of their effectiveness is difficult; it usually involves either manual examination of differing warnings on real code, or the bias-prone construction of artificial test cases.
 In practice, comparisons tend to be limited to superficial, anecdotal discussions in the informal literature (e.g., blog posts by software developers), or purely research-community-oriented evaluations made by the authors of new tools seeking to publish their results.

### 2020-01-24 conclusions.tex
more python
initial submission

@@
 ground-truth/false positive problem that makes comparing static
-analysis tools so labor-intensive.  It cannot help with precision
+analysis tools so labor-intensive.  Our approach cannot identify precision
 problems, directly, but combined with finding counts for un-mutated
 code, our approach can identify tools that detect mutants well,
-adjusted for their general tendency to flag any code as faulty.  We
+adjusted for their general tendency to flag any code as faulty, and
+thus identify imprecise tools.  We
 evaluated 9 popular static analysis tools, for Solidity smart contracts, Java,
-and Python, and offer advice to users of these tools.  We were also
-able to use our methods, plus a novel mutant prioritization scheme, to
+and Python, and offer advice to users of these tools.  Our mutation
+results strongly confirm the wisdom of using multiple tools; with the
+exception of one Java tool, all tools we investigated uniquely
+detected over 1,000 mutants, and for Java and Python there were no
+mutants detected by all tools. For Solidity,
+academic research evaluations of the tools generally agreed strongly
+with our conclusions, but lacked the detail mutant analysis contributed.
+We were also able to use our methods, plus a novel mutant prioritization scheme, to
 identify three useful new detectors for the Slither smart contract
@@
 learning \cite{scholkopf1998nonlinear,tipping1999probabilistic}) are
-the most obvious and interesting possibilites for a better metric.
+the most obvious and interesting possibilities for a better metric.
 Finally, mutant prioritization should be applicable to improving

### 2020-01-24 eold.tex
initial submission

--- 2020-01-23 experiments.tex
@@
 \begin{itemize}
-\item {\bf RQ1:}  Does mutation analysis of static analysis tools produce actionable results?
+\item {\bf RQ1:}  Does our approach produce actionable results?
 That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar in terms of mutation detection capability?
@@
 Can we confirm results from other evaluation methods, or informal opinion?
-\item {\bf RQ4:}  Do mutation scores for clean programs differ from those for programs for which tools report findings before mutation?
-\item {\bf RQ5:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
- different tools are good at/bad at?  Can we extend this information to improve tools by addressing identified weaknesses?
-\item {\bf RQ6:}  How easy is it identify low-hanging fruit for improving tools without prioritization to remove conceptually redundant mutants?
+\item {\bf RQ4:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that different tools are good at/bad at?
+Can we extend this information to improve tools by addressing identified weaknesses?
   \end{itemize}
@@
 This expectation that (some subset of the) mutants can serve as proxies for real, detectable faults is the core concept of our approach.
-{\bf RQ4} addresses a concern briefly mentioned in the introduction:  it is possible that warnings for the original code interfere with our definition of detection.
-The ideal case for our approach is when a tool report no findings for un-mutated code, and either reports a finding (kills the mutant) or not (does not kill it) when the mutant is introduced.
-Chekam et al.
-showed that the ``clean program assumption'' for testing, the idea that coverage will be similar for faulty and fixed versions of a program, is a threat to the validity of investigations of the relationship between coverage and fault detection \cite{CleanProgram}, and we want to establish that this is unlikely to be the case for our approach.
-{\bf RQ5} and {\bf RQ6} receive only preliminary answers, for the smart contract tools.
 
@@
 We analyzed three well-known tools for static analysis of Solidity smart contracts: Slither \cite{slither}, SmartCheck \cite{smartcheck}, and Securify \cite{Securify}.
-Slither, based on an SSA-based intermediate language (SlithIR \cite{slither}) is a tool produced by Trail of Bits.
-SmartCheck, developed by SmartDec, translates Solidity source directly to an XML-based representation, then uses \emph{XPath} patterns to define problems.
-Securify, from SRI Systems Lab at ETH Zurich, works at the bytecode level, first parsing and decompiling contracts, then translating to \emph{semantic facts} in order to look for predefined problems.
 
@@
 We therefore instead selected 100 random contracts for which EtherScan (\url{https://etherscan.io/}) has source code, and used this (quite arbitrary) set of contracts from the actual blockchain to compare tools and identify opportunities for improvement.
-The collected contracts had a total of 15,980 non-comment source lines, as measured by {\tt cloc}, with a mean size of 159.8 LOC and a median size of 108 LOC.
-The largest single contract had 1,127 lines of code.
-Smart contracts are typically smaller than other programs, for various reasons, which explains our use of 100 contracts, compared to a much smaller set of projects for Java and Python.
-The Universal Mutator generated 46,769 mutants for the 100 contracts.
 
@@
 Extracting the signal from Securify's noise will be difficult.
-We also note that while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick (it took about 6 days sequential compute time for Slither and 3 days for SmartCheck), Securify often required many hours to analyze a mutant, and frequently required a few days to analyze a mutant; the full analysis required over three months of compute time.
+We also note that while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick (it took about 8 days sequential compute time for Slither and 3 days for SmartCheck), Securify often required many hours to analyze a mutant, and frequently required a few days to analyze a mutant; the full analysis required over three months of compute time.
 
-For our research questions, {\bf RQ1} is clearly answered in the affirmative.
-Figure \ref{fig:solidityvenn} shows that the tools address quite different problems, with all tools reporting far more uniquely detected mutants than mutants in common with other tools.
-There are only 18 mutants detected by all tools, all of them involving replacement of {\tt msg.sender} (the caller of a smart contract, which may be another smart contract) with {\tt tx.origin} (the original initiator of a sequence of blockchain calls, a ``human'' account).
-Use of {\tt tx.origin} is usuallly a bad idea, can lead to incorrect behavior, and may not be relied on to be meaningful at all in future versions of Ethereum (\url{https://ethereum.stackexchange.com/questions/196/how-do-i-make-my-dapp-serenity-proof}).
+We first compare the two tools that performed best.
+Slither detected 2.37 mean issues over the 100 contracts, and found 39 contracts free of all issues.
+Slither had a mean mutation score of 0.09 over all contracts, and 0.11 over clean (according to Slither) contracts.
+SmartCheck detected 1.89 mean issues over the 100 contracts, but only found 27 contracts free of all issues.
+The mean issue counts for the tools were not statistically significantly different ($p=0.34$).
+SmartCheck had a mean mutation score of 0.05 over all contracts, and 0.03 over clean (according to SmartCheck) contracts.
+The ratio between mean issues (Slither / SmartCheck) is 1.25, and the ratio of numbers of clean contracts is similar (1.44).
+The mutation score ratio, however, is 1.8, suggesting that Slither is not out-scoring SmartCheck \emph{only} because it produces more false positives.
+Slither detects 3,520 mutants not detected by SmartCheck, while SmartCheck only detects 1,629 mutants not detected by Slither (of these, 1,256 were also not detected by Securify) .
+Over the 7 contracts that both tools mark as clean, Slither has a mean mutation score of 0.08, and SmartCheck has a mean mutation score of 0.03; the difference, even with only 7 contracts, is statistically significant by paired Wilcoxon test, at $p=0.03$.
 
-{\bf RQ2} is also answered in the affirmative.
-Counting findings for un-mutated code might suggest that Securify is the best tool, by a wide margin, but in the context of its near-zero mutant ratio, we must suspect that many of the warnings are false positives.
-Slither has the best mutant ratio, but the margin between it and SmartCheck confirms that both tools likely provide value.
+Securify was an outlier, compared to the other tools.
+First, while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick, Securify often required hours to analyze a contract; the full analysis required over a month.
+Second, Securify identified a mean of 24.65 issues on the un-mutated contracts, with a median of 17 issues found; this was a significant difference in issue counts from either other tool, with $p<1.0^{-15}$ in both cases.
+Despite this huge number of non-informational warnings for the original contracts, the mean mutation score was even worse than for SmartCheck, 0.03.
+For contracts where Securify found no issues in the original contract, it \emph{never} detected any mutants.
+Securify, however, did detect 1,449 mutants not detected by Slither, 1,076 of which were not detected by SmartCheck.
+On the other hand, given the ratio of issue counts compared to both tools is over 10.0, it is fairly likely many of these \emph{are} false positives, despite the probability mutants are actual faults.
 
-For {\bf RQ3}, there are only a few tool comparisons in the literature; this is probably due to the fast-moving nature of the blockchain analysis world; the oldest of these tools' publication dates is 2018.
-The most extensive is that of Durieux et al.
-\cite{durieux2019empirical}.
-Slither detected 17\% of known vulnerabilities in their analyis, vs.
-11\% for SmartCheck and 9\% for Securify.
-Slither and SmartCheck were also among the four (out of 9) tools that detected vulnerabilities in the most categories; Securify was not.
-The overall recommendation of Durieux et al.
-was to use a combination of Slither and Mythril \cite{mythril-code} for contract analysis.
-Parizi et al.
-\cite{Parizi} also offer a ranking of tools, and determined that SmartCheck was the most effective, and far more so than Securify; unfortunately, they did not include Slither in their set of evaluated tools.
+Our recommendation, based on mutation analysis results, is that smart contract analysis should at least use both Slither and SmartCheck, perhaps applying only Slither until it reports no interesting issues.
+Securify does offer a substantial non-overlapping set of issues reported, however, but produces a very large number of false positives.
+Available human resources and total number of issues produced (as well as high runtime) might justify not using Securify in some cases.
 
-The Slither paper \cite{slither} also provides an evaluation of all three tools.
-Their findings counts differ from ours because of different choices (we threw out merely informational results), but these are unrelated to mutation analysis, in any case.
-The evaluation only considered reentrancy faults \cite{SurveyAttacks,FC20} (which are sometimes, but only rarely, introduced by mutants).
-For reentrancy, Slither performed best on two real-world large contracts, finding subtle bugs in both, SmartCheck detected the problem in one of the two, and Securify detected neither.
-For a dataset of 1,000 contracts, SmartCheck had a high false positive rate (over 70\%) but detected more actual reentrancies (209) than Slither (99) or Securify (6).
-On the other hand, Slither's low false positive rate of 11\%  makes its results possibly more useful in practice.
-It seems safe to say that
-
-For {\bf RQ4}, on the changes seen when restricting analysis to clean contracts, Slither did slightly better at detecting mutants when the original contract was clean for Slither, and the other two tools did somewhat worse on contracts for which they reported no findings.
-For the three contracts clean for all tools, Slither performed almost exactly as it did over contracts in general, and the other tools performed worse, by about the same margin as they did for their own clean contracts.
-For our approach, we only need a week version of the ``clean program assumption'':  the threat is that kills may be under-reported for non-clean programs, due to interference with findings for the original code.
-It is not a problem if mutation scores are \emph{worse} for programs where a tool reports no findings for the un-mutated code.
-We therefore, for smart contracts, find no threat to our approach arising from the presence of findings on un-mutated code.
-We speculate that ``clean'' results for some tools result from contracts where the tool has trouble with the contract code, but does not actually crash; Slither may do better on clean code because it has fewer such failures, and clean contracts are probably generally simpler and easier to analyze.
-
-\subsubsection{Improving Slither {\bf (RQ5 and RQ6)}}
-
-\begin{figure}
-  {\scriptsize
-\noindent \begin{code}
- 0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.mutant.275.sol:
- if (!p.recipient.send(p.amount)) \{  // Make the payment
- ==>          if (true) \{  // Make the payment
- if (true) \{  // Make the payment
-      \end{code}
-      }
-      \caption{Mutant showing {\bf Boolean constant misuse}.}
-      \label{fig:boolean}
-    \end{figure}
-
-    \begin{figure}
-      {\scriptsize
-\begin{code}
- 0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.mutant.311.sol:
- require(nextDiscountTTMTokenId6 >= 361 \&\& nextDiscountTTMTokenId6 <= 391);
- ==>  ...361...==>...0...
- require(nextDiscountTTMTokenId6 >= 0 \&\& nextDiscountTTMTokenId6 <= 391);
-      \end{code}
-      }
-      \caption{Mutant showing {\bf Type-based tautologies}.}
-      \label{fig:tautology}      
-\end{figure}
-
-\begin{figure}
-  {\scriptsize
-\begin{code}
- 0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.mutant.480.sol:
- byte char = byte(bytes32(uint(x) * 2 ** (8 * j)));
- ==>  ...*...==>.../...
- byte char = byte(bytes32(uint(x) * 2 ** (8 / j)));
-      \end{code}
-      }
-      \caption{Mutant showing {\bf Loss of precision}.}
-      \label{fig:divmul}            
-      \end{figure}
+\subsubsection{Improving Slither {\bf (RQ 4)}}
 
@@
 Examining the first few mutants, we identified three such, based on mutants killed by either SmartCheck or Securify, or both:
+
+\begin{figure*}
+\begin{code}
+  0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.mutant.275.sol:
+  424contracts/0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.sol:249
+        if (!p.recipient.send(p.amount)) \{  // Make the payment
+ ==>          if (true) \{  // Make the payment
+        if (true) \{  // Make the payment
+      \end{code}
+      \caption{Mutant showing {\bf Boolean constant misuse}.}
+      \label{fig:boolean}
+    \end{figure*}
+
+    \begin{figure*}
+\begin{code}
+  0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.mutant.311.sol:
+  424contracts/0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.sol:175
+        require(nextDiscountTTMTokenId6 >= 361 \&\& nextDiscountTTMTokenId6 <= 391);
+ ==>  ...361...==>...0...
+        require(nextDiscountTTMTokenId6 >= 0 \&\& nextDiscountTTMTokenId6 <= 391);
+      \end{code}
+      \caption{Mutant showing {\bf Type-based tautologies}.}
+      \label{fig:tautology}      
+\end{figure*}
+
+\begin{figure*}
+\begin{code}
+  0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.mutant.480.sol:
+  424contracts/0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.sol:650
+        byte char = byte(bytes32(uint(x) * 2 ** (8 * j)));
+ ==>  ...*...==>.../...
+        byte char = byte(bytes32(uint(x) * 2 ** (8 / j)));
+      \end{code}
+      \caption{Mutant showing {\bf Loss of precision due to divide before multiply}.}
+      \label{fig:divmul}            
+      \end{figure*}
 
@@
 
-\item {\bf Loss of precision:}  Solidity only supports integer types.
+\item {\bf Loss of precision due to divide before multiply:}  Solidity only supports integer types.
 This means that performing division before multiplication can introduce rounding that is not present when the multiplication is performed first.
@@
 
-Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants, suggesting that for {\bf RQ6} it is indeed hard to use mutation analysis results without prioritization.
-A large majority of the mutants we inspected involved either the missing
-{\tt return} problem noted in the introduction, or replacing {\tt msg.sender} with {\tt tx.origin}; Slither, of course, has a detector for misuses of {\tt tx.origin}, and we believe (but are not sure) that almost all of the differences with respect to such changes are due to intentional behavior.
-SmartCheck and Securify tend to identify most (though not all) uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule, intended to reduce false positives.
-It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ5} and {\bf RQ6}.
+Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants.
+The majority of cases involved either the missing return problem or replacing {\tt msg.sender} with {\tt tx.origin}; Slither has a detector for misuses of {\tt tx.origin}, and we believe (but are not sure) that almost all of these are due to intentional behavior: SmartCheck and Securify tend to identify most uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule intended to avoid false positives.
+
+It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ4}.
 
 \subsection{Java Tools}
-
-70,825 mutants
 
@@
 \subsubsection{Analysis Results}
-
-Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools, and Tables \ref{tab:scorejava} and \ref{tab:cleanjava} provide numeric results for projects and clean projects, respectively.
-
-In terms of {\bf RQ1}, the raw results are clear:  SpotBugs finds vastly more uniquely killed mutants than the other tools, but PMD contributes a large number of its own unique detections, and clearly offers some independent utility.
-Infer on the other hand, is only able to detect two mutants, but these are unique.
-
-{\bf RQ2} is also answered in the affirmative.
-While the raw kills for SpotBugs are impressive, it also produced fewer findings for clean code, giving it an excellent mutant ratio, making its dominance even more clear.
-PMD's performance was acceptable, and better than any of the Solidity tools (note that comparing ratios across languages is potentially invalid, but not obviously so).
-Infer is confirmed as nearly useless, again.
-
-\begin{figure}
-  \url{https://stackoverflow.com/questions/4297014/what-are-the-differences-between-pmd-and-findbugs}
-  \url{https://www.sw-engineering-candies.com/blog-1/comparison-of-findbugs-pmd-and-checkstyle}
-  \url{https://www.reddit.com/r/java/comments/3i7w6n/checkstyle_vs_pmd_vs_findbugs_for_dummies_why/}
-\caption{Blog posts and forum threads discussing Java static analysis tools.}
-\label{fig:blog}
-\end{figure}
-
-For {\bf RQ3}, to our knowledge there is no academic comparison of these tools.
-However, the user postings listed in Figure \ref{fig:blog}, plus personal communications with security analysts who use these tools (citation removed for now, due to potential to compromise blinding) supported some basic conclusions.
-SpotBugs/FindBugs is indeed the best tool for finding bugs; PMD focuses more on stylistic issues and has a weaker semantic model.
-Mutants are much more likely to be bugs than stylistic issues (they alter semantics, not code style, in almost all cases: deletion or constant/operator changes obviously have no style impact).
-Infer is closer to a model checker focusing on resource leaks than a truly general-purpose tool, arguably.
-
-For {\bf RQ4} there were very few clean projects, but the mutation scores for Infer were unchanged, and for SpotBugs the mean mutation score improved and median mutation score was worse.
-Again, we see no evidence that non-clean code is a source of significant degradation on mutation detection.
 
@@
     \hline
-    SpotBugs & 0.47 & 0.00 & 0.37 & 0.21 & 0.7847 \\
+    SpotBugs & 0.26 & 0.00 & 0.30 & 0.07 & 1.132 \\
     \hline
-    PMD & 1.27 & 0.00 & 0.07 & 0.07 & 0.0562 \\
+    PMD & 0.48 & 0.00 & 0.09 & 0.00 & 0.183 \\
     \hline
-    Infer & 0.07 & 0.00 & 0.00 & 0.00 &  0.0002 \\
+    Infer & 0.00 & 0.00 & 0.00 & 0.00 &  0.006 \\
     \hline
   \end{tabular}
-  \caption{Java tools: number of findings and mutation scores over all projects.}
+  \caption{Java tools: number of findings and mutation scores over all files.}
   \label{tab:scorejava}
@@
 \begin{table}
-  \begin{tabular}{l|r|r|r}
-    & & \multicolumn{2}{|c|}{} \\
-    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} \\
-    Tool & Projects & Mean & Median \\
+  \begin{tabular}{l|r|r|r|r|r}
+    & & \multicolumn{2}{|c|}{} & \multicolumn{2}{|c}{Clean For All (1,185)} \\
+    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} &  \multicolumn{2}{|c}{Mutation Score}\\
+    Tool & Files & Mean & Median & Mean & Median\\
     \hline
     \hline
-    SpotBugs & 3 & 0.41 & 0.17 \\
+    SpotBugs & 1,426 & 0.32 & 0.05 & 0.37 & 0.06 \\
     \hline
-    PMD & 0 & N/A & N/A \\
+    PMD & 1,359 & 0.08 & 0.00 & 0.08 & 0.00 \\
     \hline
-    Infer & 6 & 0.00 & 0.00 \\
+    Infer & 1,598 & 0.00 & 0.00 & 0.00 & 0.00 \\
     \hline
   \end{tabular}
-  \caption{Java tools: clean project counts and mutation scores.  No projects were clean for all tools, due to PMD producing findings for all projects.}
+  \caption{Java tools: clean file counts and mutation scores.}
   \label{tab:cleanjava}
@@
 
+Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools.
+
 \subsection{Python Tools}
-
-167,511 mutants
 
@@
 
-Figure \ref{fig:pythonvenn} shows the mutants killed by the Java analysis tools, and Tables \ref{tab:scorepython} and \ref{tab:cleanpython} provide numeric results for projects and clean projects, respectively.
-
-For {\bf RQ1}, Pylint and pyflakes both killed more than 15,000 mutants killed by no other tool; Pylint killed more mutants, but the relative difference was much smalller than in our Solidity or Java analysis.
-It is clear both tools are well worth using, assuming these are not false positives.
-Pyflakes' mutation score is lower than unique kills might suggest because it has almost no overlap with the other two tools.
-It is doing something quite different than they are, apparently.
-
-The mutant ratios {\bf RQ2} suggest that PyChecker might be better than pyflakes, in fact, and better than Pylint, because while it killed fewer mutants, it also produced fewer findings by far, overall.
-PyChecker would be a good choice for quick and dirty analysis of code on the fly, perhaps, in that it had excellent mutation scores for some programs (hence a high mean) but is relatively ``quiet'' compared to the other tools.
-However, its very low median mutation score is troubling; it may simply not work for some code.
-A combination of pyflakes and Pylint seems safer, though adding the relatively low-findings PyChecker seems likely to have little cost.
-For Python, we see that all of our measures are required to get a good picture of the tools.
-
-\begin{figure}
-  \url{https://stackoverflow.com/questions/1428872/pylint-pychecker-or-pyflakes}
-  \url{https://www.reddit.com/r/Python/comments/ii3gm/experience_with_pylint_pychecker_pyflakes/}
-  \url{https://www.slant.co/versus/12630/12631/~pylint_vs_pyflakes}
-  \url{https://news.ycombinator.com/item?id=12748885}
-  \url{https://doughellmann.com/blog/2008/03/01/static-code-analizers-for-python/}
-\caption{Blog posts and forum threads discussing Python static analysis tools.}
-\label{fig:blogpython}
-\end{figure}
-
-For {\bf RQ3}, there were again no academic comparisons we could find.
-However, opinions on the web were quite common (see Figure \ref{fig:blogpython}, which lists ones we examined).
-It is hard to summarize the overall opinion here, since it ranges considerably.
-There is probably general agreement that PyChecker is old and maybe less useful, but also terse and sometimes helpful.
-Pylint is the most recommended tool, and the general complaint that it is too picky was mitigated in our results by turning off clearly informative-only findings; users not configuring Pylint will probably see even more killed mutants, but likely a worse mutant ratio because mutants tend to introduce non-stylistic problems.
-Pyflakes is also well liked, and supposedly less verbose than Pylint; we did not see this in our results, but that may be because we configured Pylint, and pyflakes aims to focus on ``actual bugs'' so it is equally verbose for mutants, which are likely to be actual bugs.
-
-For {\bf RQ4} there were no clean projects for the known-to-be-aggressive Pylint.
-Performance for the other two tools was worse on projects for which they were clean, again failing to support the possibility that our definition of killing a mutant under-reports detection for non-clean code.
-
 \begin{figure}
@@
   \caption{Mutants killed by Python static analysis tools.}
-  \label{fig:pythonvenn}
+  \label{fig:javavenn}
 \end{figure}
@@
     \hline
-    Pylint & 2.24 & 1.00 & 0.22 & 0.20 & 0.097 \\
+    SpotBugs & 0.26 & 0.00 & 0.30 & 0.07 & 1.132 \\
     \hline
-    pyflakes & 2.4 & 1.00 & 0.03 & 0.00 & 0.012 \\
+    PMD & 0.48 & 0.00 & 0.09 & 0.00 & 0.183 \\
     \hline
-    PyChecker & 0.56& 0.00 & 0.10 & 0.00 &  0.183 \\
+    Infer & 0.00 & 0.00 & 0.00 & 0.00 &  0.006 \\
     \hline
   \end{tabular}
-  \caption{Python tools: number of findings and mutation scores over all projects.}
-  \label{tab:scorepython}
+  \caption{Java tools: number of findings and mutation scores over all files.}
+  \label{tab:scorejava}
 \end{table}
@@
 \begin{table}
-  \begin{tabular}{l|r|r|r}
-    & & \multicolumn{2}{|c|}{} \\
-    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} \\
-    Tool & Projects & Mean & Median \\
+  \begin{tabular}{l|r|r|r|r|r}
+    & & \multicolumn{2}{|c|}{} & \multicolumn{2}{|c}{Clean For All (1,185)} \\
+    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} &  \multicolumn{2}{|c}{Mutation Score}\\
+    Tool & Files & Mean & Median & Mean & Median\\
     \hline
     \hline
-    Pylint & 0 & N/A & N/A \\
+    SpotBugs & 1,426 & 0.32 & 0.05 & 0.37 & 0.06 \\
     \hline
-    pyflakes & 12 & 0.02 & 0.00 \\
+    PMD & 1,359 & 0.08 & 0.00 & 0.08 & 0.00 \\
     \hline
-    PyChecker & 21& 0.01 & 0.00 \\
+    Infer & 1,598 & 0.00 & 0.00 & 0.00 & 0.00 \\
     \hline
   \end{tabular}
-  \caption{Python tools: clean project counts and mutation scores.  No projects were clean for all tools, due to Pylint producing findings for all projects.}
-  \label{tab:cleanpython}
+  \caption{Java tools: clean file counts and mutation scores.}
+  \label{tab:cleanjava}
 \end{table}
@@
 \subsection{Threats to Validity}
-
-The primary threat to validity in terms of generalization is that we only examined nine static analysis tools, and our analyses were restricted to 100 smart contracts, 15 Java projects, and 25 Python projects.
-Because it is hard to identify a ground truth to compare with (the motivation for our approach), we cannot be certain that our rankings of tools are correct even for these tools and this code.
-However, where there are existing discussions of the tools, our results seem to agree with these, but add substantial detail.
-
-A general threat to our approach is that mutants may differ substantially from ``real'' faults, in some way.
-This is certainly true, in a sense \cite{GopinathMutants}, but for static analysis purposes, the risk is that some mutation operators align with patterns a particular tool identifies, biasing the evaluation in favor of that tool.
-Such faults may be dis-proportionately present in mutants vs.
-real code.
-However, we consider this unlikely.
-The vast majority of applied mutation operations for all of our experiments were highly generic, and do not plausibly represent a pattern in which some tool might specialize.
-Statement deletions leave no ``trace'' for a tool to match against, but only an omission.
-Changing arithmetic and comparison operators and numeric constants (incrementing, decrementing, or changing to 0 or 1) account for most of the non-statement-deletion mutants.
-We cannot see how any tool could have a (useful) detection rule that looks for these kinds of changes (a tool obviously would not flag all instances of 0 and 1 constants, or addition).
-Finally, many mutants (though many fewer than in these categories) add {\tt break} or {\tt continue} statements.
-Again, we do not see how a tool could take advantage of this without either genuinely understanding likely-bad control flow patterns (in which case, we think it ``deserves'' to kill more mutants) or flagging so many instances of these it would have a bad mutant ratio.

### 2020-01-24 experiments.tex
initial submission

@@
 There are only 18 mutants detected by all tools, all of them involving replacement of {\tt msg.sender} (the caller of a smart contract, which may be another smart contract) with {\tt tx.origin} (the original initiator of a sequence of blockchain calls, a ``human'' account).
-Use of {\tt tx.origin} is usuallly a bad idea, can lead to incorrect behavior, and may not be relied on to be meaningful at all in future versions of Ethereum (\url{https://ethereum.stackexchange.com/questions/196/how-do-i-make-my-dapp-serenity-proof}).
+Use of {\tt tx.origin} is usually a bad idea, can lead to incorrect behavior, and may not be relied on to be meaningful at all in future versions of Ethereum (\url{https://ethereum.stackexchange.com/questions/196/how-do-i-make-my-dapp-serenity-proof}).
 
@@
 \cite{durieux2019empirical}.
-Slither detected 17\% of known vulnerabilities in their analyis, vs.
+Slither detected 17\% of known vulnerabilities in their analysis, vs.
 11\% for SmartCheck and 9\% for Securify.
@@
 
-70,825 mutants
-
 \subsubsection{Static Analysis Tools Compared}
 
+For Java, we again compared three tools.
+SpotBugs (\url{https://spotbugs.github.io/}) is the ``spiritual successor'' of FindBugs \cite{FindBugs,CompareJavaTools}.
+PMD (\url{https://pmd.github.io/}) \cite{CompareJavaTools} is an extensible cross-language static code analyzer.
+FaceBook's Infer (\url{https://fbinfer.com/}) \cite{Infer} focuses on diff-related detection of serious errors (concurrency, memory safety, and information flow) and aims to scale to very large code bases.
+
 \subsubsection{Project Selection}
+
+For Java and Python, we did not have to worry about invalidating tool improvements by basing our results on benchmark code.
+We therefore aimed to use realistic, important source code.
+We selected top GitHub projects (defined by number of stars) for each language, and removed projects with fewer than 5 developers or less than six months of commit history (as well as projects that did not build).
+For Java, we analyzed the top 15 projects satisfying our criteria, with a maximum of 623,355 LOC and a minimum of 3,957 LOC, and a total size of 1.8 million LOC.
+The Java compiler rejected a large number of the generated mutants, leaving us with just over 70,000 mutants to analyze, ranging from 136 to 10,016 per project.
 
@@
 Infer on the other hand, is only able to detect two mutants, but these are unique.
+Unfortunately, these mutants only served to indicate that we were actually running Infer in such a way that it \emph{could} detect problems:  both were concurrency warnings, but the code change in both cases was simply removing a (semantically inoperative) {\tt \@Override} annotation.
+It may be that the diff sizes in our code were simply too small for Infer's approach.
 
@@
 
-For {\bf RQ3}, to our knowledge there is no academic comparison of these tools.
+For {\bf RQ3}, to our knowledge there is no recent academic comparison of these tools; the most relevant study dates from 2004 \cite{CompareJavaTools}, used FindBugs, not SpotBugs, and reached no strong conclusions with respect to FindBugs vs.
+PMD.
 However, the user postings listed in Figure \ref{fig:blog}, plus personal communications with security analysts who use these tools (citation removed for now, due to potential to compromise blinding) supported some basic conclusions.
@@
 
-167,511 mutants
-
 \subsubsection{Static Analysis Tools Compared}
 
+We compared three widely used and well-known Python tools:  Pylint \url{https://www.pylint.org/} (probably the most widely used of Python bug finding tools), pyflakes \url{https://pypi.org/project/pyflakes/}, designed to be faster, lighter-weight, and more focused on bugs (without configuration) than Pylint, and PyChecker \url{http://pychecker.sourceforge.net/}, an older, but still used (e.g., we found numerous {\tt .travis.yml} files installing and running PyChecker), tool.
+
 \subsubsection{Project Selection}
+
+For Python, we analyzed the top 25 GitHub projects by our criteria (see above), due to the smaller size of Python projects.
+These ranged in size from 3,377 LOC to 34,671 LOC, with a total size of just under 100 KLOC.
+For Python, we were able to produce 167,511 valid mutants to analyze.
 
@@
 
-Figure \ref{fig:pythonvenn} shows the mutants killed by the Java analysis tools, and Tables \ref{tab:scorepython} and \ref{tab:cleanpython} provide numeric results for projects and clean projects, respectively.
+Figure \ref{fig:pythonvenn} shows the mutants killed by the Python analysis tools, and Tables \ref{tab:scorepython} and \ref{tab:cleanpython} provide numeric results for projects and clean projects, respectively.
 
-For {\bf RQ1}, Pylint and pyflakes both killed more than 15,000 mutants killed by no other tool; Pylint killed more mutants, but the relative difference was much smalller than in our Solidity or Java analysis.
+For {\bf RQ1}, Pylint and pyflakes both killed more than 15,000 mutants killed by no other tool; Pylint killed more mutants, but the relative difference was much smaller than in our Solidity or Java analysis.
 It is clear both tools are well worth using, assuming these are not false positives.

### 2020-01-24 introduction.tex
initial submission

@@
 SmartCheck, on the other hand, detects removing the {\tt return} in {\tt get} or replacing it with a {\tt revert}, or removing the {\tt public} visibility modifier for {\tt get} \footnote{Slither's ``missing return'' detector is only available in the private version of slither, or through the {\tt crytic} service provided by Trail of Bits.}.
-If we restrict our analyis to findings with a severity greater than informational, SmartCheck detects no mutants of the contract, while Slither still reports that some mutants \emph{allow an arbitrary caller to cause the contract to self destruct}.
+If we restrict our analysis to findings with a severity greater than informational, SmartCheck detects no mutants of the contract, while Slither still reports that some mutants \emph{allow an arbitrary caller to cause the contract to self destruct}.
 This simple example shows how our approach works on a small scale.

### 2020-01-24 old.tex
initial submission

--- 2020-01-24 eold.tex
@@
 \subsection{Java Tools}
-
-\subsubsection{Static Analysis Tools Compared}
-
-\subsubsection{Project Selection}
-
-\subsubsection{Analysis Results}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{java.png}
-  \caption{Mutants killed by Java static analysis tools.}
-  \label{fig:javavenn}
-\end{figure}
-
-\begin{table}
-  \begin{tabular}{l|r|r|r|r|r}
-    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
-    Tool & Mean & Median & Mean & Median & Ratio\\
-    \hline
-    \hline
-    SpotBugs & 0.26 & 0.00 & 0.30 & 0.07 & 1.132 \\
-    \hline
-    PMD & 0.48 & 0.00 & 0.09 & 0.00 & 0.183 \\
-    \hline
-    Infer & 0.00 & 0.00 & 0.00 & 0.00 &  0.006 \\
-    \hline
-  \end{tabular}
-  \caption{Java tools: number of findings and mutation scores over all files.}
-  \label{tab:scorejava}
-\end{table}
-
-\begin{table}
-  \begin{tabular}{l|r|r|r|r|r}
-    & & \multicolumn{2}{|c|}{} & \multicolumn{2}{|c}{Clean For All (1,185)} \\
-    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} &  \multicolumn{2}{|c}{Mutation Score}\\
-    Tool & Files & Mean & Median & Mean & Median\\
-    \hline
-    \hline
-    SpotBugs & 1,426 & 0.32 & 0.05 & 0.37 & 0.06 \\
-    \hline
-    PMD & 1,359 & 0.08 & 0.00 & 0.08 & 0.00 \\
-    \hline
-    Infer & 1,598 & 0.00 & 0.00 & 0.00 & 0.00 \\
-    \hline
-  \end{tabular}
-  \caption{Java tools: clean file counts and mutation scores.}
-  \label{tab:cleanjava}
-\end{table}
-
-Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools.
-
-\subsection{Python Tools}
-
-\subsubsection{Static Analysis Tools Compared}
-
-\subsubsection{Project Selection}
-
-\subsubsection{Analysis Results}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{python.png}
-  \caption{Mutants killed by Python static analysis tools.}
-  \label{fig:javavenn}
-\end{figure}
-
-\begin{table}
-  \begin{tabular}{l|r|r|r|r|r}
-    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
-    Tool & Mean & Median & Mean & Median & Ratio\\
-    \hline
-    \hline
-    SpotBugs & 0.26 & 0.00 & 0.30 & 0.07 & 1.132 \\
-    \hline
-    PMD & 0.48 & 0.00 & 0.09 & 0.00 & 0.183 \\
-    \hline
-    Infer & 0.00 & 0.00 & 0.00 & 0.00 &  0.006 \\
-    \hline
-  \end{tabular}
-  \caption{Java tools: number of findings and mutation scores over all files.}
-  \label{tab:scorejava}
-\end{table}
-
-\begin{table}
-  \begin{tabular}{l|r|r|r|r|r}
-    & & \multicolumn{2}{|c|}{} & \multicolumn{2}{|c}{Clean For All (1,185)} \\
-    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} &  \multicolumn{2}{|c}{Mutation Score}\\
-    Tool & Files & Mean & Median & Mean & Median\\
-    \hline
-    \hline
-    SpotBugs & 1,426 & 0.32 & 0.05 & 0.37 & 0.06 \\
-    \hline
-    PMD & 1,359 & 0.08 & 0.00 & 0.08 & 0.00 \\
-    \hline
-    Infer & 1,598 & 0.00 & 0.00 & 0.00 & 0.00 \\
-    \hline
-  \end{tabular}
-  \caption{Java tools: clean file counts and mutation scores.}
-  \label{tab:cleanjava}
-\end{table}
-
-\subsection{Threats to Validity}

### 2020-01-24 relatedwork.tex
initial submission

@@
 This is a result of both algorithmic constraints and basic engineering trade-offs.
+While comparisons of static analysis tools \cite{CompareJavaTools,durieux2019empirical, Parizi,slither} have appeared in the literature for years, these generally involved large human effort and resulting smaller scale, or did not make a strong effort to address false positives, or restricted analysis to, e.g., a known defects set \cite{AllBugs}.
 
@@
 The primary differences are that their seed programs must be warning-free (which greatly limits the set of input programs available) and their tool must parse and understand the programs, and that the new versions are based on adding new assertions, not ``breaking'' the original code.
-We allow arbitarily buggy seed programs (thus many more real programs can be used), and can, due to the any-language nature of the mutation generator we use, operate even in new languages without further development effort.
+We allow arbitrarily buggy seed programs (thus many more real programs can be used), and can, due to the any-language nature of the mutation generator we use, operate even in new languages without further development effort.
 Further, their approach only identifies problems when tools are outliers compared to numerous other tools in either detecting a bug (precision) or not detecting it (soundness), and so requires comparing multiple tools.

### 2020-01-24 experiments.tex
josselin good changes

@@
 
-We used the Universal Mutator \cite{universalmutator,regexpMut}, which aggressively produces large numbers of mutants, to generate all mutants.
-We will make available a repository containing our full raw results plus details on contracts and projects selected, for replication and further analysis, upon acceptance (to avoid breaking blinding).
-
 \subsection{Solidity Smart Contract Tools}
@@
 The largest single contract had 1,127 lines of code.
-Smart contracts are typically smaller than other programs, for various reasons, which explains our use of 100 contracts, compared to a much smaller set of projects for Java and Python.
-The Universal Mutator generated 46,769 mutants for the 100 contracts.
+The Universal Mutator generated 46,769 valid mutants for these 100 contracts.
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{solidity.png}
+  \includegraphics[width=0.8\columnwidth]{solidity.png}
   \caption{Mutants killed by Solidity static analysis tools.}
@@
 Without prioritization, on average a developer would have to look at more than 200, 80, and 400 mutants, respectively, to find instances of these problems.
+Interestingly, the very fact that these instances are ``needles in a haystack'' among the mutants not killed by Slither means that the results in Figure \ref{fig:solidityvenn} and Tables \ref{tab:scoresolidity} and \ref{tab:cleansolidity} are almost unaltered by our improvements to Slither: our analysis is fairly robust to modest tool improvements, unless added detectors account for a large number of mutants not detected by the tool.
+Adding such detectors will also only improve mutation ratio if they do not add many false positives.
+Substantial changes in results therefore require adding very effective (for mutants) detectors that seldom trigger for correct code (or at least trigger much less than for mutants).
+``Cheating'' with respect to a mutation benchmark is thus, we hope, difficult without actual major tool improvements.
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{java.png}
+  \includegraphics[width=0.8\columnwidth]{java.png}
   \caption{Mutants killed by Java static analysis tools.}
@@
 Infer is closer to a model checker focusing on resource leaks than a truly general-purpose tool, arguably.
+In fact, we suspect Infer would perform much better if we crafted complex mutation operators targeting some important subtle Java bugs; it is still fair to say that Infer is probably not a good \emph{general-purpose} static analysis tool for Java.
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{python.png}
+  \includegraphics[width=0.8\columnwidth]{python.png}
   \caption{Mutants killed by Python static analysis tools.}
@@
 
-A general threat to our approach is that mutants may differ substantially from ``real'' faults, in some way.
-This is certainly true, in a sense \cite{GopinathMutants}, but for static analysis purposes, the risk is that some mutation operators align with patterns a particular tool identifies, biasing the evaluation in favor of that tool.
-Such faults may be dis-proportionately present in mutants vs.
-real code.
-However, we consider this unlikely.
-The vast majority of applied mutation operations for all of our experiments were highly generic, and do not plausibly represent a pattern in which some tool might specialize.
-Statement deletions leave no ``trace'' for a tool to match against, but only an omission.
-Changing arithmetic and comparison operators and numeric constants (incrementing, decrementing, or changing to 0 or 1) account for most of the non-statement-deletion mutants.
-We cannot see how any tool could have a (useful) detection rule that looks for these kinds of changes (a tool obviously would not flag all instances of 0 and 1 constants, or addition).
-Finally, many mutants (though many fewer than in these categories) add {\tt break} or {\tt continue} statements.
-Again, we do not see how a tool could take advantage of this without either genuinely understanding likely-bad control flow patterns (in which case, we think it ``deserves'' to kill more mutants) or flagging so many instances of these it would have a bad mutant ratio.
+We used the Universal Mutator \cite{universalmutator,regexpMut}, which aggressively produces large numbers of mutants, but does not target any particular software defect patterns, to generate all mutants.
+We will make available a repository containing our full raw results plus details on contracts and projects selected, for replication and further analysis, upon acceptance (to avoid breaking blinding).

### 2020-01-24 introduction.tex
josselin good changes

@@
 
-We can quantitatively express this relationship between detecting mutants and findings for the original program.
+We can quantitatively express the relationship between detecting mutants and findings for the original program.
 Since we are comparing over the same program or set of programs, we simply divide the mean \emph{mutation score} ($\frac{|\mathit{killed}||}{|\mathit{mutants}|}$, the ratio of killed mutants to all mutants) by the mean number of findings.
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{example.png}
+  \includegraphics[width=0.8\columnwidth]{example.png}
   \caption{Mutants killed by three static analysis tools.}
@@
 We therefore also introduce the idea of prioritizing mutants, using a Furthest-Point-First \cite{Gonzalez} algorithm and distance metric inspired by previous work on helping developers sort through failing test cases and avoid duplicates \cite{PLDI13}, to help static analysis tool developers find interesting patterns without wading through numerous uninteresting duplicates.
+
+A general objection to our approach is that mutants may differ substantially from ``real'' faults, in some way.
+This is certainly true, in a sense \cite{GopinathMutants}, but for static analysis purposes does not matter.
+The real risk is that some mutation operators align with patterns a particular tool identifies, biasing the evaluation in favor of that tool.
+Such faults may be dis-proportionately present in mutants vs.
+real code.
+However, we consider this unlikely.
+The vast majority of applied mutation operations for all of our experiments were highly generic, and do not plausibly represent a pattern in which some tool might specialize.
+Statement deletions leave no ``trace'' for a tool to match against, but only an omission.
+Changing arithmetic and comparison operators and numeric constants (incrementing, decrementing, or changing to 0 or 1) account for most of the non-statement-deletion mutants.
+We cannot see how any tool could have a (useful) detection rule that looks for these kinds of changes (a tool obviously would not flag all instances of 0 and 1 constants, or of addition, and remain a useful tool).
+Finally, many mutants (though many fewer than in these categories) add {\tt break} or {\tt continue} statements.
+Again, we do not see how a tool could take advantage of this without either genuinely understanding likely-bad control flow patterns (in which case, we think it ``deserves'' to kill more mutants) or flagging so many instances of these it would have a bad mutant ratio.
 
@@
 Consider the code in Figure \ref{fig:sol424intro}, from the Solidity 0.4.24 ``Introduction to Smart Contracts''.
-The Universal Mutator tool \cite{universalmutator,regexpMut}, which has been extensively tuned for Solidity mutation (and is the only smart contract mutation tool referenced in the Solidity documentation (\url{https://solidity.readthedocs.io/en/v0.5.12/resources.html})) produces seven valid, non-redundant mutants for this trivial example code.
+The Universal Mutator tool \cite{universalmutator,regexpMut}, which has been extensively tuned for Solidity's grammar (though not to target any particular vulnerabilities), and is the only smart contract mutation tool referenced in the Solidity documentation (\url{https://solidity.readthedocs.io/en/v0.5.12/resources.html}), produces seven valid, non-redundant mutants for this trivial example code.
 Both the public version of Slither \cite{slither} and SmartCheck \cite{smartcheck} produce a small number (three and two, respectively) of low-severity, informational, warnings for this code.

### 2020-01-25 introduction.tex
Gustavo edits

@@
 The tool cannot simply flag all code as having a problem or it will perform poorly at the task of \emph{distinguishing} the mutated code from non-mutated (and presumably at least \emph{more} correct) code.
-Based on verification and testing uses of mutation, it is safe to say that at least 50, and likely 60-70\% or more, of mutants that are not semantically equivalent to the original code are actually fault-inducing, so the task presented to a static analysis tool is the generalization of the task we ideally expect static analysis to perform:  to identify faulty code, without executing it, and, most critically, \emph{to distinguish faulty from correct code}.
+Based on verification and testing uses of mutation, it is safe to say that usually at least 40\%, and often 50-60\% or more, of mutants are not semantically equivalent to the original code, and are actually fault-inducing \cite{TCE,groce2015verified,groce2018verified,impactEquiv,smith2009should}, so the task presented to a static analysis tool is the generalization of the task we ideally expect static analysis to perform:  to identify faulty code, without executing it, and, most critically, \emph{to distinguish faulty from correct code}.
 Obviously, many faults cannot be identified statically without a complete specification, or without unreasonable analysis cost and precision, but the measure of performance here is meant to be mostly \emph{relative} to other tools applied to the same code; this is primarily a \emph{differential} approach.

### 2020-01-25 prioritizing.tex
Gustavo edits

@@
 
-We do not claim this is an optimal, or even tuned, metric.  It
-provided useful results for a different problem, ranking mutants in an
+We do not claim this is an optimal, or even tuned, metric; in the long run, it may be completely replaced.  It did
+provide useful results for a different problem, ranking mutants in an
 effort similar to that of Groce et
 al. \cite{groce2015verified,groce2018verified} to examine unkilled
-mutants for a smart contract library tested using the Echidna fuzzer,
+mutants for a smart contract library tested using the Echidna fuzzer \cite{echidna-code},
 and we simply adopted it for our differential mutation analysis.

### 2020-01-25 experiments.tex
Various fixes

@@
 \begin{figure}
-  \includegraphics[width=0.8\columnwidth]{solidity.png}
+  \includegraphics[width=0.7\columnwidth]{solidity.png}
   \caption{Mutants killed by Solidity static analysis tools.}
@@
 
+Based on the differential mutation analysis, we identified three low-hanging fruit to improve the performance of Slither.
+The process was simple.
+First, we produced a list of all mutants killed by either SmartCheck or Securify, but not killed by Slither.
+We then applied the prioritization method based on the FPF algorithm and the distance metric described in Section \ref{sec:prioritizing}, and examined the mutants in rank order.
+Many of the mutants were difficult to identify as true or false positives, absent context.
+Some opportunities for enhancement were clear, but seemed likely to require considerable effort to implement without producing a large number of false positives.
+For example, Securify often detected when an ERC20 token contract's guard preventing making the special 0x0 address the owner of a contract was removed, and issued the error {\tt Violation for MissingInputValidation}.
+Detecting such missing guards is probably useful, but formulating a way to do it without producing false positives is non-trivial.
+We wanted to show that mutants could identify \emph{useful} but \emph{easy to implement} missing detectors.
+Examining the first few mutants, we identified three such, based on mutants killed by either SmartCheck or Securify, or both:
+
 \begin{figure}
@@
 
-Based on the differential mutation analysis, we identified three low-hanging fruit to improve the performance of Slither.
-The process was simple.
-First, we produced a list of all mutants killed by either SmartCheck or Securify, but not killed by Slither.
-We then applied the prioritization method based on the FPF algorithm and the distance metric described in Section \ref{sec:prioritizing}, and examined the mutants in rank order.
-Many of the mutants were difficult to identify as true or false positives, absent context.
-Some opportunities for enhancement were clear, but seemed likely to require considerable effort to implement without producing a large number of false positives.
-For example, Securify often detected when an ERC20 token contract's guard preventing making the special 0x0 address the owner of a contract was removed, and issued the error {\tt Violation for MissingInputValidation}.
-Detecting such missing guards is probably useful, but formulating a way to do it without producing false positives is non-trivial.
-We wanted to show that mutants could identify \emph{useful} but \emph{easy to implement} missing detectors.
-Examining the first few mutants, we identified three such, based on mutants killed by either SmartCheck or Securify, or both:
-
 \begin{enumerate}
@@
 For example, if {\tt x} is an unsigned integer type, the comparison {\tt x >= 0} is always true and {\tt x < 0} is always false.
-This detector is a generalization of the SmartCheck detector \url{https://github.com/smartdec/smartcheck/blob/master/rule\_descriptions/SOLIDITY\_UINT\_CANT\_BE\_NEGATIVE/description\_en.html}, modified to actually compute the ranges of types and identify more general instances of tautological comparisons, e.g.
+This detector is a generalization of the SmartCheck detector \url{https://github.com/smartdec/smartcheck/blob/master/rule\_descriptions/SOLIDITY\_UINT\_CANT\_BE\_NEGATIVE/}, modified to actually compute the ranges of types and identify more general instances of tautological comparisons, e.g.
 {\tt y < 512} where {\tt y}'s type is {\tt int8}.
@@
 This is a fairly important problem, given the frequency with which Solidity code performs financial calculations where maximum precision is desired.
-SmartCheck provides a detector for such precision losses \url{https://github.com/smartdec/smartcheck/blob/master/rule\_descriptions/SOLIDITY\_DIV\_MUL/description\_en.html}, which enabled it to detect mutants such as the one shown in Figure\ref{fig:divmul}.
+SmartCheck provides a detector for such precision losses \url{https://github.com/smartdec/smartcheck/blob/master/rule\_descriptions/SOLIDITY\_DIV\_MUL/}, which enabled it to detect mutants such as the one shown in Figure\ref{fig:divmul}.
 \end{enumerate}
@@
 \begin{figure}
-  \includegraphics[width=0.8\columnwidth]{java.png}
+  \includegraphics[width=0.7\columnwidth]{java.png}
   \caption{Mutants killed by Java static analysis tools.}
@@
 \begin{figure}
-  \includegraphics[width=0.8\columnwidth]{python.png}
+  \includegraphics[width=0.7\columnwidth]{python.png}
   \caption{Mutants killed by Python static analysis tools.}

### 2020-01-25 introduction.tex
Various fixes

--- 2020-01-24 introduction.tex
@@
 On the other hand, if two tools are very similar in what they detect, the effort of wading through duplicate results may not be a good use of time.
-Developers of static analysis tools also want to be able to compare their tools to others targeting the same domain, in order to see what detection methods (or tweaks to precision/soundness tradeoffs) they might want to imitate.
-Unfortunately, comparing tools is hard, and would seem to require vast manual effort to inspect findings and determine ground truth, to obtain any statistical validity.
+Developers of static analysis tools also want to be able to compare their tools to others targeting the same domain, in order to see what detection methods (or tweaks to precision/soundness trade-offs) they might want to imitate.
+Unfortunately, comparing static analysis tools is hard, and would seem to require vast manual effort to inspect findings and determine ground truth on a scale that would provide statistical confidence.
 
@@
 
-Combining a differential approach (not differential \emph{testing}, precisely, in that individual differences are not always worth inspecting) and mutation analysis, however, offers a novel way to compare static analysis tools, one useful both to users wishing to select a good tool or set of tools, and to developers hoping to improve their own tools.
+Combining a differential approach (not differential \emph{testing}, precisely, in that individual differences are not always worth inspecting) and mutation analysis, however, offers a novel way to compare static analysis tools, one useful to users wishing to select a good tool or set of tools, researchers interested in, e.g., the impact of precision/soundness trade-offs or different intermediate languages, and developers of static analysis tools hoping to improve their tools by systematic examination of code changes that are likely fault-inducing but not detected.
 
@@
 The tool cannot simply flag all code as having a problem or it will perform poorly at the task of \emph{distinguishing} the mutated code from non-mutated (and presumably at least \emph{more} correct) code.
-Based on verification and testing uses of mutation, it is safe to say that at least 50, and likely 60-70\% or more, of mutants that are not semantically equivalent to the original code are actually fault-inducing, so the task presented to a static analysis tool is the generalization of the task we ideally expect static analysis to perform:  to identify faulty code, without executing it, and, most critically, \emph{to distinguish faulty from correct code}.
+Based on verification and testing uses of mutation, it is safe to say that usually at minimum 40\%, often 50-60\%, and frequently up to 80\%+ \cite{mutKernel,groce2018verified}, of mutants are not semantically equivalent to the original code, and are actually fault-inducing \cite{TCE,impactEquiv,smith2009should}, so the task presented to a static analysis tool is the generalization of the task we ideally expect static analysis to perform:  to identify faulty code, without executing it, and, most critically, \emph{to distinguish faulty from correct code}.
 Obviously, many faults cannot be identified statically without a complete specification, or without unreasonable analysis cost and precision, but the measure of performance here is meant to be mostly \emph{relative} to other tools applied to the same code; this is primarily a \emph{differential} approach.
@@
 \begin{figure}
-  \includegraphics[width=0.8\columnwidth]{example.png}
+  \includegraphics[width=0.7\columnwidth]{example.png}
   \caption{Mutants killed by three static analysis tools.}
@@
 As with efforts to improve test suites, manually searching through all mutants can be an onerous task, especially for large-scale evaluations.
-We therefore also introduce the idea of prioritizing mutants, using a Furthest-Point-First \cite{Gonzalez} algorithm and distance metric inspired by previous work on helping developers sort through failing test cases and avoid duplicates \cite{PLDI13}, to help static analysis tool developers find interesting patterns without wading through numerous uninteresting duplicates.
+We therefore also introduce the idea of prioritizing mutants, using a Furthest-Point-First \cite{Gonzalez} algorithm and distance metric inspired by previous work on helping developers sort through failing test cases to find distinct faults \cite{PLDI13,mutComp}, to help static analysis tool developers find interesting patterns without wading through numerous uninteresting mutants not identified for similar reasons.
 
@@
 Finally, many mutants (though many fewer than in these categories) add {\tt break} or {\tt continue} statements.
-Again, we do not see how a tool could take advantage of this without either genuinely understanding likely-bad control flow patterns (in which case, we think it ``deserves'' to kill more mutants) or flagging so many instances of these it would have a bad mutant ratio.
+Again, we do not see how a tool could take advantage of this without either genuinely understanding likely-bad control flow patterns (in which case, we think it ``deserves'' to kill more mutants) or flagging so many legitimate uses of {\tt break} and {\tt continue} in un-mutated code that it would have a bad mutant ratio, due to producing more warnings than other tools.
 
@@
 Consider the code in Figure \ref{fig:sol424intro}, from the Solidity 0.4.24 ``Introduction to Smart Contracts''.
-The Universal Mutator tool \cite{universalmutator,regexpMut}, which has been extensively tuned for Solidity's grammar (though not to target any particular vulnerabilities), and is the only smart contract mutation tool referenced in the Solidity documentation (\url{https://solidity.readthedocs.io/en/v0.5.12/resources.html}), produces seven valid, non-redundant mutants for this trivial example code.
+The Universal Mutator tool \cite{universalmutator,regexpMut}, which has been extensively tuned for Solidity's grammar (though not to target any particular vulnerabilities), and is the only smart contract mutation tool referenced in the Solidity documentation (\url{https://solidity.readthedocs.io/en/v0.5.12/resources.html}), produces seven valid, non-redundant (by Trivial Compiler Equivalence \cite{TCE}) mutants for this trivial example code.
 Both the public version of Slither \cite{slither} and SmartCheck \cite{smartcheck} produce a small number (three and two, respectively) of low-severity, informational, warnings for this code.

### 2020-01-25 prioritizing.tex
Various fixes

@@
 
-We do not claim this is an optimal, or even tuned, metric; in the long run, it may be completely replaced.  It did
-provide useful results for a different problem, ranking mutants in an
+We do not claim this is an optimal, or even tuned, metric; in the long run, it even may be completely replaced.
+One element that will likely remain is the focus on source code, rather than bytecode; static analysis tools generally analyze source, and some even work for code that does not compile.
+The metric we use
+provided useful results for a different problem, ranking mutants in an
 effort similar to that of Groce et
@@
 considerable advantage over wading through an un-ordered list of
-mutants, and introduce the idea of using FPF for mutants, not just tests.
+mutants, and introduce the idea of using FPF for mutants, not just for tests: FPF is useful for failures in general, however discovered.

### 2020-01-25 experiments.tex
Time

@@
 Extracting the signal from Securify's noise will be difficult.
-We also note that while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick (it took about 6 days sequential compute time for Slither and 3 days for SmartCheck), Securify often required many hours to analyze a mutant, and frequently required a few days to analyze a mutant; the full analysis required over three months of compute time.
+We also note that while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick (it took about 6 days sequential compute time for Slither and 3 days for SmartCheck, i.e., about 5-15 seconds per mutant for both tools), Securify often required many hours to analyze a mutant, and frequently required a few days to analyze a mutant; the full analysis required over three months of compute time.
 

### 2020-01-25 experiments.tex
Shrink and add

@@
 \begin{figure}
-  \includegraphics[width=0.7\columnwidth]{solidity.png}
+  \includegraphics[width=0.6\columnwidth]{solidity.png}
   \caption{Mutants killed by Solidity static analysis tools.}
@@
 
+\subsubsection{Confirmation of Results}  Because our random contracts' quality might be low, we also checked our results on 30 contracts from the Solidity documentation, the \emph{Mastering Ethereum} book \cite{masteringEth}, and a handful of selected, recent, higher quality blockchain contracts.
+Slither had a mean mutation score of 0.11, vs.
+0.4 for SmartCheck and 0.01 for Securify.
+Associated mutant ratios were 0.38, 0.08, and close to zero.
+Mutant kill overlap was also similar (in fact, Figure \ref{fig:examplevenn} shows the results; Slither is A, SmartCheck is B, and Securify is C.
+
 \subsection{Java Tools}
@@
 \begin{figure}
-  \includegraphics[width=0.7\columnwidth]{java.png}
+  \includegraphics[width=0.6\columnwidth]{java.png}
   \caption{Mutants killed by Java static analysis tools.}
@@
 \begin{figure}
-  \includegraphics[width=0.7\columnwidth]{python.png}
+  \includegraphics[width=0.6\columnwidth]{python.png}
   \caption{Mutants killed by Python static analysis tools.}

### 2020-01-25 experiments.tex
Add confirmation stuff

@@
 Slither had a mean mutation score of 0.11, vs.
-0.4 for SmartCheck and 0.01 for Securify.
-Associated mutant ratios were 0.38, 0.08, and close to zero.
-Mutant kill overlap was also similar (in fact, Figure \ref{fig:examplevenn} shows the results; Slither is A, SmartCheck is B, and Securify is C.
+0.4 for SmartCheck, and 0.01 for Securify.
+Associated mutant ratios were 0.38, 0.08, and $<$ 0.0001.
+Mutant kill overlap was also similar; in fact, Figure \ref{fig:examplevenn} shows the results: Slither is A, SmartCheck is B, and Securify is C.
 

### 2020-01-25 conclusions.tex
Conc

@@
 software testing, as well \cite{groce2018verified}.
+We are also interested in combining static analysis and testing mutation analysis: it might be useful to consider a mutant killed if it is detected either by static or dynamic analysis, using static kills to lower the number of required costly testing/verification runs for mutants and the combined results as a useful guide to the quality of overall \emph{defense in depth} for high-importance code.

### 2020-01-25 conclusions.tex
Tiny fix

--- 2020-01-24 conclusions.tex
@@
 software testing, as well \cite{groce2018verified}.
+We are also interested in combining static analysis and testing mutation analysis: it might be useful to consider a mutant killed if it is detected either by static or dynamic analysis, using static kills to lower the number of required costly testing/verification runs for mutants and the combined result as a useful guide to the quality of overall \emph{defense in depth} for high-importance code.

### 2020-01-25 abstract.tex
Gustavo

@@
 In addition to making tool differences quantitatively observable without extensive manual effort, this approach offers a new way to detect and fix omissions in a static analysis tool's set of detectors.
-We present an extensive comparison of three well-known Solidity smart contract static analysis tools, and show how using an automatic prioritization of our results allowed us to add three new detectors to the best of the tools.
+We present an extensive comparison of three well-known Solidity smart contract static analysis tools, and show how using an automatic prioritization of our results allowed us to add three effective new detectors as an open source contribution to the best of the tools.
 We also evaluate popular Java and Python static analysis tools and discuss their strengths and weaknesses.

### 2020-01-25 conclusions.tex
Gustavo

@@
 We were also able to use our methods, plus a novel mutant prioritization scheme, to
-identify three useful new detectors for the Slither smart contract
-analyzer.
+identify and implement three useful new detectors for the open source Slither smart contract
+analyzer, the best-performing of the tools.
 

### 2020-01-25 experiments.tex
Conc
Gustavo

@@
 We analyzed three well-known tools for static analysis of Solidity smart contracts: Slither \cite{slither}, SmartCheck \cite{smartcheck}, and Securify \cite{Securify}.
-Slither, based on an SSA-based intermediate language (SlithIR \cite{slither}) is a tool produced by Trail of Bits.
+Slither, based on an SSA-based intermediate language (SlithIR \cite{slither}) is an open-source tool from Trail of Bits.
 SmartCheck, developed by SmartDec, translates Solidity source directly to an XML-based representation, then uses \emph{XPath} patterns to define problems.
@@
 For reentrancy, Slither performed best on two real-world large contracts, finding subtle bugs in both, SmartCheck detected the problem in one of the two, and Securify detected neither.
-For a dataset of 1,000 contracts, SmartCheck had a high false positive rate (over 70\%) but detected more actual reentrancies (209) than Slither (99) or Securify (6).
+For a set of 1,000 contracts, SmartCheck had a high false positive rate (over 70\%) but detected more actual reentrancies (209) than Slither (99) or Securify (6).
 On the other hand, Slither's low false positive rate of 11\%  makes its results possibly more useful in practice.
-It seems safe to say that
 
@@
 0.4 for SmartCheck, and 0.01 for Securify.
-Associated mutant ratios were 0.38, 0.08, and $<$ 0.0001.
+Associated mutant ratios were 0.38, 0.08, and 0.007.
 Mutant kill overlap was also similar; in fact, Figure \ref{fig:examplevenn} shows the results: Slither is A, SmartCheck is B, and Securify is C.

### 2020-01-26 introduction.tex
Shrink and add
bibfix

@@
 \begin{figure}
-  \includegraphics[width=0.7\columnwidth]{example.png}
+  \includegraphics[width=0.5\columnwidth]{example.png}
   \caption{Mutants killed by three static analysis tools.}
@@
 As with efforts to improve test suites, manually searching through all mutants can be an onerous task, especially for large-scale evaluations.
-We therefore also introduce the idea of prioritizing mutants, using a Furthest-Point-First \cite{Gonzalez} algorithm and distance metric inspired by previous work on helping developers sort through failing test cases to find distinct faults \cite{PLDI13,mutComp}, to help static analysis tool developers find interesting patterns without wading through numerous uninteresting mutants not identified for similar reasons.
+We therefore also introduce the idea of prioritizing mutants, using a Furthest-Point-First \cite{Gonzalez} algorithm and distance metric inspired by previous work on helping developers sort through failing test cases to find distinct faults \cite{PLDI13,distMut}, to help static analysis tool developers find interesting patterns without wading through numerous uninteresting mutants not identified for similar reasons.
 

### 2020-01-26 introduction.tex
fixes

@@
 
-Mutation testing \cite{demillo1978hints,budd1980theoretical} (or mutation analysis, the term we will use in this work, for reasons that will become clear) uses small syntactic changes to a program to introduce synthetic ``faults,'' under the assumption that if the original version of a program is (mostly) correct, most small changes will therefore introduce a fault.
+Mutation testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical} (or mutation analysis, the term we will use in this work, for reasons that will become clear) uses small syntactic changes to a program to introduce synthetic ``faults,'' under the assumption that if the original version of a program is (mostly) correct, most small changes will therefore introduce a fault.
 For the most part, mutation analysis has been used to evaluate test suites by computing a score (the ratio of mutants the suite detects, or ``kills'').
@@
 The tool cannot simply flag all code as having a problem or it will perform poorly at the task of \emph{distinguishing} the mutated code from non-mutated (and presumably at least \emph{more} correct) code.
-Based on verification and testing uses of mutation, it is safe to say that usually at minimum 40\%, often 50-60\%, and frequently up to 80\%+ \cite{mutKernel,groce2018verified}, of mutants are not semantically equivalent to the original code, and are actually fault-inducing \cite{TCE,impactEquiv,smith2009should}, so the task presented to a static analysis tool is the generalization of the task we ideally expect static analysis to perform:  to identify faulty code, without executing it, and, most critically, \emph{to distinguish faulty from correct code}.
+Based on verification and testing uses of mutation, it is safe to say that usually at minimum 40\%, often 50-60\%, and frequently up to 80\%+ \cite{mutKernel,groce2018verified,le2014mucheck}, of mutants are not semantically equivalent to the original code, and are actually fault-inducing and detectable by some test or property \cite{TCE,impactEquiv,smith2009should}, so the task presented to a static analysis tool is the generalization of the task we ideally expect static analysis to perform:  to identify faulty code, without executing it, and, most critically, \emph{to distinguish faulty from correct code}.
 Obviously, many faults cannot be identified statically without a complete specification, or without unreasonable analysis cost and precision, but the measure of performance here is meant to be mostly \emph{relative} to other tools applied to the same code; this is primarily a \emph{differential} approach.

### 2020-01-26 relatedwork.tex
fixes

@@
 
-Finally, the large body of work on using mutants in software testing \cite{demillo1978hints,budd1980theoretical, groce2015verified,groce2018verified,MutGoogle,ivankovic2018industrial,mutKernel} is obviously relevant, in that we basically adopt its approach, but re-define killing a mutation for a static analysis context.
+Finally, the large body of work on using mutants in software testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical, groce2015verified,groce2018verified,MutGoogle,ivankovic2018industrial,mutKernel} is obviously relevant, in that we basically adopt its approach, but re-define killing a mutation for a static analysis context.
 Methods for improving mutation analysis in testing will often (but not always, e.g.

### 2020-01-26 experiments.tex
Typo in score for smartcheck
big change, fix SpotBugs results

@@
 Slither had a mean mutation score of 0.11, vs.
-0.4 for SmartCheck, and 0.01 for Securify.
+0.04 for SmartCheck, and 0.01 for Securify.
 Associated mutant ratios were 0.38, 0.08, and 0.007.
@@
 PMD (\url{https://pmd.github.io/}) \cite{CompareJavaTools} is an extensible cross-language static code analyzer.
-FaceBook's Infer (\url{https://fbinfer.com/}) \cite{Infer} focuses on diff-related detection of serious errors (concurrency, memory safety, and information flow) and aims to scale to very large code bases.
+FaceBook's Infer (\url{https://fbinfer.com/}) \cite{Infer} focuses on diff-related detection of serious errors (concurrency, memory safety, and information flow).
 
@@
 For Java, we analyzed the top 15 projects satisfying our criteria, with a maximum of 623,355 LOC and a minimum of 3,957 LOC, and a total size of 1.8 million LOC.
-The Java compiler rejected a large number of the generated mutants, leaving us with just over 70,000 mutants to analyze, ranging from 136 to 10,016 per project.
+Because the Universal Mutator does not ``know'' Java syntax, the Java compiler rejected a large number of the generated mutants (e.g., deleting definitions or declarations), leaving us with just over 70,000 mutants to analyze, ranging from 136 to 10,016 per project.
 
@@
     \hline
-    SpotBugs & 0.47 & 0.00 & 0.37 & 0.21 & 0.7847 \\
+    SpotBugs & 28.93 & 14.00 & 0.07 & 0.07 & 0.002 \\
     \hline
-    PMD & 1.27 & 0.00 & 0.07 & 0.07 & 0.0562 \\
+    PMD & 53.73 & 32.00 & 0.07 & 0.07 & 0.001 \\
     \hline
-    Infer & 0.07 & 0.00 & 0.00 & 0.00 &  0.0002 \\
+    Infer & 11.60 & 3.00 & 0.00 & 0.00 &  0.000 \\
     \hline
@@
     \hline
-    SpotBugs & 3 & 0.41 & 0.17 \\
+    SpotBugs & 3 & 0.05 & 0.06 \\
     \hline
@@
 
-In terms of {\bf RQ1}, the raw results are clear:  SpotBugs finds vastly more uniquely killed mutants than the other tools, but PMD contributes a large number of its own unique detections, and clearly offers some independent utility.
+In terms of {\bf RQ1}, the raw kills results suggest there is considereable value in running both SpotBugs and PMD.
+Both produce a large number of unique detections, though PMD produces about 20\% more than SpotBugs.
 Infer on the other hand, is only able to detect two mutants, but these are unique.
@@
 
-{\bf RQ2} is also answered in the affirmative.
-While the raw kills for SpotBugs are impressive, it also produced fewer findings for clean code, giving it an excellent mutant ratio, making its dominance even more clear.
-PMD's performance was acceptable, and better than any of the Solidity tools (note that comparing ratios across languages is potentially invalid, but not obviously so).
-Infer is confirmed as nearly useless, again.
+{\bf RQ2} is also answered in the affirmative.  While the raw kills for SpotBugs are not as good as for PMD, it produced fewer findings, giving it a mutant ratio approximately twice that of PMD.
+
+We note that SpotBugs crashed for many more Java programs than PMD and Infer (neither crashed for any original file in our experiments).
+SpotBugs ``failed to detect'' 23,000 of the mutants because it did not process the un-mutated file for 383 files, just over 23\% of the 1,664 total files.
+Removing files for which SpotBugs failed, however, did not dramatically change results; SpotBugs mutant ratio rose to 0.003, but PMD's rose to 0.002.
 
@@
   {\scriptsize
-  \begin{itemize}
-  \item \url{https://stackoverflow.com/questions/4297014/what-are-the-differences-between-pmd-and-findbugs}
-  \item \url{https://www.sw-engineering-candies.com/blog-1/comparison-of-findbugs-pmd-and-checkstyle}
-  \item \url{https://www.reddit.com/r/java/comments/3i7w6n/checkstyle_vs_pmd_vs_findbugs_for_dummies_why/}
-  \end{itemize}
+    % \begin{itemize}
+    \raggedright    
+  \url{https://stackoverflow.com/questions/4297014/what-are-the-differences-between-pmd-and-findbugs} \\
+  \url{https://www.sw-engineering-candies.com/blog-1/comparison-of-findbugs-pmd-and-checkstyle} \\
+  \url{https://www.reddit.com/r/java/comments/3i7w6n/checkstyle_vs_pmd_vs_findbugs_for_dummies_why/} \\
+  %\end{itemize}
   }
@@
 However, the user postings listed in Figure \ref{fig:blog}, plus personal communications with security analysts who use these tools (citation removed for now, due to potential to compromise blinding) supported some basic conclusions.
-SpotBugs/FindBugs is indeed the best tool for finding bugs; PMD focuses more on stylistic issues and has a weaker semantic model.
-Mutants are much more likely to be bugs than stylistic issues (they alter semantics, not code style, in almost all cases: deletion or constant/operator changes obviously have no style impact).
+SpotBugs/FindBugs is supposedly the best tool for finding bugs; PMD focuses more on stylistic issues and has a weaker semantic model.
+Running both is definitely recommended, as neither is extremely effective.
 Infer is closer to a model checker focusing on resource leaks than a truly general-purpose tool, arguably.
 In fact, we suspect Infer would perform much better if we crafted complex mutation operators targeting some important subtle Java bugs; it is still fair to say that Infer is probably not a good \emph{general-purpose} static analysis tool for Java.
+Note that we did not use Infer's experimental, non-standard, detectors, however.
 
-For {\bf RQ4} there were very few clean projects, but the mutation scores for Infer were unchanged, and for SpotBugs the mean mutation score improved and median mutation score was worse.
-Again, we see no evidence that non-clean code is a source of significant degradation on mutation detection.
+For {\bf RQ4} there were very few clean projects, but the mutation scores for Infer were unchanged, and for SpotBugs they were worse.
+We see no evidence that non-clean code is a source of degradation in mutation detection.
 
@@
 Pyflakes' mutation score is lower than unique kills might suggest because it has almost no overlap with the other two tools.
-It is doing something quite different than they are, apparently.
+It is doing something different.
 
@@
   {\scriptsize
-  \begin{itemize}
-  \item \url{https://stackoverflow.com/questions/1428872/pylint-pychecker-or-pyflakes}
-  \item \url{https://www.reddit.com/r/Python/comments/ii3gm/experience_with_pylint_pychecker_pyflakes/}
-  \item \url{https://www.slant.co/versus/12630/12631/~pylint_vs_pyflakes}
-  \item \url{https://news.ycombinator.com/item?id=12748885}
-  \item \url{https://doughellmann.com/blog/2008/03/01/static-code-analizers-for-python/}
-  \end{itemize}
+    % \begin{itemize}
+    \raggedright
+  \url{https://stackoverflow.com/questions/1428872/pylint-pychecker-or-pyflakes}\\
+  \url{https://www.reddit.com/r/Python/comments/ii3gm/experience_with_pylint_pychecker_pyflakes/}\\
+  \url{https://www.slant.co/versus/12630/12631/~pylint_vs_pyflakes}\\
+  \url{https://news.ycombinator.com/item?id=12748885}\\
+  \url{https://doughellmann.com/blog/2008/03/01/static-code-analizers-for-python/}\\
+  %\end{itemize}
   }

### 2020-01-26 analyzejava.py

@@ -31,12 +31,31 @@ projects = set([])
 projectFiles = {}
 
 Mutant = 0
-    
+
+spotbugsFailed = 0
+PMDFailed = 0
+InferFailed = 0
+
+totalFiles = set([])
+failFiles = set([])
+failProjects = set([])
+
 with open("Java.csv") as javaResults:
     reader = csv.DictReader(javaResults)
     for row in reader:
+        #if row["Spotbug"] == "-1":
+        #    continue
         Mutant += 1
         fname = row["Project"]
+        totalFiles.add(row["Project"] + ":" + row["File"])
+        if row["Spotbug"] == "-1":
+            spotbugsFailed += 1
+            failFiles.add(row["Project"] + ":" + row["File"])
+            failProjects.add(row["Project"])
+        if row["PMD"] == "-1":
+            PMDFailed += 1
+        if row["Infer"] == "-1":
+            InferFailed += 1
         projects.add(row["Project"])
         if fname not in files:
             projectFiles[fname] = []
@@ -63,13 +82,13 @@ with open("Java.csv") as javaResults:
             x = int(row["PMD"])
         except:
             print("NO ISSUES FOR PMD")            
-        if row["spotbugs_killed"] == "1":
+        if row["spotbugs_killed"] == "1" and row["Spotbug"] != "-1":
             files[fname]["spotbugs_kills"] += 1
             spotbugs_killed += 1
             spotbugs_file.write(row["Mutant"] + "\n")
             if row["PMD_killed"] == "0":
                 spotbugs_not_PMD += 1
-        if row["PMD_killed"] == "1":
+        if row["PMD_killed"] == "1" and row["PMD"] != "-1":
             files[fname]["PMD_kills"] += 1
             PMD_killed += 1
             PMD_file.write(row["Mutant"] + "\n")
@@ -78,7 +97,6 @@ with open("Java.csv") as javaResults:
         if row["Infer_killed"] == "1":
             files[fname]["Infer_kills"] += 1            
             Infer_file.write(row["Mutant"] + "\n")            
-            print(row)
             Infer_killed += 1
         try:
             spotbugs_warnings += int(row["Spotbug"])
@@ -134,3 +152,17 @@ with open("Java.csv") as javaResults:
 
 print(projects, len(projects))
 print(Mutant-1)
+
+for f in files:
[…]

### 2020-01-26 conclusions.tex
Fixups

@@
 software testing, as well \cite{groce2018verified}.
-We are also interested in combining static analysis and testing mutation analysis: it might be useful to consider a mutant killed if it is detected either by static or dynamic analysis, using static kills to lower the number of required costly testing/verification runs for mutants and the combined result as a useful guide to the quality of overall \emph{defense in depth} for high-importance code.
+We are also interested in combining static and traditional mutation analysis: it might be useful to consider a mutant killed if it is detected either by static or dynamic analysis, using static kills to lower the number of required costly testing/verification runs for mutants, and the combined result as a useful guide to the quality of overall \emph{defense in depth} \cite{antunes2012defending,FC20} for high-importance code.

### 2020-01-26 experiments.tex
Fixups

@@
 For Java, we analyzed the top 15 projects satisfying our criteria, with a maximum of 623,355 LOC and a minimum of 3,957 LOC, and a total size of 1.8 million LOC.
-Because the Universal Mutator does not ``know'' Java syntax, the Java compiler rejected a large number of the generated mutants (e.g., deleting definitions or declarations), leaving us with just over 70,000 mutants to analyze, ranging from 136 to 10,016 per project.
+Because the Universal Mutator does not ``know'' Java syntax, and Java is very verbose, the Java compiler rejected a large number of the generated mutants (e.g., deleting declarations), leaving just over 70,000 mutants, ranging from 136 to 10,016 per project.
 

### 2020-01-26 introduction.tex
Fixups

@@
 Using large numbers of larger, more realistic contracts makes it possible to extract the same kind of information, on a much larger scale.
-Prioritization of mutants is not very useful here (ranking three mutant saves little effort), so we will show the utility of that approach in our full Solidity results.
+Prioritization of mutants is not very useful here (ranking three mutants saves little effort), so we will show the utility of that idea in our full Solidity results.
 

### 2020-01-26 experiments.tex
Little more on mutant ranking

@@
   \end{tabular}
-  \caption{Solidity tools: number of findings and mutation scores over all contracts.}
+  \caption{Solidity tool results over all contracts.}
   \label{tab:scoresolidity}
@@
   \end{tabular}
-  \caption{Solidity tools: clean contract counts and mutation scores.}
+  \caption{Solidity tool clean contract results.}
   \label{tab:cleansolidity}
@@
 Our three submitted detectors were not present as private detectors, and only one (the type-based tautology detector) had even been identified, via a GitHub issue, as a potential improvement (and only in the private version of Slither).
+Combining statement deletion mutants with other mutants only moved the mutants we used down to 3rd, 11th, and 14th positions.
+By default we rank statement deletions separately, since such mutants are usually easier to understand and either note as important or dismiss, and in testing (but not static analysis) they are likely to be the most critical faults not detected.
 
@@
   \end{tabular}
-  \caption{Java tools: number of findings and mutation scores over all projects.}
+  \caption{Java tool results over all projects.}
   \label{tab:scorejava}
@@
   \end{tabular}
-  \caption{Java tools: clean project counts and mutation scores.  No projects were clean for all tools, due to PMD producing findings for all projects.}
+  \caption{Java tool clean project results.}
   \label{tab:cleanjava}
@@
   \end{tabular}
-  \caption{Python tools: number of findings and mutation scores over all projects.}
+  \caption{Python tool results over all projects.}
   \label{tab:scorepython}
@@
   \end{tabular}
-  \caption{Python tools: clean project counts and mutation scores.  No projects were clean for all tools, due to Pylint producing findings for all projects.}
+  \caption{Python tools clean project results.}
   \label{tab:cleanpython}

### 2020-01-26 experiments.tex
Mention real pyfakefs bugs

@@
 We note that SpotBugs crashed for many more Java programs than PMD and Infer (neither crashed for any original file in our experiments).
-SpotBugs ``failed to detect'' 23,000 of the mutants because it did not process the un-mutated file for 383 files, just over 23\% of the 1,664 total files.
-Removing files for which SpotBugs failed, however, did not dramatically change results; SpotBugs mutant ratio rose to 0.003, but PMD's rose to 0.002.
+SpotBugs ``failed to detect'' 23,000 of the mutants because it did not process the un-mutated file for 383 files, over 12 of the 15 projects---just over 23\% of the 1,664 total files.
+Removing files for which SpotBugs failed, however, did not dramatically change results; SpotBugs' mean mutation score rose to 0.09, and mutant ratio rose to 0.003, but PMD's mean score rose to 0.10, and mutant ratio to 0.002.
 

### 2020-01-26 introduction.tex
Mention real pyfakefs bugs

@@
 proposed examining individual mutants that survive a formal verification or automated test generation process to detect and correct weaknesses in a specification or test generator.
-The approach was able to expose bugs in a heavily-tested module of the Linux kernel \cite{mutKernel} and improve a heavily used test generator for the {\tt pyfakefs} file system.
+The approach was able to expose bugs in a heavily-tested module of the Linux kernel \cite{mutKernel} and improve a heavily used test generator for the {\tt pyfakefs} file system, allowing it to detect multiple new bugs.
 Recently, mutation analysis has been adopted in industrial settings, though not for actual examination of all surviving mutants \cite{MutGoogle,ivankovic2018industrial}, a practice that is hard to scale to large code bodies.

### 2020-01-26 experiments.tex
Updates from Overleaf

@@
 For Java, we analyzed the top 15 projects satisfying our criteria, with a maximum of 623,355 LOC and a minimum of 3,957 LOC, and a total size of 1.8 million LOC.
-Because the Universal Mutator does not ``know'' Java syntax, and Java is very verbose, the Java compiler rejected a large number of the generated mutants (e.g., deleting declarations), leaving just over 70,000 mutants, ranging from 136 to 10,016 per project.
+Because the Universal Mutator does not ``know'' Java syntax, and Java is very verbose, the Java compiler rejected a large number of the generated mutants (e.g., deleting declarations).
+We still, due to the huge size of the source files and thus number of mutants (and time to compile full projects), restricted our analysis to files where Universal Mutator's implementation of TCE \cite{TCE} for Java was useful, i.e.
+files that could be compiled and the bytecode compared without processing the full project, leaving us with just over 70,000 mutants, ranging from 136 to 10,016 per project.
 
@@
 For {\bf RQ4} there were no clean projects for the known-to-be-aggressive Pylint.
-Performance for the other two tools was worse on projects for which they were clean, again failing to support the possibility that our definition of killing a mutant under-reports detection for non-clean code.
+Performance for the other two tools was worse on projects for which they were clean, again failing to support the possibility that our definition of killing a mutant is too restrictive.
 

### 2020-01-26 introduction.tex

@@
 This is strong evidence that A is simply better all-around than B; it likely has a lower false positive rate \emph{and} a lower false negative rate.
-While it is not proof of this claim, it is hard to construct another plausible explanation for reporting \emph{fewer} findings on un-mutated contracts while still detecting \emph{more} mutants.
+While it is not proof of this claim, it is hard to construct another plausible explanation for reporting \emph{fewer} findings on un-mutated code while still detecting \emph{more} mutants.
 Other than having better precision and recall, how else could a tool effectively distinguish mutated from un-mutated code?

### 2020-01-26 relatedwork.tex

@@
 in contrast to their introduction of checks that may be guaranteed to pass, so this is probably not very helpful).
-Most importantly, however, their approach only applies to tools that check assertions, rather than more general static analysis tools that may only identify bad code patterns, not compute state reachability (e.g, most of the widely used tools).
+Most importantly, however, their approach only applies to tools that check assertions, rather than more general, popular static analysis tools that only identify bad code patterns.
 
 Finally, the large body of work on using mutants in software testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical, groce2015verified,groce2018verified,MutGoogle,ivankovic2018industrial,mutKernel} is obviously relevant, in that we basically adopt its approach, but re-define killing a mutation for a static analysis context.
-Methods for improving mutation analysis in testing will often (but not always, e.g.
-those based on coverage or other dynamic information) transfer to our context as well.

### 2020-01-27 introduction.tex
Comma murder

@@
 
-Users of static analysis tools often wonder which of multiple tools available for a programming language are most effective, and, when using more than one tool is an option (e.g.
-with free tools), how much tools overlap in their results.
+Users of static analysis tools often wonder which of multiple tools available for a programming language are most effective, and, when using more than one tool is an option, which is often the case, how much tools overlap in their results.
 Given the human effort required to read static analysis results, the latter can be an important question.
-If two tools find substantially different (non-false-positive) bugs, it is wise to use both if finding all bugs is important.
+If two tools find substantially different bugs, it is wise to use both tools.
 On the other hand, if two tools are very similar in what they detect, the effort of wading through duplicate results may not be a good use of time.
-Developers of static analysis tools also want to be able to compare their tools to others targeting the same domain, in order to see what detection methods (or tweaks to precision/soundness trade-offs) they might want to imitate.
+Developers of static analysis tools also want to be able to compare their tools to others targeting the same domain, in order to see what detection methods or tweaks to precision/soundness trade-offs they might want to imitate.
 Unfortunately, comparing static analysis tools is hard, and would seem to require vast manual effort to inspect findings and determine ground truth on a scale that would provide statistical confidence.
@@
 
-Mutation testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical} (or mutation analysis, the term we will use in this work, for reasons that will become clear) uses small syntactic changes to a program to introduce synthetic ``faults,'' under the assumption that if the original version of a program is (mostly) correct, most small changes will therefore introduce a fault.
-For the most part, mutation analysis has been used to evaluate test suites by computing a score (the ratio of mutants the suite detects, or ``kills'').
+Mutation testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical} (or mutation analysis, the term we will use in this work, for reasons that will become clear) uses small syntactic changes to a program to introduce synthetic ``faults,'' under the assumption that if the original version of a program is mostly correct, most small changes will therefore introduce a fault.
+For the most part, mutation analysis has been used to evaluate test suites by computing a mutation score, the fraction of mutants the suite detects, or ``kills''.
 Most such use has been in research efforts, rather than practical testing efforts, though there has been sporadic use by interested developers.
@@
 
-We can say that a static analysis tool kills a mutant when the number of (non-informational or optimization related) warnings or errors produced with respect to the code \emph{increases} for the mutated version, compared to the original code.
-This difference is most informative and easily interpreted when the original code produces no warnings or errors (it is ``clean''); for non-clean code, a tool conceivably could detect the mutant, but only change a previously generated warning, not add an additional warning, leading to an underestimate of effectiveness.
-However, even for non-clean code, most detected mutants will produce a new warning.
+We can say that a static analysis tool kills a mutant when the number of non-informational or optimization related, that is fault-related, warnings or errors produced with respect to the code \emph{increases} for the mutated version, compared to the original code.
+This difference may be most informative and easily interpreted when the original code produces no warnings or errors; we call such code ``clean''.
+For non-clean code, a tool conceivably could detect the mutant, but only change a previously generated warning, not add an additional warning, leading to an underestimate of effectiveness.
+However, even for non-clean code, most detected mutants should produce a new warning.
 
@@
 First, this is a measure that does not reward a tool that produces too many false positives.
-The tool cannot simply flag all code as having a problem or it will perform poorly at the task of \emph{distinguishing} the mutated code from non-mutated (and presumably at least \emph{more} correct) code.
+The tool cannot simply flag all code as having a problem or it will perform poorly at the task of \emph{distinguishing} the mutated code from non-mutated, and presumably at least \emph{more} correct, code.
 Based on verification and testing uses of mutation, it is safe to say that usually at minimum 40\%, often 50-60\%, and frequently up to 80\%+ \cite{mutKernel,groce2018verified,le2014mucheck}, of mutants are not semantically equivalent to the original code, and are actually fault-inducing and detectable by some test or property \cite{TCE,impactEquiv,smith2009should}, so the task presented to a static analysis tool is the generalization of the task we ideally expect static analysis to perform:  to identify faulty code, without executing it, and, most critically, \emph{to distinguish faulty from correct code}.
@@
 It is not clear that any other fully automatic method is competitively meaningful; it is possible that methods based on code changes from version control provide some of the same benefits, but these require classification of changes into bug-fixes and non-bug-fixes, and of course require version control history.
-Also, history-based methods will be biased towards precisely those faults humans (or tools) were able to detect and fix.
+Also, history-based methods will be biased towards precisely those faults humans or tools already in use were able to detect and fix.
 
@@
 Since we are comparing over the same program or set of programs, we simply divide the mean \emph{mutation score} ($\frac{|\mathit{killed}||}{|\mathit{mutants}|}$, the ratio of killed mutants to all mutants) by the mean number of findings.
-This \emph {mutant ratio} tells us about the ability of a tool to produce findings \emph{for mutants}, relative to its tendency to produce findings in general (per line of code, per source file, etc.).
-If a tool has a tendency to produce large numbers of findings (compared to other tools), and this is paired with a tendency to detect most mutants, then the tool will not be penalized for producing many findings.
-Assuming that real faults are relatively rare in the original, un-mutated code, the best result (and best ratio score) will be for a tool that produces comparatively few findings for un-mutated code, but detects a larger portion of mutants than other tools; the worst result will be a tool that produces lots of findings, but detects few mutants.
+This \emph {mutant ratio} tells us about the ability of a tool to produce findings \emph{for mutants}, relative to its tendency to produce findings in general, (per line of code, per source file, etc.).
+If a tool has a tendency to produce large numbers of findings compared to other tools, and this is paired with a tendency to detect most mutants, then the tool will not be penalized for producing many findings.
+Assuming that real faults are relatively rare in the original, un-mutated code, the best result and best mutant ratio will be for a tool that produces comparatively few findings for un-mutated code, but detects a larger portion of mutants than other tools; the worst result will be a tool that produces lots of findings, but detects few mutants.
 We will actually see some examples of the worst case in our results for real tools.
@@
 
-Finally, even when tools have similar quantitative results, including similar ratios, examining individual mutants killed by one tool but not by another allows us to understand strengths and weaknesses of the tools, in a context that makes comprehending the cause of the detection (or lack of it) easy: the difference between the un-mutated code and mutated code will always be small and relatively simple.
+Finally, even when tools have similar quantitative results, including similar ratios, examining individual mutants killed by one tool but not by another allows us to understand strengths and weaknesses of the tools, in a context that makes comprehending the cause of the detection, or lack of it, easy: the difference between the un-mutated code and mutated code will always be small and relatively simple.
 Moreover, simply looking at how much two tools agree on mutants can answer the question of a user of static analysis tools: given that I am using tool A, would adding tool B be likely to add enough new, interesting results to make it worth my time to examine its output?
-When (rarely) one tool subsumes another in terms of mutants, it can be very clear that the tool whose killed mutants are all killed by another tool is likely strictly inferior.
+When one tool subsumes another in terms of mutants, it can be very clear that the tool whose killed mutants are all killed by another tool is likely strictly inferior.
 Interested users, e.g.
@@
 B does not do as well as A, but it is the only tool that detects a large number of mutants, and most mutants it detects are unique to it.
-Finally, Tool C may not be worth running; recall that it produces a similar number of findings to A and B on the un-mutated code, so it is notably bad at detecting faults (at least ones that look like mutants).
+Finally, Tool C \emph{may} not be worth running; recall that it produces a similar number of findings to A and B on the un-mutated code, so it is notably bad at detecting faults, at least ones that look like mutants.
 It might be a good idea to just look at the 27 mutants detected by C alone:  if they represent an important class of potential problems (perhaps C specialized in detecting potentially non-terminating loops), then C might be useful, but if the first few mutants inspected are false positives, then C is likely not useful.
 
-Comparing mutant results also leads to the idea of \emph{improving} static analysis tools by examining mutants detected by another tool (thus known to be in-principle detectable) but not by the tool to be improved.
+Comparing mutant results also leads to the idea of \emph{improving} static analysis tools by examining mutants detected by another tool, and thus known to be in-principle detectable, but not detected by the tool to be improved.
 Of course, any faults in code, not just mutants, could serve this purpose.
@@
 A general objection to our approach is that mutants may differ substantially from ``real'' faults, in some way.
-This is certainly true, in a sense \cite{GopinathMutants}, but for static analysis purposes does not matter.
+This is certainly true, in a sense \cite{GopinathMutants}, but for static analysis purposes we believe it does not matter.
 The real risk is that some mutation operators align with patterns a particular tool identifies, biasing the evaluation in favor of that tool.
@@
 Changing arithmetic and comparison operators and numeric constants (incrementing, decrementing, or changing to 0 or 1) account for most of the non-statement-deletion mutants.
-We cannot see how any tool could have a (useful) detection rule that looks for these kinds of changes (a tool obviously would not flag all instances of 0 and 1 constants, or of addition, and remain a useful tool).
-Finally, many mutants (though many fewer than in these categories) add {\tt break} or {\tt continue} statements.
-Again, we do not see how a tool could take advantage of this without either genuinely understanding likely-bad control flow patterns (in which case, we think it ``deserves'' to kill more mutants) or flagging so many legitimate uses of {\tt break} and {\tt continue} in un-mutated code that it would have a bad mutant ratio, due to producing more warnings than other tools.
+We cannot see how any tool could have a useful detection rule that looks for these kinds of changes: a tool obviously would not flag all instances of 0 and 1 constants, or of addition, and remain a useful tool.
+Finally, many mutants, though many fewer than in these categories, add {\tt break} or {\tt continue} statements.
+Again, we do not see how a tool could take advantage of this without either genuinely understanding likely-bad control flow patterns, in which case, we think it ``deserves'' to kill more mutants, or flagging so many legitimate uses of {\tt break} and {\tt continue} in un-mutated code that it would have a bad mutant ratio, due to producing more warnings than other tools.
 
@@
 SmartCheck, on the other hand, detects removing the {\tt return} in {\tt get} or replacing it with a {\tt revert}, or removing the {\tt public} visibility modifier for {\tt get} \footnote{Slither's ``missing return'' detector is only available in the private version of slither, or through the {\tt crytic} service provided by Trail of Bits.}.
-If we restrict our analysis to findings with a severity greater than informational, SmartCheck detects no mutants of the contract, while Slither still reports that some mutants \emph{allow an arbitrary caller to cause the contract to self destruct}.
+If we restrict our analysis to findings with a severity greater than informational, SmartCheck detects no mutants of the contract, while Slither still reports that some mutants \emph{allow an arbitrary caller to cause the contract to destroy itself}.
 This simple example shows how our approach works on a small scale.
@@
 The Solidity, Java, and Python case studies also serve to answer a set of research questions investigating the utility of our approach, and provide basic evidence that it provides actionable, non-obvious information that is otherwise difficult to produce.
-While there are certainly limitations to using differential mutation analysis to compare (and extend) static analysis tools, the method scales to basing comparisons on large numbers of real software source files, but has some of the advantages of humans establishing ground truth for tool findings.
+While there are certainly limitations to using differential mutation analysis to compare and extend static analysis tools, the method scales to basing comparisons on large numbers of real software source files and very large numbers of faults, but still has some of the advantages of humans establishing ground truth for tool findings.

### 2020-01-27 experiments.tex
Cites etc

@@
 
-For {\bf RQ3}, to our knowledge there is no recent academic comparison of these tools; the most relevant study dates from 2004 \cite{CompareJavaTools}, used FindBugs, not SpotBugs, and reached no strong conclusions with respect to FindBugs vs.
-PMD.
-However, the user postings listed in Figure \ref{fig:blog}, plus personal communications with security analysts who use these tools (citation removed for now, due to potential to compromise blinding) supported some basic conclusions.
-SpotBugs/FindBugs is supposedly the best tool for finding bugs; PMD focuses more on stylistic issues and has a weaker semantic model.
+For {\bf RQ3}, to our knowledge there is no academic comparison of all three tools; one study dates from 2004 \cite{CompareJavaTools}, used FindBugs, not SpotBugs, and reached no strong conclusions with respect to FindBugs vs.
+PMD; a more recent study found that SpotBugs outperformed Infer for Defects4J \cite{just2014defects4j} bugs \cite{AllBugs}, but did not compare to PMD.
+However, the user postings listed in Figure \ref{fig:blog}, plus personal communications with security analysts who use these tools \cite{personalJava} supported some basic conclusions.
+SpotBugs is perhaps the best tool for finding bugs; PMD focuses more on stylistic issues and has a weaker semantic model.
 Running both is definitely recommended, as neither is extremely effective.

### 2020-01-27 introduction.tex
Cites etc

@@
 Given the human effort required to read static analysis results, the latter can be an important question.
-If two tools find substantially different bugs, it is wise to use both tools.
-On the other hand, if two tools are very similar in what they detect, the effort of wading through duplicate results may not be a good use of time.
+If two tools find substantially different bugs, it is important to use both tools \cite{AllBugs}.
+On the other hand, if two tools are very similar in what they detect, the effort to configure and use both, and especially of wading through duplicate results, may not be a good use of time.
 Developers of static analysis tools also want to be able to compare their tools to others targeting the same domain, in order to see what detection methods or tweaks to precision/soundness trade-offs they might want to imitate.
@@
 
-Differential testing \cite{Differential,ICSEDiff} is a popular approach to comparing multiple software systems offering similar functionality, but the wide divergence of possible trade-offs, analysis focuses, and the prevalence of false positives in almost all analysis results makes na\"ive differential testing not applicable to static analysis tools \cite{regehrRandom}.
+Differential testing \cite{Differential,ICSEDiff,csmith} is a popular approach to comparing multiple software systems offering similar functionality, but the wide divergence of possible trade-offs, analysis focuses, and the prevalence of false positives in almost all analysis results makes na\"ive differential testing not applicable to static analysis tools \cite{regehrRandom}.
 

### 2020-01-27 experiments.tex
Ok

@@
 Can we confirm results from other evaluation methods, or informal opinion?
-\item {\bf RQ4:}  Do mutation scores for clean programs differ from those for programs for which tools report findings before mutation?
+\item {\bf RQ4:}  Are mutation scores for clean programs better than those for programs for which tools report findings before the program is mutated?
 \item {\bf RQ5:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
- different tools are good at/bad at?  Can we extend this information to improve tools by addressing identified weaknesses?
-\item {\bf RQ6:}  How easy is it identify low-hanging fruit for improving tools without prioritization to remove conceptually redundant mutants?
+ different tools are good at/bad at, and extend this information to improve tools by addressing identified weaknesses?
+\item {\bf RQ6:}  How easy is it identify low-hanging fruit for improving tools \emph{without} prioritization to remove conceptually redundant mutants?
   \end{itemize}
 
-In particular, we consider {\bf RQ2} to be of critical importance; if the mutant ratios for tools differ, then this is clear evidence that our hypothesis that the tendency of mutants to be faults, and to expect that mutated code will, by a sound and precise tool, be flagged as problematic more often than non-mutated code, holds.
+In particular, we consider {\bf RQ2} to be of critical importance; if the mutant ratios for tools differ, then this is clear evidence that our hypothesis that the tendency of mutants to be faults, and to expect that mutated code will, by a more precise and accurate tool, be flagged as problematic more often than non-mutated code, holds.
 This expectation that (some subset of the) mutants can serve as proxies for real, detectable faults is the core concept of our approach.

### 2020-01-27 introduction.tex
Ok

@@
 
-We can say that a static analysis tool kills a mutant when the number of non-informational or optimization related, that is fault-related, warnings or errors produced with respect to the code \emph{increases} for the mutated version, compared to the original code.
-This difference may be most informative and easily interpreted when the original code produces no warnings or errors; we call such code ``clean''.
-For non-clean code, a tool conceivably could detect the mutant, but only change a previously generated warning, not add an additional warning, leading to an underestimate of effectiveness.
-However, even for non-clean code, most detected mutants should produce a new warning.
+We can say that a static analysis tool kills a mutant when the number of warnings or errors, which we call \emph{findings}, that are not informational or optimization related, i.e., that are \emph{fault-related}, produced with respect to the code \emph{increases} for the mutated version, compared to the original code.
+That is, when the tool finds more bugs for the mutated code.
+This difference may be most easily interpreted when the original code produces no findings; we call such code \emph{clean}.
+For non-clean code, a tool conceivably could detect the mutant, but only change a previously generated finding, not add an additional finding, leading to an underestimate of effectiveness.
+However, even for non-clean code, \emph{most detected mutants should produce a new warning}.
 
@@
 Also, history-based methods will be biased towards precisely those faults humans or tools already in use were able to detect and fix.
+Rather than the hundreds \cite{just2014defects4j} or at most few thousand of faults \cite{BugSwarm} in benchmark defect sets, our approach enables the use of \emph{many tens of thousands} of hypothetical faults.
 

### 2020-01-27 prioritizing.tex
Ok

@@
 $$0.1 \times r(\mathit{mutant}_1, \mathit{mutant}_2) +$$
-$$0.5 \times \mathit{different\_files} +$$
-$$max(0.25, \frac{\mathit{line\_distance}(\mathit{mutant}_1, \mathit{mutant}_2)}{10})$$
+$$0.5 \times \mathit{not\_same\_file} +$$
+$$max(0.25, \frac{\mathit{line\_dist}(\mathit{mutant}_1, \mathit{mutant}_2)}{10})$$
 
 \noindent Where $r$ is a Levenshtein-based string similarity ratio,
-and $\mathit{line\_distance}$ is the distance in a source file between
-two locations, in lines (and zero if the locations are in different
-source files).
+$\mathit{line\_dist}$ is the distance in a source file between
+two locations, in lines (zero if the locations are in different
+files), and $\mathit{not\_same\_file}$ is 0/1.
 

### 2020-01-27 introduction.tex
Second bar removed

@@
 We can quantitatively express the relationship between detecting mutants and findings for the original program.
-Since we are comparing over the same program or set of programs, we simply divide the mean \emph{mutation score} ($\frac{|\mathit{killed}||}{|\mathit{mutants}|}$, the ratio of killed mutants to all mutants) by the mean number of findings.
+Since we are comparing over the same program or set of programs, we simply divide the mean \emph{mutation score} ($\frac{|\mathit{killed}|}{|\mathit{mutants}|}$, the ratio of killed mutants to all mutants) by the mean number of findings.
 This \emph {mutant ratio} tells us about the ability of a tool to produce findings \emph{for mutants}, relative to its tendency to produce findings in general, (per line of code, per source file, etc.).

### 2020-03-24 response.txt
issta

We would like to thank the anonymous reviewers for their helpful comments.

Below are responses to questions and clarifications.

R1:

Most importantly: https://github.com/mutantsforstaticanalysis/data now provides all the raw data used in the paper.  We hope this helps with qualitiative review.  We also, below, discuss briefly the issue of whether the mutants detected by any tools were of actual interest, which is indeed a core issue.  We do not examine mutants detected by no tools; due to the differential nature of the approach, we consider these potentially noise/uninteresting/not plausibly statically detected, but can simply ignore them.  Only when at least one tool detects a mutant do we care.  We note that the set of mutation operators can be easily extracted by examining the rule files in the open source repository for universalmutator.

Yes, we actually used "number" of warnings, as a conservative measure of detection.  If a mutant simply causes a tool to change the message it reports, then we do not consider that a kill.  Similarly, if it loses one warning and produces another (which is in a sense the same thing, from a user's perspective), then it is not a kill.  The core reason we don't consider the location/nature of the warning change is that in some cases a mutant causes a warning about code far from the mutant location.  Therefore, we strictly require that a tool produce _more warnings_ when code is mutated, in order to be counted as detecting the mutant.  We experimented with other definitions on a small set of smart contracts, and this was much better than other approaches, in terms of the results making sense.  It is possible (though we didn't see this in practice with smart contracts) that it undercounts cases where 1) the original code produces warning A and 2) the mutated code "fixes" problem A and causes detection of a different problem, B.  If the mutant doesn't change the issue, just changes the exact nature of the problem, then we argue the mutant didn't introduce a fault, it just changed some aspect of an existing fault, which is not what we are interested in.  If we want to compare tools over existing faults, that is already doable.

Good point, the source/bytecode choice is language dependent.  As to why we chose slither, see below.

R2:

We chose PMD primarily because we started by focusing on [43] as a comparison paper, rather than [28], due to the highly-focused scope of [28] (only Defects4J bugs, thus many many fewer relevant warnings/aspects of code considered, and any biases in producing the Defects4J data set).  Adding Error Prone as well would be a good idea.  In a way, [43] is a closer comparison to our work:  larger scope of interest, if higher risk of invalid results over that larger scope.

"Importing" rules from one tool to another can be very helpful for actual users.  If one tool uses a better underlying analysis engine (e.g., better intermediate language, better alias/data-flow/etc. approach), then adding a rule from another tool can result in much better _actual_ detection of bugs, or fewer false positives.  Slither's IR and analysis are probably more scalable than Securify and more precise than SmartCheck, so even just using the same rule gives everyone better analysis results.  Also, the rules are generalizations that expand on the rules in SmartCheck/Securify: the mutation approach makes generalizing a specific rule somewhat easier, we think, due to providing multiple examples, not just a high-level concept.

We used universalmutator for two reasons:  primarily because PIT is a bytecode-level mutation tool only, PMD requires source code, and we believe users want to inspect mutants for this application (and we needed to manually do some validation that the differences were meaningful).  Second because we wanted to use a consistent set of operators and mutation approach, and universalmutator was the best tool for the other languages.

"better" in RQ4 means that we hypothesized that tools would produce _higher_ mutation scores (kill more mutants) for programs for which there were no findings in the original code.

Securify is a static analysis tool, but it "symbolically analyzes the contract’s dependency graph to extract
precise semantic information from the code" before checking compliance/violation patterns.  The first step, unfortunately, like many symbolic approaches, can sometimes scale abominably on real code.

Our reasoning on the mutant ratio for Securify is that Securify does not notably produce warnings for mutants (which are probably bugs in some sense, even if not statically detectable) compared to the vast number of warnings it issues in general.  Other research supports the claim that Securify is prone to false positives over all code, and its behavior on mutants is not appreciably different (see the 20x or more higher ratios for the other tools.

R3:

Yes, we had to understand each tool's types of messages (for Pylint this was quite a painful process) and filter out ones that are essentially coding style rather than "this could be a bug!" issues, to the extent possible.  We made considerably effort to ensure these choices were fair to the tools.  Some tools made it easy by simply providing a setting to turn off "informational and style" issues, but for others it is a (minor in our opinion) threat to validity.

### 2020-03-25 response.txt
Minor change
in prog
maybe ready?

@@
 Only when at least one tool detects a mutant do we care.
-We note that the set of mutation operators can be easily extracted by examining the rule files in the open source repository for universalmutator.
+The full set of mutation operators can be easily extracted by examining the rule files in universalmutator github repo.
+
+Qualitative analysis is hard to make completely scientific, but we agree the paper omitted some of the forest for the trees, to make way for details of comparisons.
+We did in fact inspect a large number of the mutants in the "diff" (via random sampling for Java/Python).
+For Solidity, where we manually examined the majority of the mutants killed by at least one tool, most appeared to be semantically "troubling" at minimum, or clearly faults, though we think many changes of msg.sender to tx.origin flagged by non-Slither tools may be valid (if strange) code.
+For Java and Python, many mutants detected by one tool but not another involved deleted function calls or conditional changes (e.g.
+== null to != null) that would clearly introduce potential null pointer exceptions or method calls on None.
+These show differences in analysis methods rather than missing rules, but are clearly real(istic) faults.
+Other Java/Python cases were harder to understand, but many seemed to be real faults (e.g., introducing possible division by zero).
 

### 2020-03-25 response.txt
down to 1000

@@
 We would like to thank the anonymous reviewers for their helpful comments.
-
-Below are responses to questions and clarifications.
 
@@
 
-Most importantly: https://github.com/mutantsforstaticanalysis/data now provides all the raw data used in the paper.
-We hope this helps with qualitiative review.
-We also, below, discuss briefly the issue of whether the mutants detected by any tools were of actual interest, which is indeed a core issue.
-We do not examine mutants detected by no tools; due to the differential nature of the approach, we consider these potentially noise/uninteresting/not plausibly statically detected, but can simply ignore them.
-Only when at least one tool detects a mutant do we care.
+Most importantly: https://github.com/mutantsforstaticanalysis/data has the raw data.
+We hope this helps!
 The full set of mutation operators can be easily extracted by examining the rule files in universalmutator github repo.
 
-Qualitative analysis is hard to make completely scientific, but we agree the paper omitted some of the forest for the trees, to make way for details of comparisons.
+With respect to the meaningfulness of the mutants, first note that we do not examine mutants detected by zero tools, due to the differential nature of the approach; these are likely noise/uninteresting/not plausibly statically detected, but we just ignore them.
+Only when at least one tool detects a mutant do we care.
+
+Qualitative analysis is hard to make completely scientific, but we agree the paper omitted the forest for the trees, to make way for details of tool comparisons.
 We did in fact inspect a large number of the mutants in the "diff" (via random sampling for Java/Python).
-For Solidity, where we manually examined the majority of the mutants killed by at least one tool, most appeared to be semantically "troubling" at minimum, or clearly faults, though we think many changes of msg.sender to tx.origin flagged by non-Slither tools may be valid (if strange) code.
+For Solidity, where we manually examined the majority of the mutants killed by at least one tool, _most appeared to be semantically "troubling" at minimum_, or clearly faults, though we think many changes of msg.sender to tx.origin flagged by non-Slither tools may be valid (if strange) code.
 For Java and Python, many mutants detected by one tool but not another involved deleted function calls or conditional changes (e.g.
@@
 These show differences in analysis methods rather than missing rules, but are clearly real(istic) faults.
-Other Java/Python cases were harder to understand, but many seemed to be real faults (e.g., introducing possible division by zero).
+Other Java/Python cases were harder to understand, but many seemed to be real, statically detectable faults (e.g., introducing possible division by zero).
 
@@
 Adding Error Prone as well would be a good idea.
-In a way, [43] is a closer comparison to our work:  larger scope of interest, if higher risk of invalid results over that larger scope.
+In a way, [43] is a closer comparison to our work:  larger scope of interest, if higher risk of invalid results over that scope.
 
@@
 
-We used universalmutator for two reasons:  primarily because PIT is a bytecode-level mutation tool only, PMD requires source code, and we believe users want to inspect mutants for this application (and we needed to manually do some validation that the differences were meaningful).
-Second because we wanted to use a consistent set of operators and mutation approach, and universalmutator was the best tool for the other languages.
+We used universalmutator for two reasons:  primarily because PIT is a bytecode-level mutation tool only, PMD requires source code, and we believe users want to inspect mutants for this application (and we needed to manually do validation as R1 points out, which is hard for bytecode).
+We also wanted to use a consistent set of operators and mutation approach, and universalmutator was the best tool for the other languages.
 
-"better" in RQ4 means that we hypothesized that tools would produce _higher_ mutation scores (kill more mutants) for programs for which there were no findings in the original code.
+In RQ4 "better" means _higher_ mutation scores were hypothesized for programs for which there were no findings in the original code.
 
-Securify is a static analysis tool, but it "symbolically analyzes the contract’s dependency graph to extract
-precise semantic information from the code" before checking compliance/violation patterns.  The first step, unfortunately, like many symbolic approaches, can sometimes scale abominably on real code.
+Securify is a static analysis tool, but it "symbolically analyzes the contract’s dependency graph to extract precise semantic information from the code" before checking compliance/violation patterns.
+The first step, unfortunately, can sometimes scale abominably on real code.
 
@@
 
-Yes, we had to understand each tool's types of messages (for Pylint this was quite a painful process) and filter out ones that are essentially coding style rather than "this could be a bug!" issues, to the extent possible.
-We made considerably effort to ensure these choices were fair to the tools.
-Some tools made it easy by simply providing a setting to turn off "informational and style" issues, but for others it is a (minor in our opinion) threat to validity.
+Yes, we had to understand each tool's finding types (for Pylint this was quite painful) and filter out ones that are essentially stylistic rather than "here’s a bug!".
+We made considerable effort to ensure these choices were fair to the tools.
+Some tools made it easy by simply providing a setting to turn off "informational and style" issues, but for others it is a (minor, we think) threat to validity.

### 2020-03-27 response.txt
final response

@@
-We would like to thank the anonymous reviewers for their helpful comments.
+We thank the reviewers for their helpful comments.
 
@@
 
-Most importantly: https://github.com/mutantsforstaticanalysis/data has the raw data.
+Most importantly: https://github.com/mutantsforstaticanalysis/rawdata has the raw data.
 We hope this helps!
-The full set of mutation operators can be easily extracted by examining the rule files in universalmutator github repo.
+The full set of mutation operators can be easily extracted by examining the rule files in the universalmutator github repo.
 
 With respect to the meaningfulness of the mutants, first note that we do not examine mutants detected by zero tools, due to the differential nature of the approach; these are likely noise/uninteresting/not plausibly statically detected, but we just ignore them.
-Only when at least one tool detects a mutant do we care.
+Only when at least one tool detects a mutant do we care about it (and mostly when at least one does not).
 
@@
 For Solidity, where we manually examined the majority of the mutants killed by at least one tool, _most appeared to be semantically "troubling" at minimum_, or clearly faults, though we think many changes of msg.sender to tx.origin flagged by non-Slither tools may be valid (if strange) code.
-For Java and Python, many mutants detected by one tool but not another involved deleted function calls or conditional changes (e.g.
-== null to != null) that would clearly introduce potential null pointer exceptions or method calls on None.
-These show differences in analysis methods rather than missing rules, but are clearly real(istic) faults.
-Other Java/Python cases were harder to understand, but many seemed to be real, statically detectable faults (e.g., introducing possible division by zero).
+For Java, a large number of mutants detected by one tool but not another involved deleted function calls or conditional changes (e.g.
+== null to != null) that introduce potential null pointer exceptions.
+These show differences in analysis methods rather than missing detectors, and are clearly faulty code that should be detected.
+For Python, mutants producing code that would cause runtime (but not parsing) failures were common, often just an access to a now-undefined variable or obvious type mismatch.
+Again, all tools aim to detect these, so it is engine more than detector differences.
+Other Java/Python cases were harder to understand, but many seemed to be real, statically detectable faults.
+We will add this type of qualitative analysis.
 
 Yes, we actually used "number" of warnings, as a conservative measure of detection.
-If a mutant simply causes a tool to change the message it reports, then we do not consider that a kill.
-Similarly, if it loses one warning and produces another (which is in a sense the same thing, from a user's perspective), then it is not a kill.
-The core reason we don't consider the location/nature of the warning change is that in some cases a mutant causes a warning about code far from the mutant location.
-Therefore, we strictly require that a tool produce _more warnings_ when code is mutated, in order to be counted as detecting the mutant.
-We experimented with other definitions on a small set of smart contracts, and this was much better than other approaches, in terms of the results making sense.
-It is possible (though we didn't see this in practice with smart contracts) that it undercounts cases where 1) the original code produces warning A and 2) the mutated code "fixes" problem A and causes detection of a different problem, B.
-If the mutant doesn't change the issue, just changes the exact nature of the problem, then we argue the mutant didn't introduce a fault, it just changed some aspect of an existing fault, which is not what we are interested in.
-If we want to compare tools over existing faults, that is already doable.
+If a mutant only causes a tool to change finding message, then we do not consider that a kill.
+If it loses one warning and produces another then it is not a kill.
+The core reason we don't consider the location/nature of the warning change is that in some cases a mutant should cause a warning about code far from the mutant location.
+Therefore, we strictly require that a tool produce _more warnings_ when code is mutated, in order to be counted as detecting the mutant; we assume each mutant of interest potentially adds a fault, and fixes no existing problems.
+We experimented with other definitions on a set of smart contracts, and this approach produced the most useful results.
+It is thus possible (though we didn't see this happen with smart contracts) that we undercount cases where the original code produces warning A and the mutated code "fixes" problem A but introduces problem B.
+This seems like an acceptable risk to take, given mutants seldom fix faults, and better than having to parse/understand warning types/content to apply the method to every new language/tool.
 
-Good point, the source/bytecode choice is language dependent.  As to why we chose slither, see below.
+Good point, the source/bytecode choice is language dependent. As to why we chose Slither, see below.
+
+We do observe project level differences, and will report them.
+
+While not for Java, universalmutator has been extensively tuned for Python.
+
+We’ll add those references!
 
@@
 We chose PMD primarily because we started by focusing on [43] as a comparison paper, rather than [28], due to the highly-focused scope of [28] (only Defects4J bugs, thus many many fewer relevant warnings/aspects of code considered, and any biases in producing the Defects4J data set).
-Adding Error Prone as well would be a good idea.
-In a way, [43] is a closer comparison to our work:  larger scope of interest, if higher risk of invalid results over that scope.
+Adding Error Prone would be good, and should be possible by camera-ready.
+Arguably, [43] is the closer comparison to our work: larger scope of interest, albeit with higher risk of invalid results over that scope.
 
-"Importing" rules from one tool to another can be very helpful for actual users.
+"Importing" rules from one tool to another can be very helpful for users.
 If one tool uses a better underlying analysis engine (e.g., better intermediate language, better alias/data-flow/etc.
 approach), then adding a rule from another tool can result in much better _actual_ detection of bugs, or fewer false positives.
-Slither's IR and analysis are probably more scalable than Securify and more precise than SmartCheck, so even just using the same rule gives everyone better analysis results.
-Also, the rules are generalizations that expand on the rules in SmartCheck/Securify: the mutation approach makes generalizing a specific rule somewhat easier, we think, due to providing multiple examples, not just a high-level concept.
+Slither's IR and analysis are probably more scalable than Securify and more precise than SmartCheck, so even just using the same rules gives everyone better analysis results.
+Also, the rules are generalizations that expand on the rules in SmartCheck/Securify: using mutations makes generalizing a specific rule somewhat easier, we believe, due to (often multiple) concrete examples, not just a high-level concept.
 
-We used universalmutator for two reasons:  primarily because PIT is a bytecode-level mutation tool only, PMD requires source code, and we believe users want to inspect mutants for this application (and we needed to manually do validation as R1 points out, which is hard for bytecode).
-We also wanted to use a consistent set of operators and mutation approach, and universalmutator was the best tool for the other languages.
+We primarily used universalmutator because PIT is a bytecode-level mutation tool only.
+PMD requires source code, and we believe users want to inspect some mutants.
+We needed to manually do validation, as R1 points out, which is hard for bytecode.
+We wanted to use a consistent set of operators and mutation approach, and universalmutator was the best-supported tool for the other languages.
 
@@
 
-Securify is a static analysis tool, but it "symbolically analyzes the contract’s dependency graph to extract precise semantic information from the code" before checking compliance/violation patterns.
-The first step, unfortunately, can sometimes scale abominably on real code.
+Securify is a static analysis tool, but it uses souffle under the hood to solve Prolog constraints as part of building a dependency graph; this can scale very poorly.
 
-Our reasoning on the mutant ratio for Securify is that Securify does not notably produce warnings for mutants (which are probably bugs in some sense, even if not statically detectable) compared to the vast number of warnings it issues in general.
-Other research supports the claim that Securify is prone to false positives over all code, and its behavior on mutants is not appreciably different (see the 20x or more higher ratios for the other tools.
+Our reasoning on the mutant ratio for Securify is that Securify does not notably produce more warnings for mutants (which are usually bugs in some sense, even if not statically detectable) compared to the vast number of warnings it issues in general.
+Other research supports the claim that Securify is prone to false positives, and its behavior on mutants is not appreciably different than on other code, while SmartCheck and Slither have 20x better ratios.
 
@@
 
-Yes, we had to understand each tool's finding types (for Pylint this was quite painful) and filter out ones that are essentially stylistic rather than "here’s a bug!".
+Yes, we had to understand each tool's finding types (for Pylint this was painful) and filter out ones that are essentially stylistic rather than "here’s a bug!".
 We made considerable effort to ensure these choices were fair to the tools.

### 2020-04-15 experiments.tex
current draft

@@
 Can we confirm results from other evaluation methods, or informal opinion?
-\item {\bf RQ4:}  Are mutation scores for clean programs better than those for programs for which tools report findings before the program is mutated?
-\item {\bf RQ5:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
+\item {\bf RQ4:}  Are mutation scores for clean programs higher (more mutants killed) than those for programs for which tools report findings before mutation?
+\item {\bf RQ5:}  Are killed mutants reasonablely similar to actual faults?  That is, is the data above based on meaningful changes to code?
+\item {\bf RQ6:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
  different tools are good at/bad at, and extend this information to improve tools by addressing identified weaknesses?
-\item {\bf RQ6:}  How easy is it identify low-hanging fruit for improving tools \emph{without} prioritization to remove conceptually redundant mutants?
+\item {\bf RQ7:}  How easy is it to identify low-hanging fruit for improving tools \emph{without} prioritization to remove conceptually redundant mutants?
   \end{itemize}
@@
 showed that the ``clean program assumption'' for testing, the idea that coverage will be similar for faulty and fixed versions of a program, is a threat to the validity of investigations of the relationship between coverage and fault detection \cite{CleanProgram}, and we want to establish that this is unlikely to be the case for our approach.
-{\bf RQ5} and {\bf RQ6} receive only preliminary answers, for the smart contract tools.
+{\bf RQ5} is somewhat inherently qualitative and incomplete; we cannot analyze all mutants on which results are based manually, and understanding the mutants and tool warnings completely would require deep understanding of all the software systems investigated.
+However, in many cases, the impact of a mutant is clear, and the reason for warnings is obvious.
+This was often enough the case that, as we discuss below, we are confident mutants that distinguish tools are meaningful (missed) opportunities for static analysis tools.
+{\bf RQ6} and {\bf RQ7} receive only preliminary answers, for the smart contract tools.
+We used the Universal Mutator tool for all experiments, since e.g., PIT \cite{pitmut} does not produce source-level mutants and the PMD tool requires source code; we also wanted to use a consistent mutation generation method for all languages, to the extent possible.
 
@@
 For the three contracts clean for all tools, Slither performed almost exactly as it did over contracts in general, and the other tools performed worse, by about the same margin as they did for their own clean contracts.
-For our approach, we only need a week version of the ``clean program assumption'':  the threat is that kills may be under-reported for non-clean programs, due to interference with findings for the original code.
+For our approach, we only need a weak version of the ``clean program assumption'':  the threat is that kills may be under-reported for non-clean programs, due to interference with findings for the original code.
 It is not a problem if mutation scores are \emph{worse} for programs where a tool reports no findings for the un-mutated code.
@@
 
-\subsubsection{Improving Slither {\bf (RQ5 and RQ6)}}
+\subsubsection{Qualitative Examination of Mutants (RQ5)}
+
+Because examining all mutants manually is infeasible, we focused primarily on examining mutants in the \emph{difference}.
+Because our technique is differential, only mutants detected by at least one tool, but not detected by all tools, inform the comparison of static analysis tools.
+The vast majority of the mutants in this set were meaningful.
+We do not believe that all of them represent faults; some are harmless but unusual code changes.
+A large number of the instances where use of {\tt tx.origin} in place of {\tt msg.sender} was flagged seem to us to be strange but not necessarily incorrect code.
+On the other hand, at the risk of false positives, it is not unreasonable for tools to report some of these.
+Our estimate is that, ignoring {\tt tx.origin} cases, at least 70\% of the mutants detected by one, but not all, tools, represent real bugs, and the failure to detect is roughly equally due to missing detectors and imprecise analysis.
+
+\subsubsection{Improving Slither {\bf (RQ6 and RQ7)}}
 
@@
 
-Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants, suggesting that for {\bf RQ6} it is indeed hard to use mutation analysis results without prioritization.
+Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants, suggesting that for {\bf RQ7} it is indeed hard to use mutation analysis results without prioritization.
 A large majority of the mutants we inspected involved either the missing
@@
 SmartCheck and Securify tend to identify most (though not all) uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule, intended to reduce false positives.
-It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ5} and {\bf RQ6}.
+It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ6} and {\bf RQ7}.
 
@@
 
+\subsubsection{Qualitative Examination of Mutants (RQ5)}
+
+For Java, again, the large majority of mutants we inspected where at least one tool detected the mutant and at least one tool did not detect it were definitely meaningful.
+In particular, for Java, the large majority of mutants involved either deleted method calls or changes to conditionals (e.g., {\tt == null} to {\tt != null} that would clearly introduce potential null pointer exceptions (NPEs).
+While the differences between Solidity tools were often due to different detectors, the Java differences seemed mostly rooted in analysis engine methods; all tools aim to warn about potential NPEs.
+
 \subsection{Python Tools}
@@
 
+\subsubsection{Qualitative Examination of Mutants (RQ5)}
+
+For Python, as with Java, many mutants we inspected involved deletion of a call, leading to a likely method call or access to a variable containing {\tt None}.
+More Python mutants were more difficult to understand than for Solidity or Java, without a grasp of the larger program context, but a large number were still clearly differences in analysis engine capabilities, since all tools aim to detect failure to guard against improper uses of {\tt None}, the Python equivalent of a Java NPE.
+
 \subsection{Threats to Validity}
@@
 We used the Universal Mutator \cite{universalmutator,regexpMut}, which aggressively produces large numbers of mutants, but does not target any particular software defect patterns, to generate all mutants.
-We will make available a repository containing our full raw results plus details on contracts and projects selected, for replication and further analysis, upon acceptance (to avoid breaking blinding).
+This tool has been heavily tuned and used for for Solidity and Python, and moderately tuned for Java.
+The Java mutation also benefits from extensive tuning on C programs, which share syntactic mutation patterns with both Java and Solidity.
+There is no room in the paper to present the exact set of projects analyzed, but we have provided an (anonymized) github repository containing raw results for inspection by reviewers, or further analysis by other researchers.
+We will de-anonymize this if the paper is accepted.

### 2020-04-15 experiments.tex
ok

@@
 Our primary experimental results are a set of comparisons of tools using our method, for three languages: Solidity (the most popular language for smart contracts), Java, and Python.
-We also use these results to answer a set of research questions that consider the utility of our method:
+We used the Universal Mutator tool for all experiments, since e.g., PIT \cite{pitmut} does not produce source-level mutants and the PMD tool requires source code; we also wanted to use a consistent mutation generation method for all languages, to the extent possible; for Solidity and Python, we believe the Universal Mutator is simply the best available tool.
+We also our results to answer a set of research questions that consider the utility of our method:
 
@@
 \item {\bf RQ6:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
- different tools are good at/bad at, and extend this information to improve tools by addressing identified weaknesses?
-\item {\bf RQ7:}  How easy is it to identify low-hanging fruit for improving tools \emph{without} prioritization to remove conceptually redundant mutants?
+ different tools are good at/bad at, and extend this information to improve tools by addressing identified weaknesses?  How does this compare to using mutants that have not been prioritized?
   \end{itemize}
@@
 This was often enough the case that, as we discuss below, we are confident mutants that distinguish tools are meaningful (missed) opportunities for static analysis tools.
-{\bf RQ6} and {\bf RQ7} receive only preliminary answers, for the smart contract tools.
-We used the Universal Mutator tool for all experiments, since e.g., PIT \cite{pitmut} does not produce source-level mutants and the PMD tool requires source code; we also wanted to use a consistent mutation generation method for all languages, to the extent possible.
+For {\bf RQ6}  we have only preliminary answers, for the smart contract tools.
 
@@
 
-\subsubsection{Improving Slither {\bf (RQ6 and RQ7)}}
+\subsubsection{Improving Slither {\bf (RQ6)}}
 
@@
 
-Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants, suggesting that for {\bf RQ7} it is indeed hard to use mutation analysis results without prioritization.
+Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants, suggesting that it is indeed hard to use mutation analysis results without prioritization.
 A large majority of the mutants we inspected involved either the missing
@@
 SmartCheck and Securify tend to identify most (though not all) uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule, intended to reduce false positives.
-It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ6} and {\bf RQ7}.
+It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ6}.
 
@@
 We used the Universal Mutator \cite{universalmutator,regexpMut}, which aggressively produces large numbers of mutants, but does not target any particular software defect patterns, to generate all mutants.
-This tool has been heavily tuned and used for for Solidity and Python, and moderately tuned for Java.
-The Java mutation also benefits from extensive tuning on C programs, which share syntactic mutation patterns with both Java and Solidity.
-There is no room in the paper to present the exact set of projects analyzed, but we have provided an (anonymized) github repository containing raw results for inspection by reviewers, or further analysis by other researchers.
-We will de-anonymize this if the paper is accepted.
+There is no room in the paper to present the exact set of projects analyzed, but we have provided an (anonymized) github repository containing raw results for inspection by reviewers, or further analysis by other researchers (\url{the repo}).

### 2020-04-15 introduction.tex
ok

@@
 
-Static analysis of code is one of the most effective ways to avoid defects in software, and when security is a concern, applying effective static analysis tools is essential.
-Static analysis can find problems that are extremely hard to detect by testing, e.g.
-when the inputs triggering a bug are hard to find.
+Static analysis of code is one of the most effective ways to avoid defects in software, and, when security is a concern, is simply essential.
+Static analysis can find problems that are extremely hard to detect by testing, when the inputs triggering a bug are hard to find.
 Static analysis is also often more efficient than testing; a bug that takes a fuzzer like AFL days to find may be immediately identified by a good static analysis tool.
@@
 Most such use has been in research efforts, rather than practical testing efforts, though there has been sporadic use by interested developers.
-In an ASE 2015 \cite{groce2015verified} paper and a 2018 journal extension \cite{groce2018verified} of that paper, Groce et al.
-proposed examining individual mutants that survive a formal verification or automated test generation process to detect and correct weaknesses in a specification or test generator.
-The approach was able to expose bugs in a heavily-tested module of the Linux kernel \cite{mutKernel} and improve a heavily used test generator for the {\tt pyfakefs} file system, allowing it to detect multiple new bugs.
-Recently, mutation analysis has been adopted in industrial settings, though not for actual examination of all surviving mutants \cite{MutGoogle,ivankovic2018industrial}, a practice that is hard to scale to large code bodies.
+Groce et al.
+\cite{groce2015verified,groce2018verified} proposed examining individual mutants that survive a rigorous testing and verification effort to detect and correct weaknesses in such efforts.
+The approach was able to expose bugs in a heavily-tested module of the Linux kernel \cite{mutKernel} and a widely used Python file system.
+Recently, mutation analysis has been adopted in industrial settings, though not for actual examination of all surviving mutants \cite{MutGoogle,ivankovic2018industrial}, a practice that is hard to scale.
 

### 2020-04-16 experiments.tex
too large, but on the way

@@
-\section{Experimental Results: Comparison and Improvement of Static Analysis Tools}
+\section{Experimental Results}
 
@@
 \item {\bf RQ4:}  Are mutation scores for clean programs higher (more mutants killed) than those for programs for which tools report findings before mutation?
-\item {\bf RQ5:}  Are killed mutants reasonablely similar to actual faults?  That is, is the data above based on meaningful changes to code?
+\item {\bf RQ5:}  Are killed mutants flagged due to real faults, where the finding is related to the introduced fault?
+That is to say, is the evaluation based on meaningful semantic changes some (but not all) tools detect?
 \item {\bf RQ6:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
@@
 showed that the ``clean program assumption'' for testing, the idea that coverage will be similar for faulty and fixed versions of a program, is a threat to the validity of investigations of the relationship between coverage and fault detection \cite{CleanProgram}, and we want to establish that this is unlikely to be the case for our approach.
-{\bf RQ5} is somewhat inherently qualitative and incomplete; we cannot analyze all mutants on which results are based manually, and understanding the mutants and tool warnings completely would require deep understanding of all the software systems investigated.
+Our answer to {\bf RQ5} is somewhat inherently qualitative and incomplete; we cannot analyze all mutants on which results are based manually, and understanding the mutants and tool warnings completely would require deep understanding of all the software systems investigated.
 However, in many cases, the impact of a mutant is clear, and the reason for warnings is obvious.
@@
 
-\subsubsection{Qualitative Examination of Mutants (RQ5)}
+Because examining all mutants manually is infeasible,  for {\bf RQ5} we focused primarily on examining mutants in the \emph{difference}: mutants detected by at least one tool, but not detected by all tools, the only ones that actually influence the comparison of tools.
+The vast majority of the mutants in this set were meaningful semantic changes a static analysis tool could be expected to detect, and the findings produced by tools were relevant to the nature of the fault.
+We do not believe that all mutants represent definite faults; some are harmless but unusual code changes.
+A large number of the instances where use of {\tt tx.origin} in place of {\tt msg.sender} was flagged seem to us to be strange, but not necessarily incorrect, code.
+On the other hand, at the risk of false positives, it is not unreasonable for tools to report such strange code.
+Our estimate is that, ignoring {\tt tx.origin} cases, \emph{at least 70\% of the mutants detected by one, but not all, tools, represent realistic bugs}, and failure to detect is roughly equally due to missing detectors and imprecise analysis.
 
-Because examining all mutants manually is infeasible, we focused primarily on examining mutants in the \emph{difference}.
-Because our technique is differential, only mutants detected by at least one tool, but not detected by all tools, inform the comparison of static analysis tools.
-The vast majority of the mutants in this set were meaningful.
-We do not believe that all of them represent faults; some are harmless but unusual code changes.
-A large number of the instances where use of {\tt tx.origin} in place of {\tt msg.sender} was flagged seem to us to be strange but not necessarily incorrect code.
-On the other hand, at the risk of false positives, it is not unreasonable for tools to report some of these.
-Our estimate is that, ignoring {\tt tx.origin} cases, at least 70\% of the mutants detected by one, but not all, tools, represent real bugs, and the failure to detect is roughly equally due to missing detectors and imprecise analysis.
+Because our random contracts' quality might be low, we also checked our results on 30 contracts from the Solidity documentation, the \emph{Mastering Ethereum} book \cite{masteringEth}, and a handful of selected, recent, higher quality blockchain contracts.
+Slither had a mean mutation score of 0.11, vs.
+0.04 for SmartCheck and 0.01 for Securify.
+Associated mutant ratios were 0.38, 0.08, and 0.007.
+Mutant kill overlap was also similar; in fact, Figure \ref{fig:examplevenn} shows the results: Slither is A, SmartCheck is B, and Securify is C.
 
@@
 Based on the differential mutation analysis, we identified three low-hanging fruit to improve the performance of Slither.
+We chose Slither in part because it seems to have a better underlying intermediate language and analysis engine, and thus is likely to produce better results for the same rule than the other tools.
 The process was simple.
@@
   {\scriptsize
+      {\bf Mutant showing Boolean constant misuse.}
+
 \noindent \begin{code}
@@
       \end{code}
+
       }
-      \caption{Mutant showing {\bf Boolean constant misuse}.}
-      \label{fig:boolean}
-    \end{figure}
 
-    \begin{figure}
       {\scriptsize
+              {\bf Mutant showing Type-based tautologies.}
 \begin{code}
@@
       }
-      \caption{Mutant showing {\bf Type-based tautologies}.}
-      \label{fig:tautology}      
-\end{figure}
 
-\begin{figure}
       {\scriptsize
+      {\bf Mutant showing Loss of precision.}        
 \begin{code}
@@
       \end{code}
+
 }
-      \caption{Mutant showing {\bf Loss of precision}.}
-      \label{fig:divmul}            
+        \caption{Examples of mutants leading to new detectors.}
+       \label{fig:newdetect}
     \end{figure}
@@
 This detector is actually split into two detectors, one for this serious issue, and an informational/stylistic detector that notes that code such as {\tt if (x == true)}, while semantically harmless, is difficult to read.
-This problem was easily identified from cases such as the mutant in Figure \ref{fig:boolean}, killed by SmartCheck but not Slither:
+This problem was easily identified from cases such as the mutant in Figure \ref{fig:newdetect}, killed by SmartCheck but not Slither:
 
@@
 {\tt y < 512} where {\tt y}'s type is {\tt int8}.
-Again, the problem was easily identified by mutants killed by SmartCheck but not Slither such as the one in Figure \ref{fig:tautology}.
+Again, the problem was easily identified by mutants killed by SmartCheck but not Slither.
 
@@
 This is a fairly important problem, given the frequency with which Solidity code performs financial calculations where maximum precision is desired.
-SmartCheck provides a detector for such precision losses \url{https://github.com/smartdec/smartcheck/blob/master/rule\_descriptions/SOLIDITY\_DIV\_MUL/}, which enabled it to detect mutants such as the one shown in Figure\ref{fig:divmul}.
+SmartCheck provides a detector for such precision losses \url{https://github.com/smartdec/smartcheck/blob/master/rule\_descriptions/SOLIDITY\_DIV\_MUL/}.
 \end{enumerate}
@@
 A large majority of the mutants we inspected involved either the missing
-{\tt return} problem noted in the introduction, or replacing {\tt msg.sender} with {\tt tx.origin}; Slither, of course, has a detector for misuses of {\tt tx.origin}, and we believe (but are not sure) that almost all of the differences with respect to such changes are due to intentional behavior.
+{\tt return} problem noted in the introduction, or replacing {\tt msg.sender} with {\tt tx.origin}; Slither has a detector for misuses of {\tt tx.origin}.
 SmartCheck and Securify tend to identify most (though not all) uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule, intended to reduce false positives.
 It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ6}.
-
-\subsubsection{Confirmation of Results}  Because our random contracts' quality might be low, we also checked our results on 30 contracts from the Solidity documentation, the \emph{Mastering Ethereum} book \cite{masteringEth}, and a handful of selected, recent, higher quality blockchain contracts.
-Slither had a mean mutation score of 0.11, vs.
-0.04 for SmartCheck, and 0.01 for Securify.
-Associated mutant ratios were 0.38, 0.08, and 0.007.
-Mutant kill overlap was also similar; in fact, Figure \ref{fig:examplevenn} shows the results: Slither is A, SmartCheck is B, and Securify is C.
 
@@
   }
-\caption{Blog posts and forum threads discussing Java static analysis tools.}
+\caption{Discussions of Java static analysis tools.}
 \label{fig:blog}
@@
 
-\subsubsection{Qualitative Examination of Mutants (RQ5)}
-
-For Java, again, the large majority of mutants we inspected where at least one tool detected the mutant and at least one tool did not detect it were definitely meaningful.
-In particular, for Java, the large majority of mutants involved either deleted method calls or changes to conditionals (e.g., {\tt == null} to {\tt != null} that would clearly introduce potential null pointer exceptions (NPEs).
+For Java, again, the large majority of mutants we inspected for {\bf RQ5}, where at least one tool detected the mutant and at least one tool did not detect it were definitely meaningful.
+In particular, for Java, the large majority of mutants involved either deleted method calls or changes to conditionals (e.g., {\tt == null} to {\tt != null} that would clearly introduce potential null pointer exceptions (NPEs), and such a possible NPE was the produced finding.
 While the differences between Solidity tools were often due to different detectors, the Java differences seemed mostly rooted in analysis engine methods; all tools aim to warn about potential NPEs.
+Because the number of mutants we could examine and understand was smaller, we are less confident in making a probabilistic estimate than with Solidity, but it was clear the basis of the comparison was primarily realistic faults, which some tools detected and others did not.
 
@@
   }
-\caption{Blog posts and forum threads discussing Python static analysis tools.}
+\caption{Discussions of Python static analysis tools.}
 \label{fig:blogpython}
@@
 Pyflakes is also well liked, and supposedly less verbose than Pylint; we did not see this in our results, but that may be because we configured Pylint, and pyflakes aims to focus on ``actual bugs'' so it is equally verbose for mutants, which are likely to be actual bugs.
-
 For {\bf RQ4} there were no clean projects for the known-to-be-aggressive Pylint.
@@
 
-\subsubsection{Qualitative Examination of Mutants (RQ5)}
-
-For Python, as with Java, many mutants we inspected involved deletion of a call, leading to a likely method call or access to a variable containing {\tt None}.
-More Python mutants were more difficult to understand than for Solidity or Java, without a grasp of the larger program context, but a large number were still clearly differences in analysis engine capabilities, since all tools aim to detect failure to guard against improper uses of {\tt None}, the Python equivalent of a Java NPE.
+For Python, as with Java, many mutants we inspected for {\bf RQ5} involved deletion of a call or change in comparison operator, leading to a likely method call or access to a variable containing {\tt None}, and the generated finding reported this problem.
+More Python mutants were extremely difficult to understand than for Solidity or Java, without a grasp of the larger program context, but a large number were clearly due to differences in analysis engine capabilities, since all tools aim to detect failure to guard against improper uses of {\tt None}, the Python equivalent of a Java NPE.
+Again, based on a qualitative examination of many mutants, we believe the results to primarily rest on realistic faults, which some tools correctly detected and others did not.
 
@@
 We used the Universal Mutator \cite{universalmutator,regexpMut}, which aggressively produces large numbers of mutants, but does not target any particular software defect patterns, to generate all mutants.
-There is no room in the paper to present the exact set of projects analyzed, but we have provided an (anonymized) github repository containing raw results for inspection by reviewers, or further analysis by other researchers (\url{the repo}).
+There is no room in the paper to present the exact set of projects analyzed, but we have provided an (anonymized) github repository containing raw results for inspection by reviewers, or further analysis by other researchers (\url{https://github.com/mutantsforstaticanalysis/rawdata}).

### 2020-04-16 introduction.tex
too large, but on the way

@@
 The approach was able to expose bugs in a heavily-tested module of the Linux kernel \cite{mutKernel} and a widely used Python file system.
-Recently, mutation analysis has been adopted in industrial settings, though not for actual examination of all surviving mutants \cite{MutGoogle,ivankovic2018industrial}, a practice that is hard to scale.
+Recently, mutation analysis has been adopted in industrial settings, though not for actual examination of all surviving mutants \cite{MutGoogle,ivankovic2018industrial}, due to scale.
 
@@
 
-We can say that a static analysis tool kills a mutant when the number of warnings or errors, which we call \emph{findings}, that are not informational or optimization related, i.e., that are \emph{fault-related}, produced with respect to the code \emph{increases} for the mutated version, compared to the original code.
-That is, when the tool finds more bugs for the mutated code.
+We can say that a static analysis tool kills a mutant when the \emph{number of warnings or errors}, which we call \emph{findings}, that are not informational or optimization related, i.e., that are \emph{fault-related}, produced with respect to the code \emph{increases} for the mutated version, compared to the original code.
+That is, when the tool ``finds more bugs'' for the mutated code.
 This difference may be most easily interpreted when the original code produces no findings; we call such code \emph{clean}.
@@
 However, even for non-clean code, \emph{most detected mutants should produce a new warning}.
+We count findings, rather than consider their location or type, because some mutants cause a fault at a far-removed location.
+Forcing tools to produce an \emph{additonal} warning is a conservative and automatable estimate of mutant detection.
 
@@
 The tool cannot simply flag all code as having a problem or it will perform poorly at the task of \emph{distinguishing} the mutated code from non-mutated, and presumably at least \emph{more} correct, code.
-Based on verification and testing uses of mutation, it is safe to say that usually at minimum 40\%, often 50-60\%, and frequently up to 80\%+ \cite{mutKernel,groce2018verified,le2014mucheck}, of mutants are not semantically equivalent to the original code, and are actually fault-inducing and detectable by some test or property \cite{TCE,impactEquiv,smith2009should}, so the task presented to a static analysis tool is the generalization of the task we ideally expect static analysis to perform:  to identify faulty code, without executing it, and, most critically, \emph{to distinguish faulty from correct code}.
-Obviously, many faults cannot be identified statically without a complete specification, or without unreasonable analysis cost and precision, but the measure of performance here is meant to be mostly \emph{relative} to other tools applied to the same code; this is primarily a \emph{differential} approach.
-In other words, a key notion is that while most mutants cannot be detected statically, the ones that are tend to be \emph{true positives}: if they were real code changes, they would be faults.
+Based on verification and testing uses of mutation, it is safe to say that usually at minimum 40\%, often 50-60\%, and frequently up to 80\%+ \cite{mutKernel,groce2018verified,le2014mucheck}, of mutants are not semantically equivalent to the original code \cite{TCE,impactEquiv,smith2009should}, so the task presented to a static analysis tool is simply the core functionality of static analysis \emph{to distinguish faulty from correct code without execution}.
+Obviously, many faults cannot be identified statically without a complete specification, or without unreasonable analysis cost and precision, but the measure of performance here is \emph{relative} to other tools applied to the same code; this is primarily a \emph{differential} approach.
+In other words, a key notion is that while most mutants cannot be detected statically, the ones that are tend to be \emph{true positives}: \emph{if} they were real code changes, they would be faults.
+We manually confirmed that for most mutants in our experiments where any tool detected the mutant (and at least one tool did not), the changes were indeed ones that would be real faults if present in the code.
 
@@
 Comparing mutant results also leads to the idea of \emph{improving} static analysis tools by examining mutants detected by another tool, and thus known to be in-principle detectable, but not detected by the tool to be improved.
+Improving tools by adding detectors is useful because, even if all tools had the same set of detectors, they would not all report the same bugs; difference choices in intermediate language and tradeoffs made to avoid false positives would make the use of multiple tools with shared detectors essential for thorough analysis.
+And, of course, if one tool does have a much better underlying engine, it is beneficial to users that the ``best'' tool incorporate all detection rules.
+
 Of course, any faults in code, not just mutants, could serve this purpose.
@@
 Moreover, because mutants follow syntactic patterns, searching for similar mutants/faults the tool to be improved does not detect is much easier than with arbitrary faults, and can be partly automated.
-Of course, knowing which mutation patterns are of interest requires human effort.
-As with efforts to improve test suites, manually searching through all mutants can be an onerous task, especially for large-scale evaluations.
+However, as with efforts to improve test suites, manually searching through all mutants can be an onerous task, especially for large-scale evaluations.
 We therefore also introduce the idea of prioritizing mutants, using a Furthest-Point-First \cite{Gonzalez} algorithm and distance metric inspired by previous work on helping developers sort through failing test cases to find distinct faults \cite{PLDI13,distMut}, to help static analysis tool developers find interesting patterns without wading through numerous uninteresting mutants not identified for similar reasons.
@@
 Changing arithmetic and comparison operators and numeric constants (incrementing, decrementing, or changing to 0 or 1) account for most of the non-statement-deletion mutants.
-We cannot see how any tool could have a useful detection rule that looks for these kinds of changes: a tool obviously would not flag all instances of 0 and 1 constants, or of addition, and remain a useful tool.
-Finally, many mutants, though many fewer than in these categories, add {\tt break} or {\tt continue} statements.
+We cannot see how any tool could have a useful detection rule that looks for these kinds of changes: a tool obviously could not flag all instances of 0 and 1 constants, or of addition, and remain a useful tool.
+Finally, many mutants, though far fewer than in the previously noted categories, add {\tt break} or {\tt continue} statements.
 Again, we do not see how a tool could take advantage of this without either genuinely understanding likely-bad control flow patterns, in which case, we think it ``deserves'' to kill more mutants, or flagging so many legitimate uses of {\tt break} and {\tt continue} in un-mutated code that it would have a bad mutant ratio, due to producing more warnings than other tools.
@@
 \begin{code}
-pragma solidity ^0.4.0;
-
 contract SimpleStorage \{
@@
 }
-\caption{A simple example Solidity smart contract from \url{https://solidity.readthedocs.io/en/v0.4.24/introduction-to-smart-contracts.html}.}
+\caption{A simple example Solidity smart contract}
 \label{fig:sol424intro}
@@
 
-Consider the code in Figure \ref{fig:sol424intro}, from the Solidity 0.4.24 ``Introduction to Smart Contracts''.
+Consider the code in Figure \ref{fig:sol424intro} \cite{solintro}.
 The Universal Mutator tool \cite{universalmutator,regexpMut}, which has been extensively tuned for Solidity's grammar (though not to target any particular vulnerabilities), and is the only smart contract mutation tool referenced in the Solidity documentation (\url{https://solidity.readthedocs.io/en/v0.5.12/resources.html}), produces seven valid, non-redundant (by Trivial Compiler Equivalence \cite{TCE}) mutants for this trivial example code.
@@
 Both tools also detect (by increasing the number of warnings produced) four of the seven mutants.
-However, only one of the mutants detected is common to both tools: both tools detect changing the {\tt return} statement in the {\tt get} function to a call to {\tt selfdestruct} the smart contract.
+However, only one of the mutants detected is common to both tools: both tools detect changing the {\tt return} statement in the {\tt get} function to a call to {\tt selfdestruct} the smart contract, deleting it.
 Slither, but not SmartCheck, also detects replacing the assignment of {\tt storedData} in {\tt set} with either a {\tt selfdestruct} or {\tt revert}, or simply removing it altogether.

### 2020-04-16 prioritizing.tex
current draft
too large, but on the way

@@
 order to identify patterns for new detectors.  Dedicated developers may also
-simple want to scan all mutants their tool does not kill, for the same
+simply want to scan all mutants their tool does not kill, for the same
 purpose, analogous to what Groce et al. have proposed for automated
@@
 because many mutants will have the same change (e.g., changing {\tt +}
-to {\tt -}, the other ratios also often matter.  The Python
-{\tt Levenshtein} library's similarity ratio is used, as it is based
-on true minimal string edits; it reports similarity ratios between 0.0
-and 1.0.
-
+to {\tt -}, the other ratios also often matter.
+%The Python
+%{\tt Levenshtein} library's similarity ratio is used, as it is based
+%on true minimal string edits; it reports similarity ratios between 0.0
+%and 1.0.
 Our metric also incorporates a measure of the distance in the source
@@
 
-$$ 5.0 \times r(\mathit{edit}_1, \mathit{edit}_2) +$$
-$$0.1 \times r(\mathit{source}_1, \mathit{source}_2) +$$
-$$0.1 \times r(\mathit{mutant}_1, \mathit{mutant}_2) +$$
-$$0.5 \times \mathit{not\_same\_file} +$$
+$$ 5.0 \times r(\mathit{edit}_1, \mathit{edit}_2) + 0.1 \times r(\mathit{source}_1, \mathit{source}_2) +$$
+$$0.1 \times r(\mathit{mutant}_1, \mathit{mutant}_2) + 0.5 \times \mathit{not\_same\_file} +$$
 $$max(0.25, \frac{\mathit{line\_dist}(\mathit{mutant}_1, \mathit{mutant}_2)}{10})$$
@@
 
-We do not claim this is an optimal, or even tuned, metric; in the long run, it even may be completely replaced.
-One element that will likely remain is the focus on source code, rather than bytecode; static analysis tools generally analyze source, and some even work for code that does not compile.
-The metric we use
+We do not claim this is an optimal, or even tuned, metric; in the long
+run, it even may be completely replaced.  However, the metric we use
 provided useful results for a different problem, ranking mutants in an

### 2020-04-16 relatedwork.tex
too large, but on the way

@@
 This is a result of both algorithmic constraints and basic engineering trade-offs.
-While comparisons of static analysis tools \cite{CompareJavaTools,durieux2019empirical, Parizi,slither} have appeared in the literature for years, these generally involved large human effort and resulting smaller scale, or did not make a strong effort to address false positives, or restricted analysis to, e.g., a known defects set \cite{AllBugs}.
+While comparisons of static analysis tools \cite{CompareJavaTools,durieux2019empirical, Parizi,slither,pashchenko2017delta} have appeared in the literature for years, these generally involved large human effort and resulting smaller scale, or did not make a strong effort to address false positives, or restricted analysis to, e.g., a known defects set \cite{AllBugs, do2016toward}.
 

### 2020-04-20 experiments.tex
ok

@@
 Our primary experimental results are a set of comparisons of tools using our method, for three languages: Solidity (the most popular language for smart contracts), Java, and Python.
-We used the Universal Mutator tool for all experiments, since e.g., PIT \cite{pitmut} does not produce source-level mutants and the PMD tool requires source code; we also wanted to use a consistent mutation generation method for all languages, to the extent possible; for Solidity and Python, we believe the Universal Mutator is simply the best available tool.
-We also our results to answer a set of research questions that consider the utility of our method:
+We used the Universal Mutator tool for all experiments; for Solidity and Python, we believe the Universal Mutator is simply the best available tool.
+For Java, PIT \cite{pitmut} is more popular, but does not produce source-level mutants, needed for PMD and for manual inspection of results.
+We used our results to answer a set of research questions:
 
-\begin{itemize}
-\item {\bf RQ1:}  Does mutation analysis of static analysis tools produce actionable results?
-That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar in terms of mutation detection capability?
-\item {\bf RQ2:}  Does our approach provide additional information beyond simply counting findings for the original, un-mutated analyzed code?
-Do \emph{ratios} differ between programs, or does the number of mutants killed simply reflect the ``verbosity'' of each tool?
-\item {\bf RQ3:}  Do the rankings that raw kills and ratios establish agree with other sources of information about the effectiveness of the evaluated tools?
-Can we confirm results from other evaluation methods, or informal opinion?
-\item {\bf RQ4:}  Are mutation scores for clean programs higher (more mutants killed) than those for programs for which tools report findings before mutation?
-\item {\bf RQ5:}  Are killed mutants flagged due to real faults, where the finding is related to the introduced fault?
-That is to say, is the evaluation based on meaningful semantic changes some (but not all) tools detect?
+\begin{itemize}[labelsep=3pt,leftmargin=12pt]
+\item {\bf RQ1:}  Does mutation analysis of static analysis tools produce actionable results?  That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar?
+\item {\bf RQ2:}  Does our approach provide additional information beyond simply counting findings for the original, un-mutated analyzed code?  Do \emph{ratios} differ between static analysis tools?
+\item {\bf RQ3:}  Do the rankings that raw kills and ratios establish agree with other sources of information about the effectiveness of the evaluated tools?  Can we independently confirm our results?
+\item {\bf RQ4:}  Do tools detect more mutants in programs for which they produce no warnings, initially?
+\item {\bf RQ5:}  Are mutants distinguishing tools usually flagged due to real faults, where the finding is related to the introduced fault; that is, are our results usually \emph{meaningful}?  
 \item {\bf RQ6:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
- different tools are good at/bad at, and extend this information to improve tools by addressing identified weaknesses?  How does this compare to using mutants that have not been prioritized?
+ different tools are good at/bad at, and use this information to improve tools?  How does this compare to using mutants that have \emph{not} been prioritized?
   \end{itemize}
@@
 \begin{figure}
-  \includegraphics[width=0.6\columnwidth]{solidity.png}
+  \includegraphics[width=0.4\columnwidth]{solidity.png}
   \caption{Mutants killed by Solidity static analysis tools.}
@@
     Slither & 2.37 & 1.0 & 0.09 & 0.09 & 0.038 \\
+    Clean (39) & - & - & 0.11 & 0.11 & -\\
     \hline
     SmartCheck & 1.89 & 1.0 & 0.05 & 0.05 & 0.026 \\
+    Clean (27) & - & - & 0.03 & 0.01 & - \\
     \hline
     Securify & 24.65 & 17.0 & 0.03 & 0.02 &  0.001 \\
+    Clean (5) & - & - & 0.00 & 0.00 & - \\
     \hline
+
   \end{tabular}
@@
 
-\begin{table}
-  \begin{tabular}{l|r|r|r|r|r}
-    & & \multicolumn{2}{|c|}{} & \multicolumn{2}{|c}{Clean For All (3)} \\
-    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} &  \multicolumn{2}{|c}{Mutation Score}\\
-    Tool & Contracts & Mean & Median & Mean & Median\\
-    \hline
-    \hline
-    Slither & 39 & 0.11 & 0.11 & 0.09 & 0.09 \\
-    \hline
-    SmartCheck & 27 & 0.03 & 0.01 & 0.03 & 0.00 \\
-    \hline
-    Securify & 5 & 0.00 & 0.00 & 0.00 & 0.00 \\
-    \hline
-  \end{tabular}
-  \caption{Solidity tool clean contract results.}
-  \label{tab:cleansolidity}
-\end{table}
-
 Figure \ref{fig:solidityvenn} shows the mutants killed by the Solidity analysis tools.
-Tables \ref{tab:scoresolidity} and \ref{tab:cleansolidity} provide numeric details of the results, including the \emph{ratio} for each tool, adjusting its mutation scores by its' general tendency to produce findings.
-First, a user examining these results would suspect that Slither and SmartCheck are both useful tools, and should likely both be applied in a high-risk security-sensitive context like smart contract development.
-Second, a user might suspect that the large number of findings produced, and smaller number of mutants killed, for Securify, mean that including Securify in the static analysis tool stable is a more difficult decision.
+Table \ref{tab:scoresolidity} provides numeric details of the results, including the \emph{ratio} for each tool, adjusting its mutation scores by its' general tendency to produce findings.
+The second row for each tool shows the number of contracts for which it reported no findings, and the mutation scores over those contracts, only.
+A user examining these results would suspect that Slither and SmartCheck are both useful tools, and should likely both be applied in a high-risk security-sensitive context like smart contract development.
+A user might also suspect that the large number of findings produced, and smaller number of mutants killed, for Securify, means that whether to apply Securify is a more difficult decision.
 On the one hand, Securify does detect nearly as many mutants it alone can identify as SmartCheck.
@@
 For example, Securify often detected when an ERC20 token contract's guard preventing making the special 0x0 address the owner of a contract was removed, and issued the error {\tt Violation for MissingInputValidation}.
-Detecting such missing guards is probably useful, but formulating a way to do it without producing false positives is non-trivial.
+Detecting such missing guards is probably useful, but doing so without producing false positives is non-trivial.
 We wanted to show that mutants could identify \emph{useful} but \emph{easy to implement} missing detectors.
-Examining the first few mutants, we identified three such, based on mutants killed by either SmartCheck or Securify, or both:
+Examining only a few mutants, we identified three:
 
@@
 Without prioritization, on average a developer would have to look at more than 200, 80, and 400 mutants, respectively, to find instances of these problems.
-Interestingly, the very fact that these instances are ``needles in a haystack'' among the mutants not killed by Slither means that the results in Figure \ref{fig:solidityvenn} and Tables \ref{tab:scoresolidity} and \ref{tab:cleansolidity} are almost unaltered by our improvements to Slither: our analysis is fairly robust to modest tool improvements, unless added detectors account for a large number of mutants not detected by the tool.
+Interestingly, the very fact that these instances are ``needles in a haystack'' among the mutants not killed by Slither means that the results in Figure \ref{fig:solidityvenn} and Table \ref{tab:scoresolidity} are almost unaltered by our improvements to Slither: our analysis is fairly robust to modest tool improvements, unless added detectors account for a large number of mutants not detected by the tool.
 Adding such detectors will also only improve mutation ratio if they do not add many false positives.
@@
 \begin{figure}
-  \includegraphics[width=0.6\columnwidth]{java.png}
+  \includegraphics[width=0.4\columnwidth]{java.png}
   \caption{Mutants killed by Java static analysis tools.}
@@
     SpotBugs & 28.93 & 14.00 & 0.07 & 0.07 & 0.002 \\
+    Clean (3) & - & - & 0.05 & 0.06 - \\
     \hline
     PMD & 53.73 & 32.00 & 0.07 & 0.07 & 0.001 \\
+    Clean (0) & - & - & - & - & - \\
     \hline
     Infer & 11.60 & 3.00 & 0.00 & 0.00 &  0.000 \\
+    Clean (6) & - & - & 0.00 & 0.00 \\
     \hline
@@
 
-\begin{table}
-  \begin{tabular}{l|r|r|r}
-    & & \multicolumn{2}{|c|}{} \\
-    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} \\
-    Tool & Projects & Mean & Median \\
-    \hline
-    \hline
-    SpotBugs & 3 & 0.05 & 0.06 \\
-    \hline
-    PMD & 0 & N/A & N/A \\
-    \hline
-    Infer & 6 & 0.00 & 0.00 \\
-    \hline
-  \end{tabular}
-  \caption{Java tool clean project results.}
-  \label{tab:cleanjava}
-\end{table}
-
-Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools, and Tables \ref{tab:scorejava} and \ref{tab:cleanjava} provide numeric results for projects and clean projects, respectively.
+Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools, and Table \ref{tab:scorejava} provides numeric results for projects and clean projects, respectively.
 
@@
 It may be that the diff sizes in our code were simply too small for Infer's approach.
-
 {\bf RQ2} is also answered in the affirmative.  While the raw kills for SpotBugs are not as good as for PMD, it produced fewer findings, giving it a mutant ratio approximately twice that of PMD.
@@
 For Java, again, the large majority of mutants we inspected for {\bf RQ5}, where at least one tool detected the mutant and at least one tool did not detect it were definitely meaningful.
-In particular, for Java, the large majority of mutants involved either deleted method calls or changes to conditionals (e.g., {\tt == null} to {\tt != null} that would clearly introduce potential null pointer exceptions (NPEs), and such a possible NPE was the produced finding.
+In particular, for Java, the large majority of mutants involved either deleted method calls or changes to conditionals (e.g., {\tt == null} to {\tt != null}) that would clearly introduce potential null pointer exceptions (NPEs), and such a possible NPE was the produced finding.
 While the differences between Solidity tools were often due to different detectors, the Java differences seemed mostly rooted in analysis engine methods; all tools aim to warn about potential NPEs.
@@
 \begin{figure}
-  \includegraphics[width=0.6\columnwidth]{python.png}
+  \includegraphics[width=0.4\columnwidth]{python.png}
   \caption{Mutants killed by Python static analysis tools.}
@@
     Pylint & 2.24 & 1.00 & 0.22 & 0.20 & 0.097 \\
+    Clean (0) & - & - & - & - & - \\
     \hline
     pyflakes & 2.4 & 1.00 & 0.03 & 0.00 & 0.012 \\
+    Clean (12) & - & - & 0.02 & 0.00 & - \\
     \hline
     PyChecker & 0.56& 0.00 & 0.10 & 0.00 &  0.183 \\
+    Clean (21) & - & - & 0.01 & 0.00 \\
     \hline
@@
 
-\begin{table}
-  \begin{tabular}{l|r|r|r}
-    & & \multicolumn{2}{|c|}{} \\
-    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} \\
-    Tool & Projects & Mean & Median \\
-    \hline
-    \hline
-    Pylint & 0 & N/A & N/A \\
-    \hline
-    pyflakes & 12 & 0.02 & 0.00 \\
-    \hline
-    PyChecker & 21& 0.01 & 0.00 \\
-    \hline
-  \end{tabular}
-  \caption{Python tools clean project results.}
-  \label{tab:cleanpython}
-\end{table}
-
-Figure \ref{fig:pythonvenn} shows the mutants killed by the Python analysis tools, and Tables \ref{tab:scorepython} and \ref{tab:cleanpython} provide numeric results for projects and clean projects, respectively.
+Figure \ref{fig:pythonvenn} shows the mutants killed by the Python analysis tools, and Table \ref{tab:scorepython} provides numeric results for projects and clean projects, respectively.
 
@@
 A combination of pyflakes and Pylint seems safer, though adding the relatively low-findings PyChecker seems likely to have little cost.
-For Python, we see that all of our measures are required to get a good picture of the tools.
+For Python, all of our measures are required to get a good picture.
 

### 2020-04-20 introduction.tex
ok

@@
 
-Mutation testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical} (or mutation analysis, the term we will use in this work, for reasons that will become clear) uses small syntactic changes to a program to introduce synthetic ``faults,'' under the assumption that if the original version of a program is mostly correct, most small changes will therefore introduce a fault.
+Mutation analysis\cite{jia2011analysis,demillo1978hints,budd1980theoretical} uses small syntactic changes to a program to introduce synthetic ``faults,'' under the assumption that if the original version of a program is mostly correct, such changes will often introduce a fault.
 For the most part, mutation analysis has been used to evaluate test suites by computing a mutation score, the fraction of mutants the suite detects, or ``kills''.
@@
 
-Combining a differential approach (not differential \emph{testing}, precisely, in that individual differences are not always worth inspecting) and mutation analysis, however, offers a novel way to compare static analysis tools, one useful to users wishing to select a good tool or set of tools, researchers interested in, e.g., the impact of precision/soundness trade-offs or different intermediate languages, and developers of static analysis tools hoping to improve their tools by systematic examination of code changes that are likely fault-inducing but not detected.
+Combining a differential approach and mutation analysis offers a novel way to compare static analysis tools, one useful to users wishing to select a good tool or set of tools, to researchers interested in the impact of precision/soundness trade-offs or different intermediate languages, and to developers of static analysis tools hoping to improve their tools by examining missed faults.
 
@@
 \begin{figure}
-  \includegraphics[width=0.5\columnwidth]{example.png}
+  \includegraphics[width=0.4\columnwidth]{example.png}
   \caption{Mutants killed by three static analysis tools.}
@@
 Comparing mutant results also leads to the idea of \emph{improving} static analysis tools by examining mutants detected by another tool, and thus known to be in-principle detectable, but not detected by the tool to be improved.
-Improving tools by adding detectors is useful because, even if all tools had the same set of detectors, they would not all report the same bugs; difference choices in intermediate language and tradeoffs made to avoid false positives would make the use of multiple tools with shared detectors essential for thorough analysis.
-And, of course, if one tool does have a much better underlying engine, it is beneficial to users that the ``best'' tool incorporate all detection rules.
+Improving tools by adding detectors is useful because, even if all tools had the same set of detectors, they would not all report the same bugs; different choices in intermediate language and tradeoffs made to avoid false positives may make the use of multiple tools with similar detectors essential for thorough analysis.
+And if one tool simply has a superior engine, it is beneficial to users that the ``best'' tool incorporate \emph{all} detection rules.
 
@@
 However, as with efforts to improve test suites, manually searching through all mutants can be an onerous task, especially for large-scale evaluations.
-We therefore also introduce the idea of prioritizing mutants, using a Furthest-Point-First \cite{Gonzalez} algorithm and distance metric inspired by previous work on helping developers sort through failing test cases to find distinct faults \cite{PLDI13,distMut}, to help static analysis tool developers find interesting patterns without wading through numerous uninteresting mutants not identified for similar reasons.
+We therefore also introduce the idea of prioritizing mutants, using a Furthest-Point-First \cite{Gonzalez} algorithm and distance metric inspired by previous work on helping developers sort through failing test cases to find distinct faults \cite{PLDI13,distMut}, to help static analysis tool developers find interesting patterns without wading through numerous uninteresting mutants.
 
@@
 
-\begin{itemize}
+\begin{itemize}[labelsep=3pt,leftmargin=12pt]
 \item We propose a differential approach to comparing static analysis tools based on the insight that program mutants provide an automated source of simple, easy-to-understand program changes that are likely faults.
-\item We propose a definition of mutant killing that works well in a static analysis context.
-\item We introduce a simple scheme for prioritizing mutants that helps users understand the results of analysis and guide efforts to improve, rather than simply compare, tools.
+\item We propose a definition of mutant killing for static analysis.
+\item We introduce a simple scheme for prioritizing mutants that helps users understand and use the results of analysis.
 \item We apply our method to an extensive, in-depth comparison of three Solidity smart contract analysis tools, and show how prioritization allowed us to easily identify (and build) three new detectors for the most effective of these tools.

### 2020-05-07 experiments.tex
more python details

@@
 For Python, as with Java, many mutants we inspected for {\bf RQ5} involved deletion of a call or change in comparison operator, leading to a likely method call or access to a variable containing {\tt None}, and the generated finding reported this problem.
-More Python mutants were extremely difficult to understand than for Solidity or Java, without a grasp of the larger program context, but a large number were clearly due to differences in analysis engine capabilities, since all tools aim to detect failure to guard against improper uses of {\tt None}, the Python equivalent of a Java NPE.
+Another example of a possibly engine-based mutation difference is that pyflakes did a much better job than the other tools of detecting cases where a mutant caused an exception to be unwisely ignored; these accounted for more than 10\% of differentially-relevant mutants.
+We assume the issue is engine-based in that while pyflakes did by far the best at this (real) problem, there were instances where another tool found the problem and it did not.
+On the other hand, pyflakes, in the sample we examined, did worst at detecting mutants that removed a needed argument from a function call, another common pattern (Pylint performed best for this type of mutant).
 Again, based on a qualitative examination of many mutants, we believe the results to primarily rest on realistic faults, which some tools correctly detected and others did not.

### 2020-05-08 conclusions.tex
ready to ship

@@
 analysis tools so labor-intensive.  Our approach cannot identify precision
-problems, directly, but combined with finding counts for un-mutated
-code, our approach can identify tools that detect mutants well,
-adjusted for their general tendency to flag any code as faulty, and
-thus identify imprecise tools.  We
+problems directly, but combined with finding counts for un-mutated
+code, can compute a mutant ratio that implicates imprecise tools.  We
 evaluated 9 popular static analysis tools, for Solidity smart contracts, Java,

### 2020-05-08 experiments.tex
ready to ship

@@
 \begin{figure}
-  \includegraphics[width=0.4\columnwidth]{solidity.png}
+  \includegraphics[width=0.35\columnwidth]{solidity.png}
   \caption{Mutants killed by Solidity static analysis tools.}
@@
 There are only 18 mutants detected by all tools, all of them involving replacement of {\tt msg.sender} (the caller of a smart contract, which may be another smart contract) with {\tt tx.origin} (the original initiator of a sequence of blockchain calls, a ``human'' account).
-Use of {\tt tx.origin} is usually a bad idea, can lead to incorrect behavior, and may not be relied on to be meaningful at all in future versions of Ethereum (\url{https://ethereum.stackexchange.com/questions/196/how-do-i-make-my-dapp-serenity-proof}).
+Use of {\tt tx.origin} is often (though not always) a bad idea, and can lead to incorrect behavior, so it is not surprising tools all recognize some misuses of it.
 
@@
 The most extensive is that of Durieux et al.
-\cite{durieux2019empirical}.
+\cite{durieux2019empirical}, though it unfortunately was unable to provide anything other than an implicit look at false positives, somewhat limiting its practicality.
 Slither detected 17\% of known vulnerabilities in their analysis, vs.
@@
 \begin{figure}
-  \includegraphics[width=0.4\columnwidth]{java.png}
+  \includegraphics[width=0.35\columnwidth]{java.png}
   \caption{Mutants killed by Java static analysis tools.}
@@
 \begin{figure}
-  \includegraphics[width=0.4\columnwidth]{python.png}
+  \includegraphics[width=0.35\columnwidth]{python.png}
   \caption{Mutants killed by Python static analysis tools.}

### 2020-05-08 introduction.tex
ready to ship

@@
 
-We can say that a static analysis tool kills a mutant when the \emph{number of warnings or errors}, which we call \emph{findings}, that are not informational or optimization related, i.e., that are \emph{fault-related}, produced with respect to the code \emph{increases} for the mutated version, compared to the original code.
-That is, when the tool ``finds more bugs'' for the mutated code.
+We can say that a static analysis tool kills a mutant when the \emph{number of warnings or errors}, which we call \emph{findings}, \emph{increases with mutation}.
+In order to make this definition useful, we ignore informational or optimization related warnings (e.g., if a mutant is merely \emph{stylistically} suboptimal this is not ``finding a fault'').
+That is, a mutant is killed when a tool ``finds more bugs'' for the mutated code than for the un-mutated code.
 This difference may be most easily interpreted when the original code produces no findings; we call such code \emph{clean}.
@@
 \begin{figure}
-  \includegraphics[width=0.4\columnwidth]{example.png}
+  \includegraphics[width=0.35\columnwidth]{example.png}
   \caption{Mutants killed by three static analysis tools.}

### 2020-05-08 relatedwork.tex
ready to ship

--- 2020-01-26 relatedwork.tex
@@
 This is a result of both algorithmic constraints and basic engineering trade-offs.
-While comparisons of static analysis tools \cite{CompareJavaTools,durieux2019empirical, Parizi,slither} have appeared in the literature for years, these generally involved large human effort and resulting smaller scale, or did not make a strong effort to address false positives, or restricted analysis to, e.g., a known defects set \cite{AllBugs}.
+While comparisons of static analysis tools \cite{CompareJavaTools,durieux2019empirical, Parizi,slither,pashchenko2017delta} have appeared in the literature for years, these generally involved large human effort and resulting smaller scale, did not make a strong effort to address false positives, or restricted analysis to, e.g., a known defects set \cite{AllBugs, do2016toward}.
+Known defect sets are particularly vulnerable to tools intentionally overfitting/gaming the benchmark; our approach makes it easy to compare tools on ``fresh'' code to avoid this risk.
+Compared to well-known studies of Java tools \cite{AllBugs,CompareJavaTools} our approach used a larger set of subject programs (1.8 MLOC total vs.
+170-350KLOC) and thousands of \emph{detected} (and tens of thousands of potential) faults, vs.
+e.g., about 500 known defects \cite{AllBugs}.
+Results not using defect sets are even more limited in that humans can only realistically examine a few dozens of each type of warning \cite{CompareJavaTools}.
 
@@
 
-Finally, the large body of work on using mutants in software testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical, groce2015verified,groce2018verified,MutGoogle,ivankovic2018industrial,mutKernel} is obviously relevant, in that we basically adopt its approach, but re-define killing a mutation for a static analysis context.
+Finally, we essentially adopt the approach of the large body of work on using mutants in software testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical, groce2015verified,groce2018verified,MutGoogle,ivankovic2018industrial,mutKernel}, but re-define killing a mutation for a static analysis context.

### 2020-08-03 experiments.tex
move files

@@
 \item {\bf RQ4:}  Do tools detect more mutants in programs for which they produce no warnings, initially?
-\item {\bf RQ5:}  Are mutants distinguishing tools usually flagged due to real faults, where the finding is related to the introduced fault; that is, are our results usually \emph{meaningful}?  
+\item {\bf RQ5:}  Are mutants distinguishing tools usually flagged due to real faults, where the finding is related to the introduced fault; that is, are our results usually \emph{meaningful}, based on inspection of mutant code and static analysis warnings produced?
 \item {\bf RQ6:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that

### 2020-08-04 abstract.tex
first pass

--- 2020-01-24 abstract.tex
@@
 Understanding the strengths and weaknesses of tools, and performing direct comparisons of their effectiveness is difficult; it usually involves either manual examination of differing warnings on real code, or the bias-prone construction of artificial test cases.
-In practice, comparisons tend to be limited to superficial, anecdotal discussions in the informal literature (e.g., blog posts by software developers), or purely research-community-oriented evaluations made by the authors of new tools seeking to publish their results.
+%In practice, comparisons tend to be limited to superficial, anecdotal discussions in the informal literature (e.g., blog posts by software developers), or %purely research-community-oriented evaluations made by the authors of new tools seeking to publish their results.
 This paper proposes a novel automated approach to comparing static analysis tools, based on producing \emph{mutants} of real code, and comparing mutation detection rates for tools to their warning rates on the original code.
 In addition to making tool differences quantitatively observable without extensive manual effort, this approach offers a new way to detect and fix omissions in a static analysis tool's set of detectors.
-We present an extensive comparison of three well-known Solidity smart contract static analysis tools, and show how using an automatic prioritization of our results allowed us to add three new detectors to the best of the tools.
+We present an extensive comparison of three well-known Solidity smart contract static analysis tools, and show how using an automatic prioritization of our results allowed us to add three effective new detectors to the best of these.
 We also evaluate popular Java and Python static analysis tools and discuss their strengths and weaknesses.

### 2020-08-04 experiments.tex
first pass

@@
 \begin{figure}
+  \centering
   \includegraphics[width=0.35\columnwidth]{solidity.png}
@@
 
-\noindent \begin{code}
- 0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.mutant.275.sol:
+\noindent
+\begin{verbatim}
   if (!p.recipient.send(p.amount)) \{  // Make the payment
@@
   if (true) \{  // Make the payment
-      \end{code}
-
+\end{verbatim}
 }
@@
  {\bf Mutant showing Type-based tautologies.}
-\begin{code}
- 0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.mutant.311.sol:
- require(nextDiscountTTMTokenId6 >= 361 \&\& nextDiscountTTMTokenId6 <= 391);
+\begin{verbatim}
+  require(nextDiscountTTMTokenId6 >= 361 \&\& ...);
   ==>  ...361...==>...0...
- require(nextDiscountTTMTokenId6 >= 0 \&\& nextDiscountTTMTokenId6 <= 391);
-      \end{code}
+  require(nextDiscountTTMTokenId6 >= 0 \&\& ...);
+\end{verbatim}
 }
@@
       {\bf Mutant showing Loss of precision.}        
-\begin{code}
- 0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.mutant.480.sol:
+\begin{verbatim}
   byte char = byte(bytes32(uint(x) * 2 ** (8 * j)));
@@
   byte char = byte(bytes32(uint(x) * 2 ** (8 / j)));
-      \end{code}
-
+\end{verbatim}
 }
@@
 \begin{figure}
+  \centering
   \includegraphics[width=0.35\columnwidth]{java.png}
@@
 \begin{figure}
+  \centering
   \includegraphics[width=0.35\columnwidth]{python.png}

### 2020-08-04 introduction.tex
first pass

@@
 
-Static analysis of code is one of the most effective ways to avoid defects in software, and, when security is a concern, is simply essential.
+Static analysis of code is one of the most effective ways to avoid defects in software, and, when security is a concern, is essential.
 Static analysis can find problems that are extremely hard to detect by testing, when the inputs triggering a bug are hard to find.
-Static analysis is also often more efficient than testing; a bug that takes a fuzzer like AFL days to find may be immediately identified by a good static analysis tool.
-
-Users of static analysis tools often wonder which of multiple tools available for a programming language are most effective, and, when using more than one tool is an option, which is often the case, how much tools overlap in their results.
-Given the human effort required to read static analysis results, the latter can be an important question.
-If two tools find substantially different bugs, it is important to use both tools \cite{AllBugs}.
-On the other hand, if two tools are very similar in what they detect, the effort to configure and use both, and especially of wading through duplicate results, may not be a good use of time.
-Developers of static analysis tools also want to be able to compare their tools to others targeting the same domain, in order to see what detection methods or tweaks to precision/soundness trade-offs they might want to imitate.
-Unfortunately, comparing static analysis tools is hard, and would seem to require vast manual effort to inspect findings and determine ground truth on a scale that would provide statistical confidence.
+Static analysis is also often more efficient than testing; a bug that takes a fuzzer days to find may be immediately identified.
+Users of static analysis tools often wonder which of multiple tools available for a language are most effective, and how much tools overlap in their results.
+Tools often find substantially different bugs, making it important to use multiple tools \cite{AllBugs}.
+However, given the high cost of examing results, if a tool provides only marginal novelty, it may not be worth using, especially if it has a high false-positive rate.
+Developers of static analysis tools also want to be able to compare their tools to other tools, in order to see what detection patterns or precision/soundness trade-offs they might want to imitate.
+Unfortunately, comparing static analysis tools in these ways is hard, and would seem to require vast manual effort to inspect findings and determine ground truth on a scale that would provide statistical confidence.
 
 Differential testing \cite{Differential,ICSEDiff,csmith} is a popular approach to comparing multiple software systems offering similar functionality, but the wide divergence of possible trade-offs, analysis focuses, and the prevalence of false positives in almost all analysis results makes na\"ive differential testing not applicable to static analysis tools \cite{regehrRandom}.
-
 Mutation analysis\cite{jia2011analysis,demillo1978hints,budd1980theoretical} uses small syntactic changes to a program to introduce synthetic ``faults,'' under the assumption that if the original version of a program is mostly correct, such changes will often introduce a fault.
 For the most part, mutation analysis has been used to evaluate test suites by computing a mutation score, the fraction of mutants the suite detects, or ``kills''.
-Most such use has been in research efforts, rather than practical testing efforts, though there has been sporadic use by interested developers.
 Groce et al.
-\cite{groce2015verified,groce2018verified} proposed examining individual mutants that survive a rigorous testing and verification effort to detect and correct weaknesses in such efforts.
-The approach was able to expose bugs in a heavily-tested module of the Linux kernel \cite{mutKernel} and a widely used Python file system.
-Recently, mutation analysis has been adopted in industrial settings, though not for actual examination of all surviving mutants \cite{MutGoogle,ivankovic2018industrial}, due to scale.
+\cite{groce2015verified,groce2018verified} proposed examining individual mutants that survive a rigorous testing and verification effort to detect and correct weaknesses in testing, and found bugs in a heavily-tested module of the Linux kernel \cite{mutKernel} and a widely used Python file system.
+Recently, mutation analysis has been adopted in industrial settings, though not for actual examination of all surviving mutants \cite{MutGoogle,ivankovic2018industrial}.
 
@@
 
-\subsection{Differential Mutation Analysis}
+%\subsection{Differential Mutation Analysis}
 
@@
 This difference may be most easily interpreted when the original code produces no findings; we call such code \emph{clean}.
-For non-clean code, a tool conceivably could detect the mutant, but only change a previously generated finding, not add an additional finding, leading to an underestimate of effectiveness.
+For non-clean code, a tool conceivably could detect the mutant, but only change a previously generated finding, not add an additional finding.
 However, even for non-clean code, \emph{most detected mutants should produce a new warning}.
@@
 The tool cannot simply flag all code as having a problem or it will perform poorly at the task of \emph{distinguishing} the mutated code from non-mutated, and presumably at least \emph{more} correct, code.
-Based on verification and testing uses of mutation, it is safe to say that usually at minimum 40\%, often 50-60\%, and frequently up to 80\%+ \cite{mutKernel,groce2018verified,le2014mucheck}, of mutants are not semantically equivalent to the original code \cite{TCE,impactEquiv,smith2009should}, so the task presented to a static analysis tool is simply the core functionality of static analysis \emph{to distinguish faulty from correct code without execution}.
+Based on verification and testing uses of mutation, it is safe to say that usually at minimum 40\%, often 50-60\%, and frequently up to 80\%+ \cite{mutKernel,groce2018verified,le2014mucheck}, of mutants are not semantically equivalent to the original code \cite{TCE,impactEquiv,smith2009should}, so the task presented to a static analysis tool is simply the core functionality of static analysis: \emph{to distinguish faulty from correct code without execution}.
 Obviously, many faults cannot be identified statically without a complete specification, or without unreasonable analysis cost and precision, but the measure of performance here is \emph{relative} to other tools applied to the same code; this is primarily a \emph{differential} approach.
-In other words, a key notion is that while most mutants cannot be detected statically, the ones that are tend to be \emph{true positives}: \emph{if} they were real code changes, they would be faults.
-We manually confirmed that for most mutants in our experiments where any tool detected the mutant (and at least one tool did not), the changes were indeed ones that would be real faults if present in the code.
+In other words, a key notion is that while many mutants cannot be detected statically, the ones that are tend to be \emph{true positives}: \emph{if} they were real code changes, they would be faults.
+\emph{We manually confirmed that for a large portion of the mutants in our experiments where any tool detected the mutant, the changes were indeed ones that would be real faults if present in the code.}
 
@@
 If one tool emits more warnings and errors than another, it may mean that the tool is more effective at finding bugs; but it may also mean that it has a higher false positive rate.
-Without human examination of the individual findings, it is impossible to be sure, or even (in cases where the tools are reasonably comparable) to make an informed guess.
+Without human examination of the individual findings, it is impossible to be sure.
 Using mutants, however, provides a foreground to compare to this background.
 In particular, for a large set of programs, the most informative result will be when 1) tool A reports fewer findings on average than tool B over the un-mutated programs but 2) tool A also detects more mutants.
-This is strong evidence that A is simply better all-around than B; it likely has a lower false positive rate \emph{and} a lower false negative rate.
-While it is not proof of this claim, it is hard to construct another plausible explanation for reporting \emph{fewer} findings on un-mutated code while still detecting \emph{more} mutants.
-Other than having better precision and recall, how else could a tool effectively distinguish mutated from un-mutated code?
-
-We can quantitatively express the relationship between detecting mutants and findings for the original program.
-Since we are comparing over the same program or set of programs, we simply divide the mean \emph{mutation score} ($\frac{|\mathit{killed}|}{|\mathit{mutants}|}$, the ratio of killed mutants to all mutants) by the mean number of findings.
-This \emph {mutant ratio} tells us about the ability of a tool to produce findings \emph{for mutants}, relative to its tendency to produce findings in general, (per line of code, per source file, etc.).
-If a tool has a tendency to produce large numbers of findings compared to other tools, and this is paired with a tendency to detect most mutants, then the tool will not be penalized for producing many findings.
-Assuming that real faults are relatively rare in the original, un-mutated code, the best result and best mutant ratio will be for a tool that produces comparatively few findings for un-mutated code, but detects a larger portion of mutants than other tools; the worst result will be a tool that produces lots of findings, but detects few mutants.
-We will actually see some examples of the worst case in our results for real tools.
+This is strong evidence that A is simply better all-around than B; it likely has a lower false positive rate \emph{and} a lower false negative rate, since it is hard to construct another plausible explanation for reporting \emph{fewer} findings on un-mutated code while still detecting \emph{more} mutants.
+Our method (see the \emph{mutant ratio} defined below) provides a quantitative measure of this insight.
 
 \begin{figure}
+  \centering
   \includegraphics[width=0.35\columnwidth]{example.png}
@@
 
-Finally, even when tools have similar quantitative results, including similar ratios, examining individual mutants killed by one tool but not by another allows us to understand strengths and weaknesses of the tools, in a context that makes comprehending the cause of the detection, or lack of it, easy: the difference between the un-mutated code and mutated code will always be small and relatively simple.
-Moreover, simply looking at how much two tools agree on mutants can answer the question of a user of static analysis tools: given that I am using tool A, would adding tool B be likely to add enough new, interesting results to make it worth my time to examine its output?
-When one tool subsumes another in terms of mutants, it can be very clear that the tool whose killed mutants are all killed by another tool is likely strictly inferior.
+Finally, even when tools have similar quantitative results, examining individual mutants killed by one tool but not by another allows us to understand strengths and weaknesses of the tools, in a helpful context: the difference between the un-mutated code and mutated code will always be small and simple.
+Moreover, simply looking at how much two tools agree on mutants can answer the question: given that I am using tool A, would adding tool B be likely be worthwhile?
 Interested users, e.g.
 security analysts, can inspect the differences to get an idea of the particular cases when a tool might be most effective, but a more typical user can simply look at a Venn diagram of kills like that shown in Figure \ref{fig:examplevenn}.
-Assume that tools A, B, and C all produce very similar numbers of findings for un-mutated code, and have similar execution times.
+Consider hypothetical tools A, B, and C.
+A and B produce similar numbers of findings on the code in question, while tool C produces an order of magnitude more findings.
 Tool A is likely the most important tool to make use of; it detected more mutants than any other tool, and more than twice as many mutants were killed by A alone than by B alone.
-However, also running tool B is probably a very good idea, assuming, e.g., it is not a very expensive commercial tool.
+However, also running tool B is well justifed.
 B does not do as well as A, but it is the only tool that detects a large number of mutants, and most mutants it detects are unique to it.
-Finally, Tool C \emph{may} not be worth running; recall that it produces a similar number of findings to A and B on the un-mutated code, so it is notably bad at detecting faults, at least ones that look like mutants.
-It might be a good idea to just look at the 27 mutants detected by C alone:  if they represent an important class of potential problems (perhaps C specialized in detecting potentially non-terminating loops), then C might be useful, but if the first few mutants inspected are false positives, then C is likely not useful.
+Finally, Tool C \emph{may} not be worth running, since its poor performance on mutants but high finding rate suggests it may be prone to missing bugs and to false positives.
+It might be a good idea to just look at the 27 mutants detected by C alone:  \emph{if} they represent an important class of potential problems (perhaps C specialized in detecting potentially non-terminating loops), then C might be useful, but if the first few mutants inspected are false positives, then C is likely not useful.
 
-Comparing mutant results also leads to the idea of \emph{improving} static analysis tools by examining mutants detected by another tool, and thus known to be in-principle detectable, but not detected by the tool to be improved.
+Comparing mutant results also leads to the idea of \emph{improving} static analysis tools by examining mutants detected by another tool, and thus known to be in-principle detectable.
 Improving tools by adding detectors is useful because, even if all tools had the same set of detectors, they would not all report the same bugs; different choices in intermediate language and tradeoffs made to avoid false positives may make the use of multiple tools with similar detectors essential for thorough analysis.
 And if one tool simply has a superior engine, it is beneficial to users that the ``best'' tool incorporate \emph{all} detection rules.
-
-Of course, any faults in code, not just mutants, could serve this purpose.
-But, again, the automatic nature of mutation generation, and the presumption that the mutation is indeed a fault, is useful.
-Moreover, because mutants follow syntactic patterns, searching for similar mutants/faults the tool to be improved does not detect is much easier than with arbitrary faults, and can be partly automated.
 However, as with efforts to improve test suites, manually searching through all mutants can be an onerous task, especially for large-scale evaluations.
-We therefore also introduce the idea of prioritizing mutants, using a Furthest-Point-First \cite{Gonzalez} algorithm and distance metric inspired by previous work on helping developers sort through failing test cases to find distinct faults \cite{PLDI13,distMut}, to help static analysis tool developers find interesting patterns without wading through numerous uninteresting mutants.
+We therefore introduce the idea of \emph{prioritizing} mutants.
 
@@
 The vast majority of applied mutation operations for all of our experiments were highly generic, and do not plausibly represent a pattern in which some tool might specialize.
-Statement deletions leave no ``trace'' for a tool to match against, but only an omission.
+Code deletions, the most common kind of mutation by far, leave no ``trace'' for a tool to match against, but only an omission, so cannot be subject to this concern.
 Changing arithmetic and comparison operators and numeric constants (incrementing, decrementing, or changing to 0 or 1) account for most of the non-statement-deletion mutants.
 We cannot see how any tool could have a useful detection rule that looks for these kinds of changes: a tool obviously could not flag all instances of 0 and 1 constants, or of addition, and remain a useful tool.
-Finally, many mutants, though far fewer than in the previously noted categories, add {\tt break} or {\tt continue} statements.
-Again, we do not see how a tool could take advantage of this without either genuinely understanding likely-bad control flow patterns, in which case, we think it ``deserves'' to kill more mutants, or flagging so many legitimate uses of {\tt break} and {\tt continue} in un-mutated code that it would have a bad mutant ratio, due to producing more warnings than other tools.
 
@@
 Consider the code in Figure \ref{fig:sol424intro} \cite{solintro}.
-The Universal Mutator tool \cite{universalmutator,regexpMut}, which has been extensively tuned for Solidity's grammar (though not to target any particular vulnerabilities), and is the only smart contract mutation tool referenced in the Solidity documentation (\url{https://solidity.readthedocs.io/en/v0.5.12/resources.html}), produces seven valid, non-redundant (by Trivial Compiler Equivalence \cite{TCE}) mutants for this trivial example code.
+The Universal Mutator tool \cite{universalmutator,regexpMut}, which has been extensively tuned for Solidity's grammar (though not to target any particular vulnerabilities), and is the only smart contract mutation tool referenced in the Solidity documentation (\url{https://solidity.readthedocs.io/en/v0.5.12/resources.html}), produces seven valid, non-redundant (by Trivial Compiler Equivalence \cite{TCE}) mutants for it.
 Both the public version of Slither \cite{slither} and SmartCheck \cite{smartcheck} produce a small number (three and two, respectively) of low-severity, informational, warnings for this code.
@@
 If we restrict our analysis to findings with a severity greater than informational, SmartCheck detects no mutants of the contract, while Slither still reports that some mutants \emph{allow an arbitrary caller to cause the contract to destroy itself}.
-This simple example shows how our approach works on a small scale.
-Using large numbers of larger, more realistic contracts makes it possible to extract the same kind of information, on a much larger scale.
-Prioritization of mutants is not very useful here (ranking three mutants saves little effort), so we will show the utility of that idea in our full Solidity results.
 
@@
 
-The Solidity, Java, and Python case studies also serve to answer a set of research questions investigating the utility of our approach, and provide basic evidence that it provides actionable, non-obvious information that is otherwise difficult to produce.
  While there are certainly limitations to using differential mutation analysis to compare and extend static analysis tools, the method scales to basing comparisons on large numbers of real software source files and very large numbers of faults, but still has some of the advantages of humans establishing ground truth for tool findings.
+
+\section{Method}
+
+The proposed approach is simple in outline:
+
+\begin{enumerate}
+\item Run each tool on the unmutated source code target(s), and determine the \emph{baseline}: the number of (non-informational/stylistic) findings produced.
+\item Generate mutants of the source code and run each tool on each mutant.
+Consider a mutant killed if the number of findings for the mutated code is greater than the number for the baseline, un-mutated code.
+\item Compute, for each tool, the \emph{mutant ratio}:  the mutation score ($\frac{|\mathit{killed}|}{|\mathit{mutants}|}$) divided by the mean number of baseline findings.
+\item (Optional): Discard all mutants not killed by at least one tool and all mutants killed by all tools.  What remains allows \emph{differential} analysis.
+\item (Optional): Examine mutants in the difference in \emph{prioritized} order to qualitatively understand tool differences.
+\end{enumerate}
+
+The most important step here is the computation of the \emph{mutant ratio}, which tells us about the ability of a tool to produce findings \emph{for mutants}, relative to its tendency to produce findings in general.
+If a tool has a tendency to produce large numbers of findings compared to other tools, and this is paired with a tendency to detect more mutants as well, then the tool will not be penalized for producing many findings.
+Assuming that real faults are relatively rare in the original, un-mutated code, the best result and best (highest) mutant ratio will be for a tool that produces comparatively few findings for un-mutated code, but detects a larger portion of mutants than other tools; the worst result will be a tool that produces lots of findings, but detects few mutants.
+We will actually see some examples of this worst case.
+
+\subsection{Prioritizing Mutants}
+\label{sec:prioritizing}
+
+One goal of our approach is to make it easy for tool developers to examine cases where one tool kills a mutant and another fails to, in order to identify patterns for new detectors or analysis algorithm problems.
+Dedicated developers may also simply want to scan all mutants their tool does not kill, for the same purpose, analogous to what Groce et al.
+have proposed for automated verification and testing \cite{groce2015verified,groce2018verified}.
+Security analysts and other expert users who are not developers may also wish to do this, to better understand tool strengths and weaknesses.
+
+Unfortunately the full list of unkilled mutants, or differentially
+unkilled mutants, is likely to be both large and highly redundant.  In
+our results below, only one of 9 tools we examined killed fewer than
+1,000 mutants it alone detected.  Any cross-tool comparison is thus
+likely to involve hundreds or thousands of mutants.
+
+The problem of identifying unique ``faults'' (tool weaknesses) in this
+situation is very similar to the \emph{fuzzer taming} problem in
+software testing, as defined by Chen et al. \cite{PLDI13}: ``Given a
+potentially large collection of test cases, each of which triggers a
+bug, rank them in such a way that test cases triggering distinct bugs
+are early in the list.'' \cite{PLDI13}.  Their solution was
+to use Gonzalez' Furthest-Point-First \cite{Gonzalez} (FPF) algorithm
+to \emph{rank} test cases so that users can examine very different
+test cases as quickly as possible.  An
+FPF ranking requires a distance metric $d$, and ranks items so that
+dissimilar ones appear earlier.
+The hypothesis of Chen et al. was that dissimilar tests, by a
+well-chosen metric, will also fail due to different faults.  FPF is a
+greedy algorithm that proceeds by repeatedly adding the item with the
+\emph{maximum minimum distance to all previously ranked items}. Given an
+initial seed item $r_0$, a set $S$ of items to rank, and a distance
+metric $d$, FPF computes $r_i$ as
+$s \in S: \forall s' \in S: min_{ j < i}(d(s,r_j)) \geq min_{j <
+  i}(d(s',r_j))$.  The condition on $s$ is obviously true when
+$s = s'$, or when $s' = r_j$ for some $j < i$; the other cases for
+$s'$ force selection of \emph{some}
+max-min-distance $s$.
+
+In order to apply FPF ranking to examining mutants, we implemented a
+simple, somewhat \emph{ad hoc} distance metric and FPF ranker.  Our metric $d$ is
+the sum of a set of measurements.  First, it adds a similarity
+ratio based on Levenshtein distance \cite{lev} for (1) the \emph{changes} (Levenshtein edits) from
+the original source code elements to
+the two mutants,  (2) the two original source code elements changed (in
+general, lines), and (3) the actual output mutant code.  These are
+weighted with multipliers of 5.0, 0.1, and 0.1, respectively; the type
+of change (mutation operator, roughly) dominates this part of the
+distance, because it best describes ``what the mutant did''; however,
+because many mutants will have the same change (e.g., changing {\tt +}
+to {\tt -}, the other ratios also often matter.
+Our metric also incorporates a measure of the distance in the source
+code between the locations of two mutants.  If the mutants are to
+different files, this adds 0.5 to the distance; it also adds 0.25
+times the number of source lines separating the two mutants if they
+are in the same file, divided by 10, but caps the amount added at
+0.25.
+
+We do not claim this is an optimal, or even tuned, metric. Devising a better metric is left as future work, we only wish to show 
+that even a hastily-devised and somewhat arbitrary metric provides 
+considerable advantage over wading through an un-ordered list of 
+mutants, and introduce the idea of using FPF for mutants, not just for tests: FPF is useful for failures in general, however discovered.
+
+%; in the long
+%run, it even may be completely replaced.  However, the metric we use
+%provided useful results for a different problem, ranking mutants in an
+%effort similar to that of Groce et
+%al. \cite{groce2015verified,groce2018verified} to examine unkilled
+%mutants for a smart contract library tested using the Echidna fuzzer \cite{echidna-code},
+%and we simply adopted it for our differential mutation analysis.

### 2020-08-04 experiments.tex
icse draft 1 ready

@@
 \item {\bf RQ1:}  Does mutation analysis of static analysis tools produce actionable results?  That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar?
-\item {\bf RQ2:}  Does our approach provide additional information beyond simply counting findings for the original, un-mutated analyzed code?  Do \emph{ratios} differ between static analysis tools?
-\item {\bf RQ3:}  Do the rankings that raw kills and ratios establish agree with other sources of information about the effectiveness of the evaluated tools?  Can we independently confirm our results?
+\item {\bf RQ2:}  Does our approach provide additional information beyond simply counting findings for the original, un-mutated analyzed code?  Do \emph{ratios} differ between tools?
+\item {\bf RQ3:}  Do the rankings that raw kills and ratios establish agree with other sources of information about the effectiveness of the evaluated tools? 
 \item {\bf RQ4:}  Do tools detect more mutants in programs for which they produce no warnings, initially?
@@
 showed that the ``clean program assumption'' for testing, the idea that coverage will be similar for faulty and fixed versions of a program, is a threat to the validity of investigations of the relationship between coverage and fault detection \cite{CleanProgram}, and we want to establish that this is unlikely to be the case for our approach.
-Our answer to {\bf RQ5} is somewhat inherently qualitative and incomplete; we cannot analyze all mutants on which results are based manually, and understanding the mutants and tool warnings completely would require deep understanding of all the software systems investigated.
+Our answer to {\bf RQ5} is somewhat inherently qualitative and incomplete; we cannot analyze all mutants on which results are based manually, and understanding the mutants and tool warnings completely would require deep understanding of all the subject programs.
 However, in many cases, the impact of a mutant is clear, and the reason for warnings is obvious.
 This was often enough the case that, as we discuss below, we are confident mutants that distinguish tools are meaningful (missed) opportunities for static analysis tools.
-For {\bf RQ6}  we have only preliminary answers, for the smart contract tools.
+For {\bf RQ6}  we have only preliminary answers, for smart contract tools.
 
@@
 SmartCheck, developed by SmartDec, translates Solidity source directly to an XML-based representation, then uses \emph{XPath} patterns to define problems.
-Securify, from SRI Systems Lab at ETH Zurich, works at the bytecode level, first parsing and decompiling contracts, then translating to \emph{semantic facts} in order to look for predefined problems.
+Securify, from SRI Systems Lab at ETH Zurich, works at the bytecode level, first parsing and decompiling contracts, then translating to \emph{semantic facts} in order to look for problems.
 
@@
 {\bf RQ2} is also answered in the affirmative.
-Counting findings for un-mutated code might suggest that Securify is the best tool, by a wide margin, but in the context of its near-zero mutant ratio, we must suspect that many of the warnings are false positives.
+Counting findings for un-mutated code might suggest that Securify is the best tool, by a wide margin, but in the context of its near-zero mutant ratio, we must suspect (and we partially manually confirmed) that many of the warnings are false positives.
 Slither has the best mutant ratio, but the margin between it and SmartCheck confirms that both tools likely provide value.
@@
 We therefore, for smart contracts, find no threat to our approach arising from the presence of findings on un-mutated code.
-We speculate that ``clean'' results for some tools result from contracts where the tool has trouble with the contract code, but does not actually crash; Slither may do better on clean code because it has fewer such failures, and clean contracts are probably generally simpler and easier to analyze.
+We speculate that ``clean'' results for some tools result from contracts where the tool has trouble with the contract code, but does not actually crash; Slither may do better on clean code because it has fewer such failures, and clean contracts are probably somewhat easier to analyze.
 
@@
 Combining statement deletion mutants with other mutants only moved the mutants we used down to 3rd, 11th, and 14th positions.
-By default we rank statement deletions separately, since such mutants are usually easier to understand and either note as important or dismiss, and in testing (but not static analysis) they are likely to be the most critical faults not detected.
+By default we rank statement deletions separately, since such mutants are usually easier to understand and evaluate, and in testing (but not static analysis) they are likely to be the most critical faults not detected.
 
@@
 SmartCheck and Securify tend to identify most (though not all) uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule, intended to reduce false positives.
-It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ6}.
+It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ6}.
 
@@
 It may be that the diff sizes in our code were simply too small for Infer's approach.
-{\bf RQ2} is also answered in the affirmative.  While the raw kills for SpotBugs are not as good as for PMD, it produced fewer findings, giving it a mutant ratio approximately twice that of PMD.
+{\bf RQ2} is also answered in the affirmative.  While the raw kills for SpotBugs are not as good as for PMD, it had fewer findings, giving it a mutant ratio approximately twice that of PMD.
 

### 2020-08-04 introduction.tex
icse draft 1 ready

@@
 
-\section{Method}
+\section{Differential Mutation Analysis}
 
@@
 Consider a mutant killed if the number of findings for the mutated code is greater than the number for the baseline, un-mutated code.
-\item Compute, for each tool, the \emph{mutant ratio}:  the mutation score ($\frac{|\mathit{killed}|}{|\mathit{mutants}|}$) divided by the mean number of baseline findings.
+\item Compute, for each tool, the \emph{mutant ratio}:  the mutation score ($\frac{|\mathit{killed}|}{|\mathit{mutants}|}$) divided by (mean) baseline.
 \item (Optional): Discard all mutants not killed by at least one tool and all mutants killed by all tools.  What remains allows \emph{differential} analysis.
-\item (Optional): Examine mutants in the difference in \emph{prioritized} order to qualitatively understand tool differences.
+Examine the remaining mutants in the difference in \emph{prioritized} order to qualitatively understand tool differences, or improve a tool.
 \end{enumerate}

### 2020-08-04 relatedwork.tex
icse draft 1 ready

@@
 
-The goal of ``analysing the program analyser'' \cite{cadar2016analysing} and applying better automated methods to evaluate and improve analysis tools has become recently more popular and, we suspect, more possible.
-The irony of using mostly ad-hoc, manual methods to test and understand static analysis tools is apparent; however, the fundamentally incomplete and heuristic nature of effective analysis tools makes this a challenge similar to testing machine learning algorithms \cite{OnlyOracle}; most tools will not produce ``the right answer'' all the time, by their very nature.
-This is a result of both algorithmic constraints and basic engineering trade-offs.
+The goal of ``analysing the program analyser'' \cite{cadar2016analysing} is intuitively attractive.
+The irony of using mostly ad-hoc, manual methods to test and understand static analysis tools is apparent; however, the fundamentally incomplete and heuristic nature of such tools makes this a challenge similar to testing machine learning algorithms \cite{OnlyOracle}; most tools will not produce ``the right answer'' all the time, as a result of both algorithmic constraints and basic engineering trade-offs.
 While comparisons of static analysis tools \cite{CompareJavaTools,durieux2019empirical, Parizi,slither,pashchenko2017delta} have appeared in the literature for years, these generally involved large human effort and resulting smaller scale, did not make a strong effort to address false positives, or restricted analysis to, e.g., a known defects set \cite{AllBugs, do2016toward}.
-Known defect sets are particularly vulnerable to tools intentionally overfitting/gaming the benchmark; our approach makes it easy to compare tools on ``fresh'' code to avoid this risk.
+Defect sets are vulnerable to tools intentionally overfitting/gaming the benchmark; our approach makes it easy to compare tools on ``fresh'' code to avoid this risk.
 Compared to well-known studies of Java tools \cite{AllBugs,CompareJavaTools} our approach used a larger set of subject programs (1.8 MLOC total vs.
-170-350KLOC) and thousands of \emph{detected} (and tens of thousands of potential) faults, vs.
+170-350KLOC) and thousands of \emph{detected} faults, vs.
 e.g., about 500 known defects \cite{AllBugs}.
@@
 \cite{regehrRandom} proposed the generation of random programs (\emph{\'a la} Csmith \cite{csmith}) to test analysis tools aiming for soundness, in limited circumstances, but noted that na\"ive differential testing of analysis tools was not possible.
-This paper proposes a non-na\"ive differential comparison (not, exactly, differential testing, however, in that only aggregate results are possible to interpret without human intelligence), based on the observation that the ability to detect program mutants offers an automatable way to tell which of two tools is better (for a given universe of examples, at least) at telling faulty from non-faulty code.
+This paper proposes a non-na\"ive differential comparison based on the observation that the ability to detect program mutants offers an automatable comparison basis.
+We essentially adopt the approach of the large body of work on using mutants in software testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical, groce2015verified,groce2018verified,MutGoogle,ivankovic2018industrial,mutKernel}, but re-define killing a mutation for a static analysis context.
 
@@
 propose a different approach to differential testing of analysis tools \cite{klinger2019differentially}.
-Their approach is in some ways similar to ours, in that it takes as input a set of seed programs, and compares results across new versions generated from that seed.
-The primary differences are that their seed programs must be warning-free (which greatly limits the set of input programs available) and their tool must parse and understand the programs, and that the new versions are based on adding new assertions, not ``breaking'' the original code.
-We allow arbitrarily buggy seed programs (thus many more real programs can be used), and can, due to the any-language nature of the mutation generator we use, operate even in new languages without further development effort.
-Further, their approach only identifies problems when tools are outliers compared to numerous other tools in either detecting a bug (precision) or not detecting it (soundness), and so requires comparing multiple tools.
-Our approach has some utility for even a single tool (you can just examine prioritized un-detected mutants).
+Their approach is in some ways similar to ours, in that it takes as input a set of seed programs, and compares results across new versions generated from seeds.
+The primary differences are that their seed programs must be warning-free (which greatly limits the set of input programs available), and that the new versions are based on adding new assertions only.
+We allow arbitrarily buggy seed programs, and can, due to the any-language nature of the mutation generator we use, operate even in new languages.
 On the other hand, their approach can identify precision issues, while we offer no real help with false positives (in theory, you could apply their majority-vote method to mutants only a few tools flag, but mutants \emph{are} usually faults.
-in contrast to their introduction of checks that may be guaranteed to pass, so this is probably not very helpful).
-Most importantly, however, their approach only applies to tools that check assertions, rather than more general, popular static analysis tools that only identify bad code patterns.
-
-Finally, we essentially adopt the approach of the large body of work on using mutants in software testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical, groce2015verified,groce2018verified,MutGoogle,ivankovic2018industrial,mutKernel}, but re-define killing a mutation for a static analysis context.
+in contrast to their introduction of checks that may be guaranteed to pass).
+Most importantly, however, their approach \emph{only applies to tools that check assertions}, rather than the more common case of tools that identify bad code patterns (e.g.
+all tools we considered in this paper).

### 2020-08-04 conclusions.tex
operators too

@@
 for real faults, to compare (and motivate improvements to) static
-analysis tools.  Mutants are attractive in that changing a mostly
-correct program usually introduces a bug into it; this is the basis of
-mutation testing, after all, and a large body of work supports the
+analysis tools.  Mutants are attractive in that a large body of work supports the
 claim that at least 60-70\% of mutants are fault-inducing.   This
-means we can assume mutants are faulty, and escape the
+allows us to assume \emph{detected} mutants are faulty, and escape the
 ground-truth/false positive problem that makes comparing static
-analysis tools so labor-intensive.  Our approach cannot identify precision
-problems directly, but combined with finding counts for un-mutated
-code, can compute a mutant ratio that implicates imprecise tools.  We
+analysis tools so labor-intensive.   We
 evaluated 9 popular static analysis tools, for Solidity smart contracts, Java,

### 2020-08-04 experiments.tex
operators too

@@
 For Java, PIT \cite{pitmut} is more popular, but does not produce source-level mutants, needed for PMD and for manual inspection of results.
+Universal Mutator includes a large set of mutation operators, some unconventional (e.g., swapping order of function arguments) but based on real-world bugs; the complete set is described by regular expressions at \url{https://github.com/agroce/universalmutator/tree/master/universalmutator/static}.
+However, the large majority of generated compiling mutants that were detected came from a small set of  commonly-used operators \cite{mutant}, particularly 1) statement/code deletion and 2) operator, conditional, and constant replacements.
+
 We used our results to answer a set of research questions:
@@
 
-Because examining all mutants manually is infeasible,  for {\bf RQ5} we focused primarily on examining mutants in the \emph{difference}: mutants detected by at least one tool, but not detected by all tools, the only ones that actually influence the comparison of tools.
+Following the method proposed in Section \ref{sec:method},  for {\bf RQ5} we focused on examining mutants detected by at least one tool, but not detected by all tools, the only ones that actually influence the comparison of tools.
 The vast majority of the mutants in this set were meaningful semantic changes a static analysis tool could be expected to detect, and the findings produced by tools were relevant to the nature of the fault.
-We do not believe that all mutants represent definite faults; some are harmless but unusual code changes.
-A large number of the instances where use of {\tt tx.origin} in place of {\tt msg.sender} was flagged seem to us to be strange, but not necessarily incorrect, code.
-On the other hand, at the risk of false positives, it is not unreasonable for tools to report such strange code.
+We do not believe that \emph{all} mutants represent definite faults; some are harmless but unusual code changes.
+Many cases where use of {\tt tx.origin} in place of {\tt msg.sender} was flagged seem to us to be strange, but not necessarily incorrect, code.
+On the other hand, it is not at all unreasonable for tools to report such notably strange code.
 Our estimate is that, ignoring {\tt tx.origin} cases, \emph{at least 70\% of the mutants detected by one, but not all, tools, represent realistic bugs}, and failure to detect is roughly equally due to missing detectors and imprecise analysis.
@@
 
-For Java, again, the large majority of mutants we inspected for {\bf RQ5}, where at least one tool detected the mutant and at least one tool did not detect it were definitely meaningful.
-In particular, for Java, the large majority of mutants involved either deleted method calls or changes to conditionals (e.g., {\tt == null} to {\tt != null}) that would clearly introduce potential null pointer exceptions (NPEs), and such a possible NPE was the produced finding.
+For Java, again, the large majority ($>$ 75\%) of randomly chosen mutants in the tool difference set we inspected for {\bf RQ5} were definitely meaningful, essentially ``real faults.''  In particular, for Java, the large majority of mutants involved either deleted method calls or changes to conditionals (e.g., {\tt == null} to {\tt != null}) that would clearly introduce potential null pointer exceptions (NPEs), and such a possible NPE was the produced finding.
 While the differences between Solidity tools were often due to different detectors, the Java differences seemed mostly rooted in analysis engine methods; all tools aim to warn about potential NPEs.
@@
 These ranged in size from 3,377 LOC to 34,671 LOC, with a total size of just under 100 KLOC.
-For Python, we were able to produce 167,511 valid mutants to analyze.
+For Python, we were able to produce 167,511 valid mutants.
 
@@
 
-For Python, as with Java, many mutants we inspected for {\bf RQ5} involved deletion of a call or change in comparison operator, leading to a likely method call or access to a variable containing {\tt None}, and the generated finding reported this problem.
-Another example of a possibly engine-based mutation difference is that pyflakes did a much better job than the other tools of detecting cases where a mutant caused an exception to be unwisely ignored; these accounted for more than 10\% of differentially-relevant mutants.
-We assume the issue is engine-based in that while pyflakes did by far the best at this (real) problem, there were instances where another tool found the problem and it did not.
+For Python, as with Java, many mutants we inspected for {\bf RQ5} involved deletion of a call or change in comparison operator, leading to a likely method call or access to a variable containing {\tt None}, and the generated finding reported just this problem.
+Another example of a possibly engine-based mutation difference is that pyflakes did a much better job than the other tools of detecting cases where a mutant caused an exception to be unwisely ignored; these accounted for more than 10\% of differentially-relevant mutants; these may not all be faults, but are certainly worth inspecting.
+We assume the issue is engine-based in that while pyflakes did by far the best at this (real) problem, there were instances where another tool found the same notional problem and it did not.
 On the other hand, pyflakes, in the sample we examined, did worst at detecting mutants that removed a needed argument from a function call, another common pattern (Pylint performed best for this type of mutant).
-Again, based on a qualitative examination of many mutants, we believe the results to primarily rest on realistic faults, which some tools correctly detected and others did not.
+Again, based on a qualitative examination of many mutants, we believe the results to primarily ($>$ 75\%) rest on realistic faults, which some tools correctly detected and others did not.
 

### 2020-08-05 introduction.tex
ok

@@
 That is, a mutant is killed when a tool ``finds more bugs'' for the mutated code than for the un-mutated code.
-This difference may be most easily interpreted when the original code produces no findings; we call such code \emph{clean}.
+This difference may be most easily interpreted when the original code produces no findings; we call such code \emph{clean} (by analogy with Chekam et al.'s notion \cite{CleanProgram}).
 For non-clean code, a tool conceivably could detect the mutant, but only change a previously generated finding, not add an additional finding.
@@
  \section{Differential Mutation Analysis}
+ \label{sec:method}
 

### 2020-08-19 conclusions.tex
update iftekhar

@@
 software testing, as well \cite{groce2018verified}.
-We are also interested in combining static and traditional mutation analysis: it might be useful to consider a mutant killed if it is detected either by static or dynamic analysis, using static kills to lower the number of required costly testing/verification runs for mutants, and the combined result as a useful guide to the quality of overall \emph{defense in depth} \cite{antunes2012defending,FC20} for high-importance code.
+
+%We are also interested in combining static and traditional mutation analysis: it might be useful to consider a mutant killed if it is detected either by static or dynamic analysis, using static kills to lower the number of required costly testing/verification runs for mutants, and the combined result as a useful guide to the quality of overall \emph{defense in depth} \cite{antunes2012defending,FC20} for high-importance code.

### 2020-08-19 introduction.tex
update iftekhar

@@
 In order to make this definition useful, we ignore informational or optimization related warnings (e.g., if a mutant is merely \emph{stylistically} suboptimal this is not ``finding a fault'').
-That is, a mutant is killed when a tool ``finds more bugs'' for the mutated code than for the un-mutated code.
+That is, a mutant is killed when a tool ``finds more (unique) bugs'' for the mutated code than for the un-mutated code.
 This difference may be most easily interpreted when the original code produces no findings; we call such code \emph{clean} (by analogy with Chekam et al.'s notion \cite{CleanProgram}).
@@
 Obviously, many faults cannot be identified statically without a complete specification, or without unreasonable analysis cost and precision, but the measure of performance here is \emph{relative} to other tools applied to the same code; this is primarily a \emph{differential} approach.
-In other words, a key notion is that while many mutants cannot be detected statically, the ones that are tend to be \emph{true positives}: \emph{if} they were real code changes, they would be faults.
-\emph{We manually confirmed that for a large portion of the mutants in our experiments where any tool detected the mutant, the changes were indeed ones that would be real faults if present in the code.}
+While many mutants cannot be detected statically, the ones that \emph{are} tend to be \emph{true positives}: \emph{if} they were real code changes, they would be faults.
+\emph{We manually confirmed that for a large portion of the detected mutants in our experiments, the changes were indeed ones that would be real faults if present in the code.}
 
@@
 
-Comparing mutant results also leads to the idea of \emph{improving} static analysis tools by examining mutants detected by another tool, and thus known to be in-principle detectable.
-Improving tools by adding detectors is useful because, even if all tools had the same set of detectors, they would not all report the same bugs; different choices in intermediate language and tradeoffs made to avoid false positives may make the use of multiple tools with similar detectors essential for thorough analysis.
-And if one tool simply has a superior engine, it is beneficial to users that the ``best'' tool incorporate \emph{all} detection rules.
-However, as with efforts to improve test suites, manually searching through all mutants can be an onerous task, especially for large-scale evaluations.
-We therefore introduce the idea of \emph{prioritizing} mutants.
-
-A general objection to our approach is that mutants may differ substantially from ``real'' faults, in some way.
-This is certainly true, in a sense \cite{GopinathMutants}, but for static analysis purposes we believe it does not matter.
-The real risk is that some mutation operators align with patterns a particular tool identifies, biasing the evaluation in favor of that tool.
-Such faults may be dis-proportionately present in mutants vs.
-real code.
-However, we consider this unlikely.
-The vast majority of applied mutation operations for all of our experiments were highly generic, and do not plausibly represent a pattern in which some tool might specialize.
-Code deletions, the most common kind of mutation by far, leave no ``trace'' for a tool to match against, but only an omission, so cannot be subject to this concern.
-Changing arithmetic and comparison operators and numeric constants (incrementing, decrementing, or changing to 0 or 1) account for most of the non-statement-deletion mutants.
-We cannot see how any tool could have a useful detection rule that looks for these kinds of changes: a tool obviously could not flag all instances of 0 and 1 constants, or of addition, and remain a useful tool.
-
-\subsection{A Simple Example}
-
 \begin{figure}
@@
 
-Consider the code in Figure \ref{fig:sol424intro} \cite{solintro}.
+More concretely, consider the code in Figure \ref{fig:sol424intro} \cite{solintro}.
 The Universal Mutator tool \cite{universalmutator,regexpMut}, which has been extensively tuned for Solidity's grammar (though not to target any particular vulnerabilities), and is the only smart contract mutation tool referenced in the Solidity documentation (\url{https://solidity.readthedocs.io/en/v0.5.12/resources.html}), produces seven valid, non-redundant (by Trivial Compiler Equivalence \cite{TCE}) mutants for it.
-Both the public version of Slither \cite{slither} and SmartCheck \cite{smartcheck} produce a small number (three and two, respectively) of low-severity, informational, warnings for this code.
-Both tools also detect (by increasing the number of warnings produced) four of the seven mutants.
+Both the public version of Slither \cite{slither} and SmartCheck \cite{smartcheck} (two popular smart contract static analysis tools) produce a small number (three and two, respectively) of low-severity, informational, warnings for this code.
+Both tools also detect four of the seven mutants (here the number of warnings increases, and the additional warnings are clearly driven by the mutation change) .
 However, only one of the mutants detected is common to both tools: both tools detect changing the {\tt return} statement in the {\tt get} function to a call to {\tt selfdestruct} the smart contract, deleting it.
@@
 If we restrict our analysis to findings with a severity greater than informational, SmartCheck detects no mutants of the contract, while Slither still reports that some mutants \emph{allow an arbitrary caller to cause the contract to destroy itself}.
+Given that both tools, ignoring informational results, detect no problems with the original code, and only Slither detects any problems with the mutated code, we can say that Slither performs better for static analysis of mutants for this contract.
 
-\subsection{Contributions}
+Comparing mutant results also leads to the idea of \emph{improving} static analysis tools by examining mutants detected by another tool, and thus known to be in-principle detectable.
+Improving tools by adding detectors is useful because, even if all tools had the same set of detectors, they would not all report the same bugs; different choices in intermediate language and tradeoffs made to avoid false positives may make the use of multiple tools with similar detectors essential for thorough analysis.
+And if one tool simply has a superior engine, it is beneficial to users that the ``best'' tool incorporate \emph{all} detection rules.
+However, as with efforts to improve test suites, manually searching through all mutants can be an onerous task, especially for large-scale evaluations.
+We therefore introduce the idea of \emph{prioritizing} mutants to make it easier to inspect \emph{different} weaknesses in tools.
+
+A general objection to our approach is that mutants may differ substantially from ``real'' faults, in some way.
+This is certainly true, in a sense \cite{GopinathMutants}, but for static analysis purposes we believe it does not matter.
+The real risk is that some mutation operators align with patterns a particular tool identifies, biasing the evaluation in favor of that tool.
+Such faults may be dis-proportionately present in mutants vs.
+real code.
+However, we consider this unlikely.
+The vast majority of applied mutation operations for all of our experiments were highly generic, and do not plausibly represent a pattern in which some tool might specialize.
+Code deletions, the most common kind of mutation by far, leave no ``trace'' for a tool to match against, but only an omission, so cannot be subject to this concern.
+Changing arithmetic and comparison operators and numeric constants (incrementing, decrementing, or changing to 0 or 1) account for most of the non-statement-deletion mutants.
+It is difficult to imagine how any useful tool could unfairly identify these kinds of mutants.
+
+%\subsection{Contributions}
 
@@
 \item Compute, for each tool, the \emph{mutant ratio}:  the mutation score ($\frac{|\mathit{killed}|}{|\mathit{mutants}|}$) divided by (mean) baseline.
+If the (mean) baseline is zero, use a baseline equal to either one or the lowest non-zero baseline for any tool in the comparison set\footnote{Unless the number of programs used for evaluation is small, a zero mean baseline for a tool is extremely unlikely to arise in practice.}.
 \item (Optional): Discard all mutants not killed by at least one tool and all mutants killed by all tools.  What remains allows \emph{differential} analysis.

### 2020-11-21 experiments.tex
updates

@@
 We also note that while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick (it took about 6 days sequential compute time for Slither and 3 days for SmartCheck, i.e., about 5-15 seconds per mutant for both tools), Securify often required many hours to analyze a mutant, and frequently required a few days to analyze a mutant; the full analysis required over three months of compute time.
+However, similar statistical results would almost certainly be produced by running as few as 1,000 mutants, randomly selected, due to the implications of Tchebysheff’s inequality \cite{gopinath2015hard}, a technique that should improve scalability for all such analyses.
+
+If we consider results at the individual contract level, the overall picture is in some ways even clearer.
+Slither detected more mutants than Securify for 84 of the contracts, and more than SmartCheck for 85 of the contracts.
+SmartCheck detected more mutants than Securify for 83 of the contracts.
+Comparing ratios instead, Slither was better than Security for 97 contracts, but better than SmartCheck for only 75 of the contracts.
+SmartCheck's ratio was better than Securify's for 95 contracts.
+The standard deviation of contract raw scores was 0.05 for Slither, 0.03 for SmartCheck, and 0.04 for Securify.
 
@@
 Counting findings for un-mutated code might suggest that Securify is the best tool, by a wide margin, but in the context of its near-zero mutant ratio, we must suspect (and we partially manually confirmed) that many of the warnings are false positives.
-Slither has the best mutant ratio, but the margin between it and SmartCheck confirms that both tools likely provide value.
+Slither has the best mutant ratio, but the margin between it and SmartCheck confirms that both tools likely provide value; we note there is a 25\% chance that SmartCheck has a better ratio than Slither for an individual contract.
 
@@
 Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools, and Table \ref{tab:scorejava} provides numeric results for projects and clean projects, respectively.
+At the individual project level, Infer was never best; PMD had a better raw score than SpotBugs for 10 of the 15 projects, but SpotBug had a better ratio for 11 of the projects.
+There is perhaps a real tradeoff in verbosity and precision here.
+Standard deviation in project mutation scores was 0.05 for SpotBugs, 0.04 for PMD, and 0 for Infer.
 
@@
 Figure \ref{fig:pythonvenn} shows the mutants killed by the Python analysis tools, and Table \ref{tab:scorepython} provides numeric results for projects and clean projects, respectively.
+At the project level, Pylint was better than pyflakes and PyChecker for 22 of 25 projects each by raw score, and better by ratio for 23 and 22 projects, respectively.
+PyChecker was better than pyflakes for 15 projects, by both raw and ratio measures.
+Standard deviations in mutation scores were high for Python:  0.15 for Pylint, 0.06 for pyflakes, and 0.27 for PyChecker.
 

### 2020-11-21 analyzejava.py

@@ -122,6 +122,10 @@ with open("Java.csv") as javaResults:
     print()
     print(len(files), "FILES")
     print()
+
+    toolIssues = {}
+    toolScores = {}    
+    
     for tool in ["spotbugs", "PMD", "Infer"]:
         print()
         print("="*48)
@@ -142,13 +146,38 @@ with open("Java.csv") as javaResults:
                 if (files[f]["spotbugs_issues"] == 0) and (files[f]["PMD_issues"] == 0) and (files[f]["Infer_issues"] == 0):
                     allCleanCount += 1
                     allCleanScores.append(score)
-        print("ISSUES:", scipy.mean(issues), scipy.median(issues))
-        print("SCORES:", scipy.mean(scores), scipy.median(scores))
+        print("ISSUES:", scipy.mean(issues), scipy.median(issues), scipy.std(issues))
+        toolIssues[tool] = issues
+        toolScores[tool] = scores
+        print("SCORES:", scipy.mean(scores), scipy.median(scores), scipy.std(scores))
         print("RATIO:", scipy.mean(scores)/scipy.mean(issues))
         print("CLEAN:", cleanCount)
         print("CLEAN SCORES:", scipy.mean(cleanScores), scipy.median(cleanScores))
         print("#ALL CLEAN:", allCleanCount),
         print("ALL CLEAN SCORES:", scipy.mean(allCleanScores), scipy.median(allCleanScores))
+
+print()
+print()
+        
+for tool1 in ["spotbugs", "PMD", "Infer"]:
+    for tool2 in ["spotbugs", "PMD", "Infer"]:
+        if tool1 > tool2:
+            print(tool1, "VS", tool2)
+            diffAmt = []
+            for i in range(0, len(toolScores[tool1])):
+                diffAmt.append(toolScores[tool1][i]-toolScores[tool2][i])
+            print("RAW:", tool1,"BETTER THAN",tool2,len(list(filter(lambda x: x > 0, diffAmt))))
+            diffAmt = []
+            for i in range(0, len(toolScores[tool1])):
+                t1Score = toolScores[tool1][i] / max(1,toolIssues[tool1][i])
+                t2Score = toolScores[tool2][i] / max(1,toolIssues[tool2][i])                
+                diffAmt.append(t1Score-t2Score)
+            print("RATIO:", tool1,"BETTER THAN",tool2,len(list(filter(lambda x: x > 0, diffAmt))))
[…]

### 2020-11-21 analyzepython.py

@@ -79,6 +79,9 @@ with open("Python.csv") as javaResults:
     print("Pylint but not Pychecker:", Pylint_not_Pychecker)
     print("Pychecker but not Pylint:", Pychecker_not_Pylint)    
 
+    toolIssues = {}
+    toolScores = {}    
+    
     print()
     print()
     print(len(files), "FILES")
@@ -104,12 +107,31 @@ with open("Python.csv") as javaResults:
                     allCleanCount += 1
                     allCleanScores.append(score)
         print("ISSUES:", scipy.mean(issues), scipy.median(issues))
-        print("SCORES:", scipy.mean(scores), scipy.median(scores))
+        toolIssues[tool] = issues
+        toolScores[tool] = scores        
+        print("SCORES:", scipy.mean(scores), scipy.median(scores), scipy.std(scores))
         print("RATIO:", scipy.mean(scores)/scipy.mean(issues))
         print("CLEAN:", cleanCount)
         print("CLEAN SCORES:", scipy.mean(cleanScores), scipy.median(cleanScores))
         print("#ALL CLEAN:", allCleanCount),
         print("ALL CLEAN SCORES:", scipy.mean(allCleanScores), scipy.median(allCleanScores))
 
+for tool1 in ["Pylint", "Pychecker", "Pyflakes"]:
+    for tool2 in ["Pylint", "Pychecker", "Pyflakes"]:
+        if tool1 > tool2:
+            print(tool1, "VS", tool2)
+            diffAmt = []
+            for i in range(0, len(toolScores[tool1])):
+                diffAmt.append(toolScores[tool1][i]-toolScores[tool2][i])
+            print("RAW:", tool1,"BETTER THAN",tool2,len(list(filter(lambda x: x > 0, diffAmt))))
+            diffAmt = []
+            for i in range(0, len(toolScores[tool1])):
+                t1Score = toolScores[tool1][i] / max(1,toolIssues[tool1][i])
+                t2Score = toolScores[tool2][i] / max(1,toolIssues[tool2][i])                
+                diffAmt.append(t1Score-t2Score)
+            print("RATIO:", tool1,"BETTER THAN",tool2,len(list(filter(lambda x: x > 0, diffAmt))))
+            print(len(toolScores[tool1]))
+            print("="*40)
+        
 print(projects, len(projects))
 print(Mutant-1)

### 2020-11-21 randjava.py

@@ -1,0 +1,20 @@
+import csv
+import scipy
+import random
+
+data = []
+
+with open("Java.csv") as javaResults:
+    reader = csv.DictReader(javaResults)
+    for row in reader:
+        if (row["spotbugs_killed"] == "1") != (row["PMD_killed"]=="1"):
+            data.append((row["changed line"], row["diff"]))
+
+random.shuffle(data)
+
+for i in range(0,50):
+    print(data[i][0])
+    print()
+    print(data[i][1])
+    print()
+    print("="*40)

### 2020-11-21 randpy.py

@@ -1,0 +1,22 @@
+import csv
+import scipy
+import random
+
+data = []
+
+with open("Rebuttalresult_Twitter-Trends_NEW.csv") as javaResults:
+    reader = csv.DictReader(javaResults)
+    for row in reader:
+        if (((row["Pyflakeskilled"] == "1") != (row["Pycheckerkilled"]=="1")) or
+                ((row["Pyflakeskilled"] == "1") != (row["PylintKilled"]=="1")) or
+                ((row["Pycheckerkilled"] == "1") != (row["PylintKilled"]=="1"))):
+                
+            data.append((row["before"], row["after"]))
+
+random.shuffle(data)
+
+for i in range(0,50):
+    print(data[i][0].replace("\n",""))
+    print("  ==>")
+    print(data[i][1].replace("\n",""))
+    print("="*40)

### 2021-02-01 experiments.tex
fse format
version info for solidity

@@
 
-We analyzed three well-known tools for static analysis of Solidity smart contracts: Slither \cite{slither}, SmartCheck \cite{smartcheck}, and Securify \cite{Securify}.
+We analyzed three well-known tools for static analysis of Solidity smart contracts: Slither \cite{slither}\footnote{pip version 0.6.1}), SmartCheck \cite{smartcheck}\footnote{github commit 17fb98d9759eebeca1f92d48be52b85a1f5d8b89}, and Securify \cite{Securify}\footnote{github commit 0ffde61b9527be70b147d3ab687f11993d99212b}.
 Slither, based on an SSA-based intermediate language (SlithIR \cite{slither}) is an open-source tool from Trail of Bits.
@@
   \end{tabular}
-  \caption{Python tool results over all projects.}
+  \caption{Python tool results over all projects (old).}
+  \label{tab:scorepython}
+\end{table}
+
+\begin{table}
+  \begin{tabular}{l|r|r|r|r|r}
+    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
+    Tool & Mean & Median & Mean & Median & Ratio\\
+    \hline
+    \hline
+    Pylint & 11.87 & 4.00 & 0.20 & 0.19 & 0.02 \\
+    Clean (3) & - & - & 0.09 & 0.11 & - \\
+    \hline
+    pyflakes & 1.52 & 0.00 & 0.08 & 0.06 & 0.05 \\
+    Clean (15) & - & - & 0.09 & 0.08 & - \\
+    \hline
+    MyPy & 3.00 & 3.00 & 0.05 & 0.03 &  0.02 \\
+    Clean (3) & - & - & 0.03 & 0.01 & - \\
+    \hline
+  \end{tabular}
+  \caption{Python tool results over all projects (new).}
   \label{tab:scorepython}

### 2021-02-19 experiments.tex
Python new results incorporated

@@
 These ranged in size from 3,377 LOC to 34,671 LOC, with a total size of just under 100 KLOC.
-For Python, we were able to produce 167,511 valid mutants.
+We analyzed 158,418 valid mutants.
 
@@
   \centering
-  \includegraphics[width=0.35\columnwidth]{python.png}
+  \includegraphics[width=0.35\columnwidth]{python_new.png}
   \caption{Mutants killed by Python static analysis tools.}
@@
 \hline
-    Pylint & 2.24 & 1.00 & 0.22 & 0.20 & 0.097 \\
-    Clean (0) & - & - & - & - & - \\
+Pylint & 10.63 & 5.50 & 0.31 & 0.34 & 0.03 \\
+Clean (4) & - & - & 0.49 & 0.47 & - \\
 \hline
-    pyflakes & 2.4 & 1.00 & 0.03 & 0.00 & 0.012 \\
-    Clean (12) & - & - & 0.02 & 0.00 & - \\
+pyflakes & 1.46 & 0.00 & 0.14 & 0.12 & 0.09 \\
+Clean (14) & - & - & 0.14 & 0.12 & - \\
 \hline
-    PyChecker & 0.56& 0.00 & 0.10 & 0.00 &  0.183 \\
-    Clean (21) & - & - & 0.01 & 0.00 \\
+PyChecker & 11.6 & 0.00 & 0.01 & 0.00 & 0.00 \\
+Clean (4) & - & - & 0.00 & 0.00 & - \\
 \hline
 \end{tabular}
-  \caption{Python tool results over all projects (old).}
-  \label{tab:scorepython}
-\end{table}
-
-\begin{table}
-  \begin{tabular}{l|r|r|r|r|r}
-    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
-    Tool & Mean & Median & Mean & Median & Ratio\\
-    \hline
-    \hline
-    Pylint & 11.87 & 4.00 & 0.20 & 0.19 & 0.02 \\
-    Clean (3) & - & - & 0.09 & 0.11 & - \\
-    \hline
-    pyflakes & 1.52 & 0.00 & 0.08 & 0.06 & 0.05 \\
-    Clean (15) & - & - & 0.09 & 0.08 & - \\
-    \hline
-    MyPy & 3.00 & 3.00 & 0.05 & 0.03 &  0.02 \\
-    Clean (3) & - & - & 0.03 & 0.01 & - \\
-    \hline
-  \end{tabular}
-  \caption{Python tool results over all projects (new).}
+\caption{Python tool results over all projects.}
 \label{tab:scorepython}
@@
 Figure \ref{fig:pythonvenn} shows the mutants killed by the Python analysis tools, and Table \ref{tab:scorepython} provides numeric results for projects and clean projects, respectively.
-At the project level, Pylint was better than pyflakes and PyChecker for 22 of 25 projects each by raw score, and better by ratio for 23 and 22 projects, respectively.
-PyChecker was better than pyflakes for 15 projects, by both raw and ratio measures.
-Standard deviations in mutation scores were high for Python:  0.15 for Pylint, 0.06 for pyflakes, and 0.27 for PyChecker.
+At the project level, Pylint was better than pyflakes and PyChecker for 21 and 24 projects, respectively, by raw score; pyflakes was better than Pylint and PyChecker for 3 and 24 projects, respectively; PyChecker was never better than another tool by raw score.
+Switching to ratio measures, Pylint was better than pyflakes and PyChecker for 11 and 23 projects, respectively; pyflakes was better than Pylint and PyChecker for 13 and 24 projects, respectively.
+PyChecker was better than Pylint for one project, by ratio.
+Standard deviations in mutation scores were sometimes high for Python:  0.17 for Pylint, 0.08 for pyflakes, and 0.01 for PyChecker.
 
-For {\bf RQ1}, Pylint and pyflakes both killed more than 15,000 mutants killed by no other tool; Pylint killed more mutants, but the relative difference was much smaller than in our Solidity or Java analysis.
-It is clear both tools are well worth using, assuming these are not false positives.
-Pyflakes' mutation score is lower than unique kills might suggest because it has almost no overlap with the other two tools.
-It is doing something different.
+For {\bf RQ1}, there is a clear difference between tools.
+Pylint uniquely kills almost an order of magnitude more mutants than the next-best tool.
+There is a large overlap between Pylint and pyflakes, while PyChecker is both much less effective and doing something fairly different than the other tools.
+From the diagram, one might think that pyflakes acts, to some extent, as a less verbose ``subset'' of Pylint, in that most mutants detected by pyflakes are also detected by Pylint (however, the nearly 3,500 killed mutants unique to pyflakes suggest it is a useful tool, perhaps most useful after problems also reported by Pylint are fixed).
+PyChecker performs poorly in part because it crashed (due to changes in Python since the last update to the tool) for 19 of the 25 projects; however, as the clean scores show, it also performed poorly on programs where it worked.
 
-The mutant ratios {\bf RQ2} suggest that PyChecker might be better than pyflakes, in fact, and better than Pylint, because while it killed fewer mutants, it also produced fewer findings by far, overall.
-PyChecker would be a good choice for quick and dirty analysis of code on the fly, perhaps, in that it had excellent mutation scores for some programs (hence a high mean) but is relatively ``quiet'' compared to the other tools.
-However, its very low median mutation score is troubling; it may simply not work for some code.
-A combination of pyflakes and Pylint seems safer, though adding the relatively low-findings PyChecker seems likely to have little cost.
-For Python, all of our measures are required to get a good picture.
+The mutant ratios {\bf RQ2} show that pyflakes is more competitive than is obvious from raw kill comparisons.
+The mutant ratio is almost three times as good as for Pylint!
+That is, some of Pylint's advantage may be due to general verbosity, even once stylistic warnings are turned off.
+Combining the results from raw kills and ratios, a strategy of using pyflakes as a quick check for problems, then using both pyflakes and Pylint for more in-depth analysis makes sense.
+Whether to use both pyflakes and Pylint in CI is a rquestion of tolerance for handling false positives, but it is clear that the ``price'' of additional warnings from Pylint is (1) high but (2) not without ereturn on investment (in terms of additional real finds).
+PyChecker is probably too outdated, and sometimes too verbose when it works, to be useful.
 
@@
 There is probably general agreement that PyChecker is old and maybe less useful, but also terse and sometimes helpful.
-Pylint is the most recommended tool, and the general complaint that it is too picky was mitigated in our results by turning off clearly informative-only findings; users not configuring Pylint will probably see even more killed mutants, but likely a worse mutant ratio because mutants tend to introduce non-stylistic problems.
-Pyflakes is also well liked, and supposedly less verbose than Pylint; we did not see this in our results, but that may be because we configured Pylint, and pyflakes aims to focus on ``actual bugs'' so it is equally verbose for mutants, which are likely to be actual bugs.
-For {\bf RQ4} there were no clean projects for the known-to-be-aggressive Pylint.
-Performance for the other two tools was worse on projects for which they were clean, again failing to support the possibility that our definition of killing a mutant is too restrictive.
+Pylint is the most recommended tool, and the general complaint that it is too picky was somewhat mitigated in our results by turning off warnings that are clearly purely stylistic.
+Pyflakes is also well liked, and is considered much less verbose than Pylint; this was definitely reflected in our results, where Pyflakes underperformed in raw kills, but had a much better mutant ratio than Pylint.
+For {\bf RQ4}, Pylint performed significantly better on clean projects.
+Performance for the other two tools was essentially unchanged over projects for which they were clean, again failing to strongly support the possibility that our definition of killing a mutant is too restrictive.
 
-For Python, as with Java, many mutants we inspected for {\bf RQ5} involved deletion of a call or change in comparison operator, leading to a likely method call or access to a variable containing {\tt None}, and the generated finding reported just this problem.
-Another example of a possibly engine-based mutation difference is that pyflakes did a much better job than the other tools of detecting cases where a mutant caused an exception to be unwisely ignored; these accounted for more than 10\% of differentially-relevant mutants; these may not all be faults, but are certainly worth inspecting.
-We assume the issue is engine-based in that while pyflakes did by far the best at this (real) problem, there were instances where another tool found the same notional problem and it did not.
-On the other hand, pyflakes, in the sample we examined, did worst at detecting mutants that removed a needed argument from a function call, another common pattern (Pylint performed best for this type of mutant).
-Again, based on a qualitative examination of many mutants, we believe the results to primarily ($>$ 75\%) rest on realistic faults, which some tools correctly detected and others did not.
+For Python, all but three of the mutants we examined for {\bf RQ 5} (a random sample of 100 mutants with a kill difference) involved code changes we agreed would introduce a real bug in the code if it was executed.
+The large majority involved statement deletions, detected via 1) unused variables/arguments, 2) instances lacking a member field, or 3) undefined variables.
+Interestingly, this seems to be an engine issue more than a detector issue, as pyflakes and Pylint both basically support all of these kinds of checks.
+In some cases the two tools both detected a problem (but PyChecker did not) but differed as to \emph{which} variable was not used or defined, again suggesting an engine rather than detection rule difference.
+Pylint's better performance was mostly, in the sample, due to detecting more of these issues, though it also was the only tool in the sample that detected arithmetic operation changes, due as far as we could determine to constant index changes.
+Pylint also detected a few mutants no other tool did, due to the presence of unreachable code.
+In one case, it also noticed a protected member access via a change in constant index, a surprisingly complex problem to find, in our opinion.
+PyChecker was never the sole detecting tool for any mutants in our sample.
 

### 2021-02-20 experiments.tex
python loc data
clarify no bugs

@@
 For Python, we analyzed the top 25 GitHub projects by our criteria (see above), due to the smaller size of Python projects.
-These ranged in size from 3,377 LOC to 34,671 LOC, with a total size of just under 100 KLOC.
-We analyzed 158,418 valid mutants.
+These ranged in size from 137 LOC to 29,339 LOC, with a total size of about 75 KLOC, and a mean size of 3,185 LOC.
+We analyzed 158,418 valid, non-TCE redundant, mutants taken from these programs.
 
@@
 
-For Python, all but three of the mutants we examined for {\bf RQ 5} (a random sample of 100 mutants with a kill difference) involved code changes we agreed would introduce a real bug in the code if it was executed.
+For Python, all but three of the mutants we examined for {\bf RQ 5} (a random sample of 100 mutants with a kill difference) involved code changes we agreed were definitely buggy, and would cause incorrect behavior if executed (the exceptions were deletions of code with no effect on state, e.g., strings as comments).
 The large majority involved statement deletions, detected via 1) unused variables/arguments, 2) instances lacking a member field, or 3) undefined variables.

### 2021-02-28 experiments.tex
market cap

@@
 Security and correctness are thus critical in the smart contract domain, and static analysis is a key way to ensure allocation of high-value resources is not compromised.
-The most popular smart contract platform, by far, is the Ethereum blockchain, and the Solidity smart contract language \cite{buterin2013whitepaper,wood2014yellow}; the Ethereum cryptocurrency has a market capitalization as we write of over \$15 billion dollars, largely fueled by interest in the smart contract functionality.
+The most popular smart contract platform, by far, is the Ethereum blockchain, and the Solidity smart contract language \cite{buterin2013whitepaper,wood2014yellow}; the Ethereum cryptocurrency has a market capitalization as we write of over \$100 billion dollars, largely fueled by interest in the smart contract functionality.
 Ethereum contracts have been the targets of widely publicized attacks, with large financial consequences  \cite{spank,DAO}.

### 2021-04-20 experiments.tex
target ISSRE

--- 2021-02-19 experiments.tex
@@
 Universal Mutator includes a large set of mutation operators, some unconventional (e.g., swapping order of function arguments) but based on real-world bugs; the complete set is described by regular expressions at \url{https://github.com/agroce/universalmutator/tree/master/universalmutator/static}.
-However, the large majority of generated compiling mutants that were detected came from a small set of  commonly-used operators \cite{mutant}, particularly 1) statement/code deletion and 2) operator, conditional, and constant replacements.
+However, most mutants that were detected came from a small set of  commonly-used operators \cite{mutant}, particularly 1) code deletion and 2) operator, conditional, and constant replacements.
 
@@
 \item {\bf RQ4:}  Do tools detect more mutants in programs for which they produce no warnings, initially?
-\item {\bf RQ5:}  Are mutants distinguishing tools usually flagged due to real faults, where the finding is related to the introduced fault; that is, are our results usually \emph{meaningful}, based on inspection of mutant code and static analysis warnings produced?
+\item {\bf RQ5:}  Are mutants distinguishing tools usually flagged due to real faults, where the finding is related to the introduced fault; that is, are our results usually \emph{meaningful}?
 \item {\bf RQ6:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
@@
 {\bf RQ4} addresses a concern briefly mentioned in the introduction:  it is possible that warnings for the original code interfere with our definition of detection.
-The ideal case for our approach is when a tool report no findings for un-mutated code, and either reports a finding (kills the mutant) or not (does not kill it) when the mutant is introduced.
+The ideal case for our approach is when a tool reports no findings for un-mutated code, and reports a finding when the mutant is introduced.
 Chekam et al.
-showed that the ``clean program assumption'' for testing, the idea that coverage will be similar for faulty and fixed versions of a program, is a threat to the validity of investigations of the relationship between coverage and fault detection \cite{CleanProgram}, and we want to establish that this is unlikely to be the case for our approach.
+showed that the ``clean program assumption'' for testing is a threat to the validity of investigations of the relationship between coverage and fault detection \cite{CleanProgram}, but we show that this is unlikely to be the case for our approach.
 Our answer to {\bf RQ5} is somewhat inherently qualitative and incomplete; we cannot analyze all mutants on which results are based manually, and understanding the mutants and tool warnings completely would require deep understanding of all the subject programs.
@@
 This was often enough the case that, as we discuss below, we are confident mutants that distinguish tools are meaningful (missed) opportunities for static analysis tools.
-For {\bf RQ6}  we have only preliminary answers, for smart contract tools.
+For {\bf RQ6}  we have only a preliminary answer.
 
@@
 
-We analyzed three well-known tools for static analysis of Solidity smart contracts: Slither \cite{slither}\footnote{pip version 0.6.1}), SmartCheck \cite{smartcheck}\footnote{github commit 17fb98d9759eebeca1f92d48be52b85a1f5d8b89}, and Securify \cite{Securify}\footnote{github commit 0ffde61b9527be70b147d3ab687f11993d99212b}.
+%We analyzed three well-known tools for static analysis of Solidity smart contracts: Slither \cite{slither}\footnote{pip version 0.6.1}), SmartCheck \cite{smartcheck}\footnote{github commit 17fb98d9759eebeca1f92d48be52b85a1f5d8b89}, and Securify \cite{Securify}\footnote{github commit 0ffde61b9527be70b147d3ab687f11993d99212b}.
+
+We analyzed three well-known tools for static analysis of Solidity smart contracts: Slither \cite{slither}, SmartCheck \cite{smartcheck}, and Securify \cite{Securify}.
 Slither, based on an SSA-based intermediate language (SlithIR \cite{slither}) is an open-source tool from Trail of Bits.
@@
 Constant-valued conditionals tend to indicate debugging efforts that have persisted into production code, or other faults; there are almost no circumstances where a conditional should not vary with state or input.
-This detector is actually split into two detectors, one for this serious issue, and an informational/stylistic detector that notes that code such as {\tt if (x == true)}, while semantically harmless, is difficult to read.
-This problem was easily identified from cases such as the mutant in Figure \ref{fig:newdetect}, killed by SmartCheck but not Slither:
+This detector is actually split into two detectors, one for this serious issue, and an informational/stylistic detector that flags code such as {\tt if (x == true)}, which is merely difficult to read.
 
@@
 For example, if {\tt x} is an unsigned integer type, the comparison {\tt x >= 0} is always true and {\tt x < 0} is always false.
-This detector is a generalization of the SmartCheck detector \url{https://github.com/smartdec/smartcheck/blob/master/rule\_descriptions/SOLIDITY\_UINT\_CANT\_BE\_NEGATIVE/}, modified to actually compute the ranges of types and identify more general instances of tautological comparisons, e.g.
-{\tt y < 512} where {\tt y}'s type is {\tt int8}.
-Again, the problem was easily identified by mutants killed by SmartCheck but not Slither.
+This detector is a generalization of the SmartCheck detector \url{https://github.com/smartdec/smartcheck/blob/master/rule\_descriptions/SOLIDITY\_UINT\_CANT\_BE\_NEGATIVE/}, modified to actually compute the ranges of types and identify other cases such as {\tt y < 512} where {\tt y}'s type is {\tt int8}.
 
-\item {\bf Loss of precision:}  Solidity only supports integer types.
-This means that performing division before multiplication can introduce rounding that is not present when the multiplication is performed first.
-This is a fairly important problem, given the frequency with which Solidity code performs financial calculations where maximum precision is desired.
+\item {\bf Loss of precision:}  Solidity only supports integer types, so performing division before multiplication can introduce avoidable rounding.
+This is a fairly important problem, given Solidity code often performs critical financial calculations.
 SmartCheck provides a detector for such precision losses \url{https://github.com/smartdec/smartcheck/blob/master/rule\_descriptions/SOLIDITY\_DIV\_MUL/}.
@@
 Substantial changes in results therefore require adding very effective (for mutants) detectors that seldom trigger for correct code (or at least trigger much less than for mutants).
-``Cheating'' with respect to a mutation benchmark is thus, we hope, difficult without actual major tool improvements.
+``Cheating'' with respect to a mutation benchmark is thus, we hope, very difficult.
 
@@
 We still, due to the huge size of the source files and thus number of mutants (and time to compile full projects), restricted our analysis to files where Universal Mutator's implementation of TCE \cite{TCE} for Java was useful, i.e.
-files that could be compiled and the bytecode compared without processing the full project, leaving us with just over 70,000 mutants, ranging from 136 to 10,016 per project.
+individual files that could be compiled and the bytecode compared, leaving us with just over 70,000 mutants, ranging from 136 to 10,016 per project.
 
@@
   \centering
-  \includegraphics[width=0.35\columnwidth]{java.png}
+  \includegraphics[width=0.30\columnwidth]{java.png}
   \caption{Mutants killed by Java static analysis tools.}
@@
 Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools, and Table \ref{tab:scorejava} provides numeric results for projects and clean projects, respectively.
-At the individual project level, Infer was never best; PMD had a better raw score than SpotBugs for 10 of the 15 projects, but SpotBug had a better ratio for 11 of the projects.
-There is perhaps a real tradeoff in verbosity and precision here.
-Standard deviation in project mutation scores was 0.05 for SpotBugs, 0.04 for PMD, and 0 for Infer.
+At the individual project level, Infer was never best; PMD had a better raw score than SpotBugs for 10/15 projects, but SpotBug had a better \emph{ratio} for 11.
+There is likely a tradeoff between verbosity and precision.
+Standard deviation in project scores was 0.05 for SpotBugs, 0.04 for PMD, and 0 for Infer.
 
@@
 Both produce a large number of unique detections, though PMD produces about 20\% more than SpotBugs.
-Infer on the other hand, is only able to detect two mutants, but these are unique.
-Unfortunately, these mutants only served to indicate that we were actually running Infer in such a way that it \emph{could} detect problems:  both were concurrency warnings, but the code change in both cases was simply removing a (semantically inoperative) {\tt \@Override} annotation.
+Infer on the other hand, is only able to detect two mutants, but these are unique; both were, however, spurious concurrency warnings.
 It may be that the diff sizes in our code were simply too small for Infer's approach.
@@
 Infer is closer to a model checker focusing on resource leaks than a truly general-purpose tool, arguably.
-In fact, we suspect Infer would perform much better if we crafted complex mutation operators targeting some important subtle Java bugs; it is still fair to say that Infer is probably not a good \emph{general-purpose} static analysis tool for Java.
-Note that we did not use Infer's experimental, non-standard, detectors, however.
-
-For {\bf RQ4} there were very few clean projects, but the mutation scores for Infer were unchanged, and for SpotBugs they were worse.
-We see no evidence that non-clean code is a source of degradation in mutation detection.
+In fact, we suspect Infer would perform much better if we used mutation operators targeting some important subtle Java bugs; Infer is probably not a good \emph{general-purpose} static analysis tool for Java.
+Note that we did not use Infer's experimental detectors.
+For {\bf RQ4} there were very few clean projects, but we see no evidence that non-clean code is a source of degradation in mutation detection.
 
@@
 
-We compared three widely used and well-known Python tools:  Pylint \url{https://www.pylint.org/} (probably the most widely used of Python bug finding tools), pyflakes \url{https://pypi.org/project/pyflakes/}, designed to be faster, lighter-weight, and more focused on bugs (without configuration) than Pylint, and PyChecker \url{http://pychecker.sourceforge.net/}, an older, but still used (e.g., we found numerous {\tt .travis.yml} files installing and running PyChecker), tool.
+We compared three widely used and well-known Python tools:  Pylint \url{https://www.pylint.org/} (probably the most widely used of Python bug finding tools), pyflakes \url{https://pypi.org/project/pyflakes/}, designed to be faster, lighter-weight, and more focused on bugs (without configuration) than Pylint, and PyChecker \url{http://pychecker.sourceforge.net/}, an older, but still used tool.
 
@@
   \centering
-  \includegraphics[width=0.35\columnwidth]{python_new.png}
+  \includegraphics[width=0.30\columnwidth]{python_new.png}
   \caption{Mutants killed by Python static analysis tools.}
@@
 From the diagram, one might think that pyflakes acts, to some extent, as a less verbose ``subset'' of Pylint, in that most mutants detected by pyflakes are also detected by Pylint (however, the nearly 3,500 killed mutants unique to pyflakes suggest it is a useful tool, perhaps most useful after problems also reported by Pylint are fixed).
-PyChecker performs poorly in part because it crashed (due to changes in Python since the last update to the tool) for 19 of the 25 projects; however, as the clean scores show, it also performed poorly on programs where it worked.
+PyChecker performs poorly in part because it crashed (due to changes in Python since the last update to the tool) for 19 of the 25 projects; however, it also performed poorly on programs where it worked.
 
@@
 Pyflakes is also well liked, and is considered much less verbose than Pylint; this was definitely reflected in our results, where Pyflakes underperformed in raw kills, but had a much better mutant ratio than Pylint.
-For {\bf RQ4}, Pylint performed significantly better on clean projects.
-Performance for the other two tools was essentially unchanged over projects for which they were clean, again failing to strongly support the possibility that our definition of killing a mutant is too restrictive.
+For {\bf RQ4}, Pylint performed significantly better on clean projects.  Performance for the other two tools was essentially unchanged.
 

### 2021-04-20 introduction.tex
target ISSRE

@@
 
-Combining a differential approach and mutation analysis offers a novel way to compare static analysis tools, one useful to users wishing to select a good tool or set of tools, to researchers interested in the impact of precision/soundness trade-offs or different intermediate languages, and to developers of static analysis tools hoping to improve their tools by examining missed faults.
+Combining a differential approach and mutation analysis offers a novel way to compare static analysis tools, one useful to users wishing to select a good tool or set of tools, to researchers interested in the impact of precision/soundness trade-offs or different intermediate languages, and to developers of static analysis tools hoping to improve their tools.
 
@@
   \centering
-  \includegraphics[width=0.35\columnwidth]{example.png}
+  \includegraphics[width=0.30\columnwidth]{example.png}
   \caption{Mutants killed by three static analysis tools.}
@@
 If we restrict our analysis to findings with a severity greater than informational, SmartCheck detects no mutants of the contract, while Slither still reports that some mutants \emph{allow an arbitrary caller to cause the contract to destroy itself}.
-Given that both tools, ignoring informational results, detect no problems with the original code, and only Slither detects any problems with the mutated code, we can say that Slither performs better for static analysis of mutants for this contract.
+Given that both tools, ignoring informational results, detect no problems with the original code, and only Slither detects any problems with the mutants, we can say that Slither performs better for this contract.
 
@@
 Code deletions, the most common kind of mutation by far, leave no ``trace'' for a tool to match against, but only an omission, so cannot be subject to this concern.
-Changing arithmetic and comparison operators and numeric constants (incrementing, decrementing, or changing to 0 or 1) account for most of the non-statement-deletion mutants.
-It is difficult to imagine how any useful tool could unfairly identify these kinds of mutants.
+Changing arithmetic and comparison operators and numeric constants (incrementing, decrementing, or changing to 0 or 1) account for most of the non-statement-deletion mutants, and it is difficult to imagine how any tool could unfairly identify these.
 
@@
 \begin{itemize}[labelsep=3pt,leftmargin=12pt]
-\item We propose a differential approach to comparing static analysis tools based on the insight that program mutants provide an automated source of simple, easy-to-understand program changes that are likely faults.
+\item We propose a differential approach to comparing static analysis tools based on the insight that program mutants are easy-to-understand, likely-faulty, program changes.
 \item We propose a definition of mutant killing for static analysis.
@@
 \item We apply our method to an extensive, in-depth comparison of three Solidity smart contract analysis tools, and show how prioritization allowed us to easily identify (and build) three new detectors for the most effective of these tools.
-\item We further provide results for comparisons of popular Java and Python static analysis tools, showing the general usefulness of our methods, and giving a new picture of the comparative effectiveness of these tools.
+\item We also provide results for popular Java and Python static analysis tools, further demonstrating our approach and showing strengths and weaknesses of these tools.
 \end{itemize}
 
- While there are certainly limitations to using differential mutation analysis to compare and extend static analysis tools, the method scales to basing comparisons on large numbers of real software source files and very large numbers of faults, but still has some of the advantages of humans establishing ground truth for tool findings.
+ While there are limitations to using differential mutation analysis to compare/improve static analysis tools, it scales to basing comparisons on many real software source files and very many ``faults,'' but still offers some of the advantages of having humans establish ground truth.
 
@@
 \item Compute, for each tool, the \emph{mutant ratio}:  the mutation score ($\frac{|\mathit{killed}|}{|\mathit{mutants}|}$) divided by (mean) baseline.
-If the (mean) baseline is zero, use a baseline equal to either one or the lowest non-zero baseline for any tool in the comparison set\footnote{Unless the number of programs used for evaluation is small, a zero mean baseline for a tool is extremely unlikely to arise in practice.}.
+If it is zero, use a baseline equal to either one or the lowest non-zero baseline for any tool in the comparison set\footnote{This problem seldom arises in practice.}.
 \item (Optional): Discard all mutants not killed by at least one tool and all mutants killed by all tools.  What remains allows \emph{differential} analysis.
-Examine the remaining mutants in the difference in \emph{prioritized} order to qualitatively understand tool differences, or improve a tool.
+Examine the remaining mutants in the difference in \emph{prioritized} order.
 \end{enumerate}
@@
 code between the locations of two mutants.  If the mutants are to
-different files, this adds 0.5 to the distance; it also adds 0.25
+different files, this adds 0.5; it also adds 0.025
 times the number of source lines separating the two mutants if they
-are in the same file, divided by 10, but caps the amount added at
+are in the same file, but caps the amount added at
 0.25.

### 2021-04-20 relatedwork.tex
target ISSRE

@@
 in contrast to their introduction of checks that may be guaranteed to pass).
-Most importantly, however, their approach \emph{only applies to tools that check assertions}, rather than the more common case of tools that identify bad code patterns (e.g.
-all tools we considered in this paper).
+Most importantly, however, their approach \emph{only applies to tools that check assertions}, rather than the much more common case of tools that identify bad code patterns.

### 2021-04-20 conclusions.tex
other changes

@@
 learning \cite{scholkopf1998nonlinear,tipping1999probabilistic}) are
-the most obvious and interesting possibilities for a better metric.
-Finally, mutant prioritization should be applicable to improving
-software testing, as well \cite{groce2018verified}.
+the most obvious and interesting possibilities.
+%for a better metric.
+%Finally, mutant prioritization should be applicable to improving
+%software testing, as well \cite{groce2018verified}.
 

### 2021-04-20 prioritizing.tex
other changes

@@
 code between the locations of two mutants.  If the mutants are to
-different files, this adds 0.5 to the distance; it also adds 0.25
+different files, this adds 0.5; it also adds 0.25
 times the number of source lines separating the two mutants if they

### 2021-08-08 abstract.tex
QRS

--- 2020-01-24 abstract.tex
@@
-Many programming languages offer multiple static analysis tools that offer to detect faults in code without executing it.
-Understanding the strengths and weaknesses of tools, and performing direct comparisons of their effectiveness is difficult; it usually involves either manual examination of differing warnings on real code, or the bias-prone construction of artificial test cases.
-In practice, comparisons tend to be limited to superficial, anecdotal discussions in the informal literature (e.g., blog posts by software developers), or purely research-community-oriented evaluations made by the authors of new tools seeking to publish their results.
-This paper proposes a novel automated approach to comparing static analysis tools, based on producing \emph{mutants} of real code, and comparing mutation detection rates for tools to their warning rates on the original code.
+Static analysis tools attempt to detect faults in code without executing it.
+Understanding the strengths and weaknesses of such tools, and performing direct comparisons of their effectiveness, is difficult, involving either manual examination of differing warnings on real code, or the bias-prone construction of artificial test cases.
+This paper proposes a novel automated approach to comparing static analysis tools, based on producing mutants of real code, and comparing detection rates over these mutants.
 In addition to making tool differences quantitatively observable without extensive manual effort, this approach offers a new way to detect and fix omissions in a static analysis tool's set of detectors.
-We present an extensive comparison of three well-known Solidity smart contract static analysis tools, and show how using an automatic prioritization of our results allowed us to add three new detectors to the best of the tools.
+We present an extensive comparison of three smart contract static analysis tools, and show how our approach allowed us to add three effective new detectors to the best of these.
 We also evaluate popular Java and Python static analysis tools and discuss their strengths and weaknesses.

### 2021-10-26 experiments.tex
submitted
fix small issues

@@
 All three detectors produce some true positives (actual problems, though not always exploitable) in benchmark contracts, have acceptably low false positive rates, and were deemed valuable enough to include as non-informational (medium severity) detectors.
-The first mutants in prioritized rank exhibiting the issues, shown above, were the 2nd, 9th, and 12th non-statement-deletion mutants ranked for SmartCheck, out of over 800 such mutants.
+Moreover, the loss of precision detector has, to date, detected two real-world security vulnerabilities:  \url{https://github.com/crytic/slither/blob/master/trophies.md} --- search for ``Dangerous divide before multiply operations.''
+
+The first mutants in prioritized rank exhibiting the issues inspiring these detectors, shown above, were the 2nd, 9th, and 12th non-statement-deletion mutants ranked for SmartCheck, out of over 800 such mutants.
 Using our prioritization, it was possible to identify these issues by examining fewer than 20 unkilled mutants.
@@
     SpotBugs & 28.93 & 14.00 & 0.07 & 0.07 & 0.002 \\
-    Clean (3) & - & - & 0.05 & 0.06 - \\
+    Clean (3) & - & - & 0.05 & 0.06 & - \\
     \hline
@@
     Infer & 11.60 & 3.00 & 0.00 & 0.00 &  0.000 \\
-    Clean (6) & - & - & 0.00 & 0.00 \\
+    Clean (6) & - & - & 0.00 & 0.00 & - \\
     \hline

### 2021-10-26 introduction.tex
fix small issues

@@
 Slither, but not SmartCheck, also detects replacing the assignment of {\tt storedData} in {\tt set} with either a {\tt selfdestruct} or {\tt revert}, or simply removing it altogether.
-SmartCheck, on the other hand, detects removing the {\tt return} in {\tt get} or replacing it with a {\tt revert}, or removing the {\tt public} visibility modifier for {\tt get} \footnote{Slither's ``missing return'' detector is only available in the private version of slither, or through the {\tt crytic} service provided by Trail of Bits.}.
+SmartCheck, on the other hand, detects removing the {\tt return} in {\tt get} or replacing it with a {\tt revert}, or removing the {\tt public} visibility modifier for {\tt get} \footnote{Slither's ``missing return'' detector was only available in the private version of Slither, at the time we performed these experiments.}.
 If we restrict our analysis to findings with a severity greater than informational, SmartCheck detects no mutants of the contract, while Slither still reports that some mutants \emph{allow an arbitrary caller to cause the contract to destroy itself}.
